---
title: "DHC_percentage"
author: "Camila Vargas e Ignacia Rivera"
date: "8 de diciembre de 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script builds a dataframe with the percentage of yield of each chilean species for direct humand consumption (DHC). We first looked into the database provided by Peru (Product yields by species.xlsx). We looked into all the species of this databses relevant to Chile. For the remaining species we used...XXXX

## Setup
```{r}
library(tidyverse)
library(here)
library(readxl)
library(stringr)

```

## Peru's database

```{r}
## read in the Peru's data set
body_part_ratios <- read_excel(here("information/raw_databases/product_yield/product_yields_sp.xlsx"), sheet = "A", range = "A1:D167") %>% 
  rename(species_en = "Common name (English)", sc_name = "Scientific name", body_part = "Body parts", yield = "Body part to total weight (%)")

##Calculate mean yield form all body parts
mean_yield_body_parts <- body_part_ratios %>% 
  group_by(species_en, sc_name) %>% 
  summarise(mean_yield = mean(yield)) ## yield is in percentage

##Second data set witn min and max per body parts
body_part_min_max <- read_excel(here("information/raw_databases/product_yield/product_yields_sp.xlsx"), sheet = "B", range = "A1:E99") %>% 
  rename(species_en = "Common name (Eng)", sc_name = "Scientific name", body_part = "Presentations", min_yield = "Yield (% of whole fish weight)", max_yield = "X__1") %>% 
  filter(!is.na(species_en)) %>% 
  mutate(min_yield = as.numeric(min_yield)) %>% 
  mutate(max_yield = as.numeric(max_yield))
  


##Calculate mean yield from all body parts

mean_yield_body_parts_min_max <- body_part_min_max %>% 
  group_by(species_en, sc_name) %>% 
  summarise(mean_yield_min = mean(min_yield),
            mean_yield_max = mean(max_yield)) %>%  ## yield is in percentage
  mutate(mean_yield = (mean_yield_min + mean_yield_max)/2) 


##NOTE:
#setdiff(mean_yield_body_parts$sc_name, mean_yield_body_parts_min_max$sc_name)
#What species of mean_yield_body_parts are not in mean_yield_body_parts_min_max?
# 9 species difference so mean_yield_body_parts is more complete.


##Third data set with overall yield

yield_c <- read_excel(here("information/raw_databases/product_yield/product_yields_sp.xlsx"), sheet = "C", range = "A1:F111") %>% 
  select(species_en = "Common name (Eng)", sc_name = "Scientific name", min_yield = "Yield range (min)", max_yield = "Yield range (max)") %>% 
  filter(!is.na(species_en)) %>% 
  mutate(min_yield = as.numeric(min_yield)) %>% 
  mutate(max_yield = as.numeric(max_yield))

##Calculate mean between min adn max

mean_yield_c <- yield_c %>% 
mutate(mean_yield = (min_yield + max_yield)/2)

setdiff(mean_yield_c$sc_name, mean_yield_body_parts$sc_name) ## 89 differences
setdiff(mean_yield_body_parts$sc_name, mean_yield_c$sc_name) ##8 differences

##Relevant species for Chile from mean_yield_body_part-- need this three species to the list

mean_yield_bp_cl <- mean_yield_body_parts %>% 
  filter(sc_name %in% c("Mesodesma donacium","Oncorhynchus mykiss", "Sardinops sagax")) %>% 
  mutate(mean_yield = mean_yield/100) # from precentge to proportion



## Save yield dataset

write.csv(mean_yield_c, here("clean_databases/intermediate/yield_DHC.csv"), row.names=F)


```


##Compare with list of chilean species

```{r}
DHC_yield <- read_csv("clean_databases/intermediate/yield_DHC.csv") %>% 
  select(-min_yield, -max_yield)

sp_cl <- read.csv(here("clean_databases/intermediate/sp_model.csv"))


match_DHC <- sp_cl %>% 
  left_join(DHC_yield, by = "sc_name")

##Save match_DHC in order to complete the database by hand in excel considering similar species or other data source
write.csv(match_DHC, here("clean_databases/intermediate/sp_cl_DHC.csv"), row.names = F)


```


