---
title: "Cleaning landings data SERNAPESCA"
author: "Ignacia Rivera"
date: "August 17, 2018"
output: pdf_document
---
This script uses raw data of landings from SERNAPESCA to generate two databases:

1) **mean_landings_SERNAPESCA_month**: contains the mean tons per species per subsector and month aggregated at the national level. The mean was calculated across 2014-2016.

2) **mean_landings_SERNAPESCA_region**: contains the mean tons per species per subsector and month aggregated at the anual level. The mean was calculated across 2014-2016.

Although, SERNAPESCA counts with databases at a higher spatial resolution (ports), this databases cannot be matched with the specific contribution of each subsector.  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(magrittr)
library(dplyr)
library(tidyr)
library(readxl)
```

```{r importing landings data for 2014, 2015, 2016 per species, subsector, month and region}

#################################################################################
# Importing and formating raw artisanal landings 2014, 2015, 2016. 
#################################################################################

# Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Artesanal. Accessed August 15, 2016. Data for 2017 was not available online but was requested to SERNAPESCA. They answered it will be uploaded by September. 

# We need still need to know: 

## 1) if - in SERNAPESCA databases means 0 or NA. Probably 0.  
## 2) if landings in international waters means that fisher are fishing in international waters or  ## that they are landing in international waters. 

# I'm waiting for an answer from SERNAPESCA under the request 460245718
##################################################################################

# Artisanal landings per month

artisanal_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2014.xls"), sheet = "des_art_mes", range = "A5:M147", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2015.xls"), sheet = "des_art_mes", range = "A5:M150", na = "-")%>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2016.xls"), sheet = "des_art_mes", range = "A5:M147", na = "-")%>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Artisanal landings per region

artisanal_region_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2014.xls"), sheet = "des_art_region", range = "A5:O153", na = "-") %>% 
  gather(key = region, value = tons, 2:15) %>% 
  mutate (subsector = 'artisanal', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2015.xls"), sheet = "des_art_region", range = "A5:O150", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'artisanal', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2016.xls"), sheet = "des_art_region", range = "A5:O147", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'artisanal', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Artisanal landings international waters per month

artisanal_iw_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2014.xls"), sheet = "des_art_ai_mes", range = "A5:M11", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2015.xls"), sheet = "des_art_ai_mes", range = "A5:M9", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2016.xls"), sheet = "des_art_ai_mes", range = "A5:M8", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Artisanal landings international waters per region

artisanal_iw_region_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2014.xls"), sheet = "des_art_ai_reg", range = "A5:O11", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'artisanal', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2015.xls"), sheet = "des_art_ai_region", range = "A5:P9", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'artisanal', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2016.xls"), sheet = "des_art_ai_region", range = "A5:P8", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'artisanal', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

##################################################################################
# Importing and formating raw industrial landings 2014, 2015, 2016. 
##################################################################################

# Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Industrial. Accessed August 15, 2016. Data for 2014 and 2017 was not available online but was requested to SERNAPESCA. They answered that 2017 should be uploaded by September.

# We need still need to know, 

## 1) if - in SERNAPESCA databases means 0 or NA. Probably 0.  
## 2) if landings from industrials include fabric vessels. Probably not. 
## 3) if landings in international waters means that fisher are fishing in international waters or  ## that they are landing in international waters. 

# I'm waiting for an answer from SERNAPESCA under the request 460245718
################################################################################

# Industrial landings per month

industrial_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2014.xls"), sheet = "des_ind_mes", range = "A5:M52", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2015.xls"), sheet = "des_ind_mes", range = "A5:M52", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2016.xls"), sheet = "des_ind_mes", range = "A5:M48", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Industrial landings per region (missing for 2014)

industrial_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_per_species_region_2015.xls"), sheet = "Des_ind_region", range = "A5:O52", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'industrial', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_per_species_region_2016.xls"), sheet = "Des_ind_region", range = "A5:O48", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'industrial', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Industrial landings international waters per month

industrial_iw_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2014.xls"), sheet = "industriales_ai_mes", range = "A5:M13", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2015.xls"), sheet = "ind_ai_mes", range = "A5:M12", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2016.xls"), sheet = "ind_ai_mes", range = "A5:M13", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Industrial landings international waters per region

industrial_iw_region_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2014.xls"), sheet = "industriales_ai_region", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2015.xls"), sheet = "ind_ai_region", range = "A5:O12", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2016.xls"), sheet = "ind_ai_region", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

################################################################################
# Importing and formating raw landings from fabric vessels 2014, 2015, 2016. 
#################################################################################

# Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Industrial. Accessed August 15, 2016. Data for 2017 was not available online but was requested to SERNAPESCA. Supossed to be uploaded by September.

# We need still need to know, 

## 1) if - in SERNAPESCA databases means 0 or NA. Probably 0.  
## 2) if landings from industrials include fabric vessels. Probably not. 
## 3) if landings in international waters means that fisher are fishing in international waters or  ## that they are landing in international waters. 

# I'm waiting for an answer from SERNAPESCA under the request 460245718
################################################################################

# Landings from fabric vessels (fv) per month

fv_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2014.xls"), sheet = "des_bf_nac_mes", range = "A5:M19", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2015.xls"), sheet = "bf_nac_mes", range = "A5:M20", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2016.xls"), sheet = "bf_nac_mes", range = "A5:M21", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Landings from fabric vessels (fv) per region

fv_region_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2014.xls"), sheet = "des_bf_nac_reg", range = "A5:O19", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2015.xls"), sheet = "bf_nac_region", range = "A5:O20", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2016.xls"), sheet = "bf_nac_region", range = "A5:O21", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


# Landings from fabric vessels (fv) in international waters per month

fv_iw_month_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2014.xls"), sheet = "bf_ai_mes", range = "A5:M13", na = "-")%>%   gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_month_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2015.xls"), sheet = "bf_ai_mes", range = "A5:M11", na = "-")%>%   gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_month_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2016.xls"), sheet = "bf_ai_mes", range = "A5:M23", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

# Fabric vessels landings international waters per region

fv_iw_region_2014 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2014.xls"), sheet = "bf_ai_reg", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'fabric_vessel', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_region_2015 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2015.xls"), sheet = "bf_ai_region", range = "A5:O11", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'fabric_vessel', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_region_2016 <- read_excel(here("raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2016.xls"), sheet = "bf_ai_region", range = "A5:P23", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'fabric_vessel', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

#######################################################################
# Combining all landings data for 2014, 2015, 2016. 
#######################################################################


landings_SERNAPESCA <- rbind(artisanal_month_2014, artisanal_month_2015, artisanal_month_2016, artisanal_region_2014, artisanal_region_2015, artisanal_region_2015, artisanal_region_2016, artisanal_iw_month_2014, artisanal_iw_month_2015, artisanal_iw_month_2016, artisanal_iw_region_2014,artisanal_iw_region_2015, artisanal_iw_region_2016, industrial_month_2014, industrial_month_2015, industrial_month_2016, industrial_region_2015, industrial_region_2015, industrial_region_2016, industrial_iw_month_2014, industrial_iw_month_2015, industrial_iw_month_2016, industrial_iw_region_2014, industrial_iw_region_2015, industrial_iw_region_2016, fv_month_2014, fv_month_2015, fv_month_2016, fv_region_2014, fv_region_2015, fv_region_2016, fv_iw_month_2014, fv_iw_month_2015, fv_iw_month_2016, fv_iw_region_2014, fv_iw_region_2015, fv_iw_region_2016) %>% 
  filter(!(species =='NA'))

```

```{r creating landings_SERNAPESCA_month}

#######################################################################
#  Generating landings_SERNAPESCA_month.csv
#######################################################################

# calculates mean landing per species, subsector, waters and month

mean_landings_SERNAPESCA_month <- landings_SERNAPESCA %>% 
  filter(region == 'national') %>%  # removing anual estimates per region
  group_by(subsector, species, month, international_waters) %>% 
    summarise(tons = mean(tons, na.rm = TRUE))

# Writes cleaned database to clean_databases

write.csv(mean_landings_SERNAPESCA_month,  here("clean_databases/mean_landings_SERNAPESCA_month.csv"), row.names=T)

```

```{r creating landings_SERNAPESCA_region}

#######################################################################
#  Generating landings_SERNAPESCA_region.csv
#######################################################################

# calculates mean landing per species, subsector, waters and region

mean_landings_SERNAPESCA_region <- landings_SERNAPESCA %>% 
  filter(month == 'anual') %>%  # removing national estimates per month
  group_by(subsector, species, region, international_waters) %>% 
    summarise(tons = mean(tons, na.rm = TRUE))

# Writes cleaned database to clean_databases

write.csv(mean_landings_SERNAPESCA_region,  here("clean_databases/mean_landings_SERNAPESCA_region.csv"), row.names=T)

```
