---
title: "Cleaning landings data SERNAPESCA"
author: "Ignacia Rivera"
date: "August 17, 2018"
output: pdf_document
---
## Description

This script uses raw data of landings from different sectors from SERNAPESCA to generate:

1) **landings_SERNAPESCA**: contains tons per species per subsector per month or region between 2014-2017.

Although, SERNAPESCA counts with databases at a higher spatial resolution (ports), this databases cannot be matched with the specific contribution of each subsector. 

## Q & A with SERNAPESCA regarding these databases

Q: Algunas de las celdas en los datos de desembarque presentan el simbolo -. Quiero aclarar si esto significa un valor de cero toneladas o que no se pudieron tomar datos para esa especie.

A: El simbolo – significa que las toneladas desembarcadas fueron 0 o menor que 0.5, por lo tanto en los Anuarios no se consideran

Q: Las estadisticas de desembarque para el sector industrial disponible en la seccion Infomes y Datos/ Estadísticas/ Pesca Industrial incluye los datos, ubicados en a misma sección, para barcos fabrica o son estos número independientes?

A: La operación de barcos fábrica y de barcos industriales se encuentran en la misma sección porque ambos se consideran naves industriales. Son independientes

Q: Tanto para barcos fabrica, artesanales e industriales en general, existen bases de datos para desembarques en aguas internacionales. Esto quiere decir, que son desembarques de capturas en aguas nacionales en puertos internacionales ó, desembarques de capturas en aguas internacionales en puertos nacionales?

A: Corresponden a capturas realizadas en aguas internacionales y desembarcadas en puertos chilenos.

Q: Las estadisticas de desembarque para de areas de manejo, estan inlcuidas en los desembarques del sector artesanal o son independientes?

Q:  A que corresponde la columna AI en las bases de datos de desembarque de barcos fabrica en aguas internacionales por mes en 2016 y 2017? Por que no esta en anos anteriores?


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here) #install.packages('here')
library(magrittr) #install.packages('magrittr')
library(dplyr) #install.packages('dplyr')
library(tidyr) #install.packages('tidyr')
library(readxl) #install.packages('readxl')


```

## Importing and formating raw artisanal landings 2014, 2015, 2016, 2017

Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Artesanal. 

**Data downloaded** August 15, 2018 (data years 2014-2016), September 24, 2018 (data 2017)

**Data range** 2014 -2017

```{r importing artisanal landings data for 2014, 2015, 2016 per species, subsector, month and region}

## Artisanal landings per month

artisanal_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2014.xls"), sheet = "des_art_mes", range = "A5:M147", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


artisanal_month_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2015.xls"), sheet = "des_art_mes", range = "A5:M150", na = "-")%>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_month_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2016.xls"), sheet = "des_art_mes", range = "A5:M147", na = "-")%>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_month_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_month_2017.xlsx"), sheet = "des_art_mes", range = "A5:M153", na = "-")%>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'artisanal', year = 2017, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

## Artisanal landings per region

artisanal_region_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2014.xls"), sheet = "des_art_region", range = "A5:O153", na = "-") %>% 
  gather(key = region, value = tons, 2:15) %>% 
  mutate (subsector = 'artisanal', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_region_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2015.xls"), sheet = "des_art_region", range = "A5:O150", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'artisanal', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_region_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2016.xls"), sheet = "des_art_region", range = "A5:O147", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'artisanal', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_region_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_per_species_region_2017.xlsx"), sheet = "des_art_region", range = "A5:O153", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'artisanal', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Artisanal landings international waters per month

artisanal_iw_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2014.xls"), sheet = "des_art_ai_mes", range = "A5:M11", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_month_2015 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2015.xls"), sheet = "des_art_ai_mes", range = "A5:M9", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_month_2016 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2016.xls"), sheet = "des_art_ai_mes", range = "A5:M8", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_month_2017 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2017.xlsx"), sheet = "des_art_ai_mes", range = "A5:M6", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'artisanal', year = 2017, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Artisanal landings international waters per region

artisanal_iw_region_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2014.xls"), sheet = "des_art_ai_reg", range = "A5:O11", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'artisanal', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_region_2015 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2015.xls"), sheet = "des_art_ai_region", range = "A5:P9", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'artisanal', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_region_2016 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2016.xls"), sheet = "des_art_ai_region", range = "A5:P8", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'artisanal', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

artisanal_iw_region_2017 <-read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_artisanal_international_waters_per_species_region_month_2017.xlsx"), sheet = "des_art_ai_region", range = "A5:P6", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'artisanal', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


```



## Importing and formating raw industrial landings 2014, 2015, 2016, 2017. 


Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Industrial. Accessed August 15, 2018. Data for 2014 and 2017 was not available online but was requested to SERNAPESCA. 
Data for 2017 was downloaded September 24, 2018

```{r importing industrial landings data for 2014, 2015, 2016 per species, subsector, month and region} 
# Industrial landings per month

industrial_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2014.xls"), sheet = "des_ind_mes", range = "A5:M52", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_month_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2015.xls"), sheet = "des_ind_mes", range = "A5:M52", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_month_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2016.xls"), sheet = "des_ind_mes", range = "A5:M48", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_month_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_month_2017.xlsx"), sheet = "des_ind_mes", range = "A5:M48", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'industrial', year = 2017, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

## Industrial landings per region (missing for 2014)

industrial_region_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_region_2015.xls"), sheet = "Des_ind_region", range = "A5:O52", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'industrial', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_region_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_region_2016.xls"), sheet = "Des_ind_region", range = "A5:O48", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'industrial', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_region_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_per_species_region_2017.xls"), sheet = "Des_ind_region", range = "A5:O48", na = "-")%>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'industrial', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 0)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Industrial landings international waters per month

industrial_iw_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2014.xls"), sheet = "industriales_ai_mes", range = "A5:M13", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_month_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2015.xls"), sheet = "ind_ai_mes", range = "A5:M12", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_month_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2016.xls"), sheet = "ind_ai_mes", range = "A5:M13", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


industrial_iw_month_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2017.xlsx"), sheet = "ind_ai_mes", range = "A5:M12", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'industrial', year = 2017, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)



## Industrial landings international waters per region

industrial_iw_region_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2014.xls"), sheet = "industriales_ai_region", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_region_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2015.xls"), sheet = "ind_ai_region", range = "A5:O12", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_region_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2016.xls"), sheet = "ind_ai_region", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

industrial_iw_region_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_industrial_international_waters_per_species_region_month_2017.xlsx"), sheet = "ind_ai_region", range = "A5:O12", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'industrial', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

```

## Importing and formating raw landings from fabric vessels 2014, 2015, 2016, 2017. 

Datasets are from SERNAPESCA/ Informes y Datos/ Estadisticas/ Pesca Industrial. Accessed August 15, 2018 (for data from 2014-16) and on September 24, 2018 for 2017 data.

```{r importing and formating raw landings from fabric vessels 2014, 2015, 2016}
## Landings from fabric vessels (fv) per month

fv_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2014.xls"), sheet = "des_bf_nac_mes", range = "A5:M19", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2014, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_month_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2015.xls"), sheet = "bf_nac_mes", range = "A5:M20", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2015, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_month_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2016.xls"), sheet = "bf_nac_mes", range = "A5:M21", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2016, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_month_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2017.xlsx"), sheet = "bf_nac_mes", range = "A5:M20", na = "-") %>% 
  gather(key = month, value = tons, 2:13)%>% 
  mutate (subsector = 'fabric_vessel', year = 2017, region = 'national', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Landings from fabric vessels (fv) per region

fv_region_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2014.xls"), sheet = "des_bf_nac_reg", range = "A5:O19", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_region_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2015.xls"), sheet = "bf_nac_region", range = "A5:O20", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_region_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2016.xls"), sheet = "bf_nac_region", range = "A5:O21", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_region_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_per_species_region_month_2017.xlsx"), sheet = "bf_nac_region", range = "A5:O20", na = "-") %>% 
  gather(key = region, value = tons, 2:15)%>% 
  mutate (subsector = 'fabric_vessel', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 0) %>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Landings from fabric vessels (fv) in international waters per month

fv_iw_month_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2014.xls"), sheet = "bf_ai_mes", range = "A5:M13", na = "-")%>%   gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2014, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_month_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2015.xls"), sheet = "bf_ai_mes", range = "A5:M11", na = "-")%>%   gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2015, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_month_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2016.xls"), sheet = "bf_ai_mes", range = "A5:M23", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2016, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_month_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2017.xlsx"), sheet = "bf_ai_mes", range = "A5:M12", na = "-") %>% 
  gather(key = month, value = tons,2:13) %>% 
  mutate (subsector = 'fabric_vessel', year = 2017, region = 'national', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


## Fabric vessels landings international waters per region

fv_iw_region_2014 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2014.xls"), sheet = "bf_ai_reg", range = "A5:O13", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'fabric_vessel', year = 2014, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_region_2015 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2015.xls"), sheet = "bf_ai_region", range = "A5:O11", na = "-")%>% 
  gather(key = region, value = tons,2:15) %>% 
  mutate (subsector = 'fabric_vessel', year = 2015, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)

fv_iw_region_2016 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2016.xls"), sheet = "bf_ai_region", range = "A5:P23", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'fabric_vessel', year = 2016, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


fv_iw_region_2017 <- read_excel(here("information/raw_databases/SERNAPESCA/landings/landings_fabric_vessels_international_waters_per_species_region_month_2017.xlsx"), sheet = "bf_ai_region", range = "A5:P17", na = "-")%>% 
  gather(key = region, value = tons,2:16) %>% 
  mutate (subsector = 'fabric_vessel', year = 2017, month = 'anual', species =  ESPECIE, international_waters = 1)%>% 
  select(subsector, year, region, month, species, tons, international_waters)


```


## Combining all landings data for 2014, 2015, 2016, 2017. 


```{r}

landings_SERNAPESCA <- rbind(artisanal_month_2014, artisanal_month_2015, artisanal_month_2016, artisanal_month_2017, artisanal_region_2014, artisanal_region_2015, artisanal_region_2015, artisanal_region_2016, artisanal_region_2017, artisanal_iw_month_2014, artisanal_iw_month_2015, artisanal_iw_month_2016, artisanal_iw_month_2017, artisanal_iw_region_2014,artisanal_iw_region_2015, artisanal_iw_region_2016, artisanal_iw_region_2017, industrial_month_2014, industrial_month_2015, industrial_month_2016, industrial_month_2017, industrial_region_2015, industrial_region_2016, industrial_region_2017, industrial_iw_month_2014, industrial_iw_month_2015, industrial_iw_month_2016, industrial_iw_month_2017, industrial_iw_region_2014, industrial_iw_region_2015, industrial_iw_region_2016, industrial_iw_region_2017, fv_month_2014, fv_month_2015, fv_month_2016, fv_month_2017, fv_region_2014, fv_region_2015, fv_region_2016, fv_region_2017, fv_iw_month_2014, fv_iw_month_2015, fv_iw_month_2016, fv_iw_month_2017, fv_iw_region_2014, fv_iw_region_2015, fv_iw_region_2016, fv_iw_region_2017) %>% 
  filter(!(species =='NA'))


# Writes clean and complete database to clean_databases

write.csv(landings_SERNAPESCA,  here("clean_databases/landings_SERNAPESCA.csv"), row.names=F)


```
