---
title: "Cleaning mean beach prices data SERNAPESCA"
author: "Camila Vargas"
date: "19 de octubre de 2018"
output: pdf_document
---


## Description
This script cleans raw data on mean beach prices for artisanal landings provided by SERNAPESCA. 

**Note about this database**
- Prices are in chilean pesos per ton
- Origin of information: B = boat, L = Lancha, O = other, BN = natural bank, AM = TURF
- Destination of the product: CHD = human consumption, PP = processsing plant
- Empty spaces means that the information was not available



The following file is created through this scripts:


1) **prices_art_SERNAPESCA** - This document contains...


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here) #install.packages('here')
library(magrittr) #install.packages('magrittr')
library(dplyr) #install.packages('dplyr')
library(tidyr) #install.packages('tidyr')
library(readr) #install.packages('readr')
library(purrr)#install.packages('purr')
library(stringr)
library(regex)

```


## Importing and wrangling raw acuiculture harvest

Datasets were requested to SERNAPESCA (request #460312718) and recived via email on October 22, 2018.

**Data downloaded** October 22, 2018

**Data range** 2007 -2017 


### Beach prices per year and specie

```{r importing data, beach prices artisanal}

##Creat a list with all beach prices files

beach_price_list <- list.files(here('information/raw_databases/SERNAPESCA/prices'), pattern= 'beach_price_art_.*.csv', full.names = TRUE) %>% 
  
## Read, wrangle and combine beach prices files into one dataframe

  beach_price_all <- beach_price_list %>% 
  map(read.csv, header = FALSE, skip = 6, stringsAsFactors = FALSE) %>% 
  lapply(function(x) x[1:13]) %>% ## selects 13 columns of interest
  reduce(rbind) %>% 
  rename(year =V1, 
         month= V2, 
         region = V3, 
         cove_code =V4, 
         cove_name= V5, 
         sp_code=V6, 
         specie =V7, 
         peso_ton =V8,
         humidity=V9, 
         tons= V10, 
         origin_vessel=V11, 
         origin_place=V12, 
         use=V13)

``` 
  

## Save complete raw data

```{r}

write.csv(beach_price_all, here('clean_databases/intermediate/beach_price_art_comlete_SERNAPESCA.csv'), row.names = F)

```

  
## Mean prices per year

```{r}
beach_price_per_yr <- read.csv(here('clean_databases/intermediate/beach_price_art_comlete_SERNAPESCA.csv'), stringsAsFactors = FALSE) %>%
  mutate(peso_ton = gsub(",", " ", peso_ton)) %>% 
  mutate(peso_ton= str_replace_all(peso_ton, fixed(" "), "")) %>% 
  mutate(peso_ton = as.numeric(peso_ton)) %>% 
  select(year, specie, peso_ton) %>% 
  group_by(year, specie) %>% 
  summarise(mean_price = mean(peso_ton)) %>% 
  data.frame()

## Somthing wired is happening with year 2007!! BEach_price_per_yr shows 2007.0000 in some cases.
## and there is a year lable 8.989 not sure form which data base is comming. 


```

 
