---
title: "Cleaning mean beach prices data SERNAPESCA"
author: "Camila Vargas"
date: "19 de octubre de 2018"
output: pdf_document
---


## Description
This script cleans raw data on mean beach prices for artisanal landings provided by SERNAPESCA. 

**Note about this database**
- Prices are in chilean pesos per ton
- Origin of information: B = boat, L = Lancha, O = other, BN = natural bank, AM = TURF
- Destination of the product: CHD = human consumption, PP = processsing plant
- Empty spaces means that the information was not available



The following file is created through this scripts:


1) **prices_art_SERNAPESCA** - This document contains...


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here) #install.packages('here')
library(magrittr) #install.packages('magrittr')
library(dplyr) #install.packages('dplyr')
library(tidyr) #install.packages('tidyr')
library(readxl) #install.packages('readxl')

```


## Importing and wrangling raw acuiculture harvest

Datasets were requested to SERNAPESCA (request #460312718) and recived via email on October 22, 2018.

**Data downloaded** October 22, 2018

**Data range** 2007 -2017 


### Beach prices per year and specie

```{r importing data, beach prices artisanal}






