---
title: "Cleaning mean beach prices data SERNAPESCA"
author: "Camila Vargas"
date: "19 de octubre de 2018"
output: pdf_document
---


## Description
This script cleans raw data on mean beach prices for artisanal landings provided by SERNAPESCA. 

The following files are created through this scripts:


1) **prices_art_SERNAPESCA** - This document contains...


## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here) #install.packages('here')
library(magrittr) #install.packages('magrittr')
library(dplyr) #install.packages('dplyr')
library(tidyr) #install.packages('tidyr')
library(readxl) #install.packages('readxl')

```


## Importing and wrangling raw acuiculture harvest

Datasets were downloaded from [SERNAPESCA/ Informes y Datos/ Estadisticas/ ...](ADD LINK) VER SI HAY UPDATES CON INFO 2017 

**Data downloaded** September 24, 2018

**Data range** 2006 -2016


### Beach prices per year and specie

```{r importing data, beach prices artisanal}

###NOTE: need to come back to this scrip onece we recieve the data we need from SERNAPESCA. Nª 460312718


beach_price_art <- read_excel(here("information/raw_databases/SERNAPESCA/prices/mean_anual_beach_prices_per_specie_2006_2016.xls"), sheet = "cosechas_mes", range = "A5:M25", na = "-") %>% 
  gather(key = month, value = tons, 2:13, na.rm = TRUE)%>% 
  mutate (subsector = 'aquiculture', year = 2012, region = 'national', species =  ESPECIE) %>% 
  select(subsector, year, region, month, species, tons)






