---
title: "species_group_SERNAPESCA"
author: "Camila Vargas e Ignacia Rivera"
date: "6 de noviembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script wrangles the nesesary information for creating a dataframe that groups the species of SERNAPESCA in a broder level so they can be match with the import data.

The following file is created through this script:

- XXXXXXXXX

## Setup

```{r}
library(tidyverse)
library(here)
library(pdftools) #install.packages("pdftools")
library(tabulizer) #install.packages("tabulizer")
```


## Method

In order to group species but keeping as much detail as possible at a species level, we first identified all the species on the import data base. To this list we added a group name to each specie or group of specie. This database also has scientific names when they where identified in the Chapter 16 of the Aduana codes (Chapeter 16 address iport products)

The next step is to match this information by scientific name with the list we have of specie listes by sernapesca (information/raw_databases/ADUANAS_CHILE/arancel_aduanero_nomenclatura/2016_nomina_de_especies_SERNAPESCA.pdf). The list with all SERNAPESCA species is retrieved from the exports_ADUANA.csv.

Once the match is done, we need to identify which species still do not have a group name and assign one.

## Step 1: import imports to get the initial list of sepcies
This list will determine the level of detail we will get in terms of species for the final data base that includes imports. 


```{r}
## import the imports data set to identify the broad groups where the species should be clasify
species_import <- read.csv(here("clean_databases/imports_ADUANA.csv")) %>% 
  select(name_sp, sc_name) %>% 
  arrange(name_sp)

## save into an intermediate file to use as a reference of all the species that we can find in this database.
##write.csv(species_import, here("clean_databases/intermediate/species_import.csv"), row.names = F)

```

We created a file with the specie name, specie scientific name (when possible) and group name that we assing in order to calssify each of the species of our interest.


## Step 2: Import the dataset that has the species group

This will be use to match the specie scentific names with the SERNAPESCA list from the export file.

```{r}
##Import data with species group
species_group <- read_excel(here("clean_databases/intermediate/species_import_group.xlsx"), na="NA")

  
w_sc_name <- species_group %>% 
  filter(!is.na(sc_name)) 

w_out_sc_name <- species_group %>% 
  filter(is.na(sc_name))

##Import export list and select columnes of interest
species_list <- read_excel(here("information/raw_databases/ADUANAS_CHILE/arancel_aduanero_nomenclatura/arancel_aduanero_2017/codigos_aduana_por_especie_2017.xlsx"), na="NA")%>%
  select(especie_SERNAPESCA, sc_name=nombre_cientifico_SERNAPESCA) %>% 
  unique()


##match list with scientific names with the SERNAPESCA lits according to scientific names 
specie_list_complete <- species_list %>% 
  full_join(w_sc_name, by= "sc_name")

  
## Save this intermediate file. DO NO REWRITE!! This file will be modifues in excel.
##write.csv(specie_list_complete, here("clean_databases/intermediate/species_group_complete.csv"), row.names = F)


##Save a list of species with out scientific name to have as a reference for assigning group name to all species in the SERNAPESCA list. 

##write.csv(w_out_sc_name, here("clean_databases/intermediate/specie_group_w_out_sc_name.csv"), row.names = F)
```


## Step 3: Assing a group name to all the species with out scientific name

Each specie of species_group_complete.csv has to have a group name according to the groups in specie_group_w_out_sc_name.csv
There are some species that did not sit in any import category, so we determine a new group name when ever necesary (eg: algas, tiburones)

The complete list will be fill in in excel and then read in the next chunk. 

## Step 4: Read in the complete list
Read the csv with all the group names and do the matching with the import database

```{r}

sp_groups_all <- read.csv(here("clean_databases/intermediate/species_group_complete.csv"))

##add or modify species with different SERNAPESCA Name

write.csv(sp_groups_all, here("clean_databases/species_match.csv"), row.names = F)

```

## Second approach:
KEEP TRYING UNTILL IT WORK
Read the Nomina Sernapeca in pdf to get all the species we want in our analisis.

```{r}
nomina <- pdf_text(here("information/raw_databases/ADUANAS_CHILE/arancel_aduanero_nomenclatura/2016_nomina_de_especies_SERNAPESCA.pdf")) %>% 
  strsplit("\n")
```


