---
title: "Domestic Food Consumption"
author: "Camila Vargas e Ignacia Rivera"
date: "10 de diciembre de 2018"
output: pdf_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

##Species Index
sp_index <- read.csv(here("clean_databases/sp_index.csv")) %>% 
  select(species = chl_name, sc_name)
```

In order to calculate the domestic food compsumtion we buit a model that considers landings, aquaculture harvest, exports and imports. 
The model we are going to use is:
HC = (L_hc * DHC_prop) + A + I - E
Where,
L_hc = L- NHC_m
NHC_m = amount of raw products that enter to processing facilities to produce NHC products. 

##Step 1: Calculate Landings for human consuption
Import landing and NHC_m databases in ordet to calculate landings for human comsumption.

NOTE: Match all databese with the species index.

NOTE: We are going to do de analysis at anational level because we cannot account for transportation between the countris from place the place the product was transported to where they are prossesed and finally consumed.

```{r human consumption landings}
##Import landing data 

# landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
#   filter(!region %in% c("national", "R.M.")) %>%
#   mutate( tons = as.numeric(tons)) %>% 
#   select(-month, -international_waters) %>% 
#   group_by(year, region, species) %>% 
#   summarise(tons = sum(tons, na.rm = T)) %>%  ##adds same species from different subsector
#   rename(catch = tons)
  

# ##Import raw no human comsuption data (NHC_m)
# NHC_facilities <- read.csv(here("clean_databases/NHC_m.csv")) %>% 
#   rename(NHC_m = total_NHC)
#   
# 
# ##Join datasets and calculate L_hc
# landings_hc <-  landings %>% 
#   left_join(NHC_facilities, by= c("year", "region", "sc_name")) %>% 
#   mutate(NHC_m = ifelse(is.na(NHC_m), 0, NHC_m)) %>% 
#   mutate(L_hc = catch - NHC_m)
# 
# landing_hc_nat <- landings %>% 
#   left_join(NHC_facilities, by= c("year", "region", "sc_name")) %>% 
#   mutate(NHC_m = ifelse(is.na(NHC_m), 0, NHC_m)) %>% 
#   group_by(year, sc_name, species) %>% 
#   summarise(catch_nat = sum(catch, na.rm = T),
#             NHC_m_nat = sum(NHC_m, na.rm = T)) %>% 
#   mutate(L_hc_nat = catch_nat - NHC_m_nat) %>% 
#   mutate(L_hc_nat = ifelse(L_hc_nat < 0, 0, L_hc_nat)) %>% 
#   left_join(sp_index, by = c("species", "sc_name")) %>% 
#   select(year, sc_name, species, en_name, catch_nat, NHC_m_nat, L_hc_nat)
#   

```


## Step 2: Multiply landings for human consumptions 

```{r}
##Import yield proportion data
DHC_yield <- read_xlsx(here("clean_databases/intermediate/sp_cl_DHC.xlsx")) %>% 
  select(-category, -reference)

##Calculate yield at a national level
species_yield_nat <- landing_hc_nat %>% 
  left_join(DHC_yield, by= c("species", "sc_name")) %>% 
  select(year, sc_name, species, L_hc_nat, mean_yield) %>%
  mutate(L_hc_yield = L_hc_nat*mean_yield)
  
  
```


## Step 3: Add aquaculture to L_hc and subtract Exports


```{r}
##Import Aquaculture data

aquaculture_nat <- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv")) %>% 
  filter(region != "national",
         year!=2012) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  select(-subsector, -month) %>% 
  rename(aquaculture = tons) %>% 
  left_join(sp_index, by= "species") %>% 
  group_by(year, species, sc_name, en_name) %>% 
  summarise(aquaculture_nat = sum(aquaculture, na.rm = T))
  


##Import Export data
exports_nat <- read_csv(here("clean_databases/exports_ADUANA.csv")) %>% 
  select(year, region, exports= tons, species= especie_SERNAPESCA, sc_name=nombre_cientifico_SERNAPESCA, value) %>%
  filter(year!= 2018,
         region != 20) %>% 
  mutate(region = ifelse(region==0, NA, region)) %>%
  mutate(region = ifelse(region==1, "I", region)) %>% 
  mutate(region = ifelse(region==2, "II", region)) %>% 
  mutate(region = ifelse(region==3, "III", region)) %>% 
  mutate(region = ifelse(region==4, "IV", region)) %>% 
  mutate(region = ifelse(region==5, "V", region)) %>% 
  mutate(region = ifelse(region==6, "VI", region)) %>% 
  mutate(region = ifelse(region==7, "VII", region)) %>% 
  mutate(region = ifelse(region==8, "VIII", region)) %>% 
  mutate(region = ifelse(region==9, "IX", region)) %>% 
  mutate(region = ifelse(region==10, "X", region)) %>% 
  mutate(region = ifelse(region==11, "XI", region)) %>% 
  mutate(region = ifelse(region==12, "XII", region)) %>% 
  mutate(region = ifelse(region==14, "XIV", region)) %>% 
  mutate(region = ifelse(region==15, "XV", region)) %>% 
  mutate(region = ifelse(region==13, "R.M", region)) %>%
  mutate(species = str_to_upper(species)) %>% 
  mutate(sc_name = ifelse(species == "SALMON OTROS", "Salmonidae", sc_name)) %>%
  mutate(sc_name = ifelse(species == "ALGAS MIXTAS", "Algas", sc_name)) %>% 
  select(year, region, species, sc_name, exports) %>% 
  group_by(year, species, sc_name) %>% 
  summarise(export_nat = round(sum(exports, na.rm = T),2)) %>% 
  left_join(sp_index, by= c("species", "sc_name")) %>% 
  select(year, species, sc_name, en_name, export_nat) %>% 
  ungroup()
exports_nat[is.na(exports_nat)] <- 0

# %>%
#   data.frame() %>% 
#   mutate(year = as.integer(year)) %>% 
#   mutate(exports_nat = as.numeric(exports_nat))


##Add aquaculture and sutract exports
model_hc_nat <-landing_hc_nat %>%
  left_join(species_yield_nat, by= c("year", "species", "sc_name", "L_hc_nat")) %>% 
  left_join(aquaculture_nat, by= c("year", "species", "sc_name", "en_name")) %>% 
  left_join(exports_nat, by= c("year", "species", "sc_name", "en_name")) %>% 
  mutate(aquaculture_nat = ifelse(is.na(aquaculture_nat), 0, aquaculture_nat))

model_hc_nat$export_nat[is.na(model_hc_nat$export_nat)] <- 0

model_hc_nat <- model_hc_nat %>% 
  mutate(hc_total = (L_hc_nat+aquaculture_nat)-export_nat)
  


  


```



## Step 4: Add imports

Imports are not at a species level therfore we are going to group species by assinged categories according to import data. 

```{r}
##read in group index
group_index <- read.csv(here("clean_databases/sp_index.csv")) %>% 
  rename(species = chl_name) 

##Add groups to model dataframe
model_hc_group <- model_hc_nat %>% 
  left_join(group_index, by=c("sc_name", "species", "en_name"))


##import import data
imports <- read_csv(here("clean_databases/imports_ADUANA.csv")) %>% 
  select(year, tons, sp_group=name_sp, tons)


  


```



### Simple model

```{r}
sp_index <- read_csv(here("clean_databases/sp_index.csv"))

chl_to_sc <- sp_index %>% 
  select(chl_name, sc_name) %>% 
  unique()

export_gp_index <- read.csv(here("clean_databases/intermediate/export_groups_index.csv"))

##Landings from anuarios
landings_an <- read.csv(here("clean_databases/landings_anuario.csv")) %>% ##all ready match with sp_index
  select(year, sc_name, catch = tons)

##landings normales
landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(!region %in% c("national", "R.M.")) %>% 
  select(-month, -international_waters, -subsector) %>% 
  rename(chl_name = species) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  left_join(chl_to_sc, by= "chl_name") %>% 
  select(year, sc_name, tons) %>% 
  group_by(year, sc_name) %>% 
  summarise(catch = sum(tons, na.rm = T)) %>% 
  left_join(export_gp_index, by= "sc_name") %>% 
  mutate(export_group = as.character(grupo_exportacion)) %>% 
  mutate(group_name = ifelse(is.na(export_group), sc_name, export_group)) %>%
  select(year, sc_name, group_name, catch) %>% 
  ungroup()
  
  
  

##Aquaculture
aquaculture <- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv")) %>% 
  filter(region != "national",
         year!=2012) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  rename(chl_name = species) %>% 
  select(-subsector, -month) %>% 
  rename(aquaculture = tons) %>% 
   left_join(chl_to_sc, by= "chl_name") %>% 
  select(year, sc_name, aquaculture) %>% 
  group_by(year,sc_name) %>% 
  summarise(aquaculture= sum(aquaculture, na.rm = T)) %>% 
  left_join(export_gp_index, by= "sc_name") %>% 
  mutate(export_group = as.character(grupo_exportacion)) %>% 
  mutate(group_name = ifelse(is.na(export_group), sc_name, export_group)) %>%
  select(year, sc_name, group_name, aquaculture) %>% 
  ungroup()


##Exports
exports <- read_csv(here("clean_databases/exports_ADUANA.csv")) %>% 
  select(year, region, group_name=grupo_exportacion, exports= tons) %>%
  filter(year!= 2018) %>% 
  group_by(year, group_name) %>% 
  summarise(export = round(sum(exports, na.rm = T)))

non_human_cons <- read.csv(here("clean_databases/NHC_m.csv")) %>% 
  group_by(year, sc_name) %>% 
  mutate(total_NHC = as.numeric(total_NHC)) %>% 
  summarise(total_NHC = sum(total_NHC, na.rm = T)) %>% 
  left_join(export_gp_index, by= "sc_name") %>% 
  mutate(export_group = as.character(grupo_exportacion)) %>% 
  mutate(group_name = ifelse(is.na(export_group), sc_name, export_group)) %>%
  select(year, sc_name, group_name, total_NHC) %>% 
  ungroup()

intermediate_df <- landings %>% 
  full_join(aquaculture, by= c("year","sc_name", "group_name")) %>% 
  full_join(non_human_cons, by= c("year","sc_name", "group_name")) %>%
  group_by(year, group_name) %>% 
  summarise(catch = sum(catch, na.rm = T),
            aquaculture = sum(aquaculture, na.rm = T),
            NHC = sum(total_NHC, na.rm = T)) %>% 
  full_join(exports, by= c("year", "group_name"))
  
  intermediate_df$export[is.na(intermediate_df$export)] <- 0

preliminary_model <- intermediate_df %>% 
  mutate(prelim = catch + aquaculture - NHC - export)


DHC_yield <- read_xlsx(here("clean_databases/intermediate/sp_cl_DHC.xlsx")) %>% 
  select(sc_name, yield = mean_yield) %>%
  unique() %>% 
  left_join(export_gp_index, by= "sc_name") %>% 
  mutate(export_group = as.character(grupo_exportacion)) %>% 
  mutate(group_name = ifelse(is.na(export_group), sc_name, export_group)) %>%
  select(sc_name, group_name, yield) %>% 
  ungroup()

dhc_yield_mean <- DHC_yield %>% 
  group_by(group_name) %>% 
  summarize(mean_yield = mean(yield, na.rm = T)) %>% 
  ungroup() %>% 
  mutate(group_name = ifelse(group_name== "ATUN ESCOFINA/KONSO", "Ruvettus pretiosus", group_name))

intermediate_yield <- intermediate_df %>% 
  left_join(dhc_yield_mean, by= "group_name") %>% 
  mutate(prelim = ((catch + aquaculture - NHC)*mean_yield)-export)

# domestic_consumption <- intermediate_df %>% 
#   mutate(HC = catch + aquaculture - export) %>% 
#   left_join(sp_index, by= "sc_name")


##Add import data
# imports <- read_csv(here("clean_databases/imports_ADUANA.csv")) %>% 
#   select(year, name_sp, imports=tons) 
# 
# group_match_imports <- read_excel(here("clean_databases/intermediate/species_import_group.xlsx"))
# 
# ##match imports with group name
# imports_group <- imports %>% 
#   left_join(group_match_imports, by= "name_sp") %>% 
#   select(year, group, imports) %>% 
#   group_by(year, group) %>% 
#   summarise(imports = sum(imports, na.rm = T))
# 
# 
# sp_groups <- read.csv(here("clean_databases/sp_index.csv")) %>% 
#   select(sc_name, group, en_group)
# 
# int_group <- intermediate_df %>% 
#   left_join(sp_groups, by= "sc_name") %>% 
#   unique() %>% 
#   group_by(year, group) %>% 
#   summarise(catch = sum(catch),
#              no_hc = sum(total_NHC, na.rm = T),
#              aquaculture = sum(aquaculture, na.rm = T),
#             export= sum(export, na.rm = T),
#             mean_yield = mean(mean_yield, na.rm = T))
# 
# int_total <- int_group %>% 
#   left_join(imports_group, by= c("year", "group")) 
# 
# int_total$imports[is.na(int_total$imports)] <- 0
# 
# 
# 
# domestic_consumption <- int_total %>% 
#   mutate(HC = ((catch- no_hc)*mean_yield) + aquaculture + imports - export)
# 
# 
# domestic_consumption_no_yield <- int_total %>% 
#   mutate(HC = (catch- no_hc) + aquaculture + imports - export)



```


##Top domestic consumption
```{r}
# top_10_hc <- domestic_consumption %>% 
#   arrange(year, desc(HC)) %>% 
#   group_by(year) %>% 
#   top_n(10, HC)
# 
# top_5_hc <- domestic_consumption %>% 
#   arrange(year, desc(HC)) %>% 
#   group_by(year) %>% 
#   top_n(5, HC)

top_10_hc <- intermediate_yield%>% 
  arrange(year, desc(prelim)) %>% 
  group_by(year) %>% 
  top_n(10, prelim)

top_10_hc_ny <- preliminary_model%>% 
  arrange(year, desc(prelim)) %>% 
  group_by(year) %>% 
  top_n(10, prelim)


top_5_hc <- domestic_consumption %>% 
  arrange(year, desc(HC)) %>% 
  group_by(year) %>% 
  top_n(5, HC)




top_5_hc_ny <- domestic_consumption_no_yield %>% 
  arrange(year, desc(HC)) %>% 
  group_by(year) %>% 
  top_n(5, HC)




```









