---
title: "modelo_final"
author: "Camila Vargas e Ignacia Rivera"
date: "14 de diciembre de 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Joining landings, aquaculture, imports and exports by trade name

Create variables that allows us to match each of the databese

```{r matiching function, echo = FALSE, message=FALSE, warning = FALSE}

##complete specie name list with corresponding trade name
sp_index <- read.csv(here('clean_databases/sp_index.csv'))

# Chilean to trade name
chl_to_trade <- sp_index %>% 
  select(chl_name, trade_name ) %>% 
  unique()

# Imports to trade name 
imp_to_trade <- read.csv(here('clean_databases/imports_index.csv')) %>% 
  rename(imports_name = X...imports_name)

# Exports to trade name 
##exp_to_trade <- read.csv(here('clean_databases/exports_index.csv')) %>% 
  ##rename(exports_name = X...export_name)

exp_to_trade <- read_delim("clean_databases/exports_index.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  rename(exports_name = export_name)

# english to trade
en_to_trade <- sp_index %>% 
  select(en_name, trade_name ) %>% 
  unique()

# scientific to trade
sc_to_trade <- sp_index %>% 
  select(sc_name, trade_name ) %>% 
  unique()

```

Convert each data base into the trade name using the dataframe created above


```{r convertion each datasat to trade name}
# Landings: Importing, cleaning and grouping by trade name 
L <- read.csv(here('clean_databases/landings_SERNAPESCA.csv')) %>% 
  filter(region != 'national') %>% 
  select(year, chl_name = species, tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(landing = sum(tons, na.rm= TRUE))%>% 
  ungroup()
  
# Aquaculture: Importing, cleaning and grouping by trade name
A <- read.csv(here('clean_databases/aquaculture_SERNAPESCA.csv')) %>% 
  filter(region != 'national') %>%
  filter(year !=2012) %>% 
  select(year, chl_name = species, tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(aquaculture = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Imports: Importing, cleaning and grouping by trade name
I <- read.csv(here('clean_databases/imports_ADUANA.csv')) %>% 
  select(year, imports_name = name_sp, tons) %>% 
  filter(year != 2018) %>% 
  left_join(imp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(import = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Exports: Importing, cleaning and grouping by trade name
E <- read.csv(here('clean_databases/exports_ADUANA.csv')) %>% 
  select(year, exports_name = grupo_exportacion, tons) %>% 
  filter(year != 2018) %>% 
  left_join(exp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(export = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Yields: Importing, cleaning and grouping by trade name
Y <- sp_index %>% 
  group_by(trade_name) %>% 
  summarise(mean_yield= mean(yield))%>% 
  ungroup()

# Non-himan consumption: Importing, cleaning and grouping by trade name
NHC <- read.csv(here("clean_databases/NHC_m.csv")) %>% 
  group_by(year, sc_name) %>% 
  summarise(NHC = sum(total_NHC, na.rm = T)) %>% 
  ungroup() %>% 
  left_join(sc_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(NHC = sum(NHC, na.rm= TRUE)) %>% 
  ungroup()
  

```

Combining all datasets by trade name to creat the model

Implementing model Dom_consumption = (L+A)*Y + I - E

```{r combining and calculatind}
  
## full-join
model_trade_name <- L %>% 
  full_join(A, by= c("year", "trade_name")) %>% 
  full_join(NHC, by= c("year", "trade_name")) %>% 
  full_join(E, by= c("year", "trade_name")) %>% 
  full_join(I, by= c("year", "trade_name")) %>% 
  full_join(Y, by= "trade_name") %>% 
  mutate(landing = as.numeric(landing))

#convert all NAs to 0
model_trade_name$landing[is.na(model_trade_name$landing)] <- 0

model_trade_name$aquaculture[is.na(model_trade_name$aquaculture)] <- 0

model_trade_name$NHC[is.na(model_trade_name$NHC)] <- 0

model_trade_name$export[is.na(model_trade_name$export)] <- 0

model_trade_name$import[is.na(model_trade_name$import)] <- 0

model_trade_name$mean_yield[is.na(model_trade_name$mean_yield)] <- 0




###LA HORA DE ELA VERDAD

dom_cons_trade_name <- model_trade_name %>% 
  mutate(HC = ((landing + aquaculture - NHC)*mean_yield) - export + import) %>% 
  mutate(per_cap_kg = format((HC/17762681)*1000000, scientific = FALSE))

per_capita_total <- dom_cons_trade_name %>% 
  group_by(year) %>% 
  summarise(HC_year = sum(HC))

## Save final results
write.csv(dom_cons_trade_name, here("clean_databases/final_model_results.cvs"), row.names = F)


```


top 10 fisheries
```{r}

top_10_hc <- dom_cons_trade_name %>%
  arrange(year, desc(HC)) %>%
  group_by(year) %>%
  top_n(10, HC)

View(top_10_hc)


top_5_hc <- dom_cons_trade_name %>%
  arrange(year, desc(HC)) %>%
  group_by(year) %>%
  top_n(5, HC)

View(top_5_hc)


top_5_per_capita <- dom_cons_trade_name %>%
  arrange(year, desc(per_cap_kg)) %>%
  group_by(year) %>%
  top_n(5, HC)

View(top_5_per_capita)


```



####TODO CON LANDINGS ANUARIOS

```{r matiching function, echo = FALSE, message=FALSE, warning = FALSE}

##complete specie name list with corresponding trade name
sp_index <- read.csv(here('clean_databases/sp_index.csv'))

# Chilean to trade name
chl_to_trade <- sp_index %>% 
  select(chl_name, trade_name ) %>% 
  unique()

# Imports to trade name 
imp_to_trade <- read.csv(here('clean_databases/imports_index.csv')) %>% 
  rename(imports_name = X...imports_name)

# Exports to trade name 
##exp_to_trade <- read.csv(here('clean_databases/exports_index.csv')) %>% 
  ##rename(exports_name = X...export_name)

exp_to_trade <- read_delim("clean_databases/exports_index.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  rename(exports_name = export_name)

# english to trade
en_to_trade <- sp_index %>% 
  select(en_name, trade_name ) %>% 
  unique()

# scientific to trade
sc_to_trade <- sp_index %>% 
  select(sc_name, trade_name ) %>% 
  unique()

```

Convert each data base into the trade name using the dataframe created above


```{r convertion each datasat to trade name}
# Landings: Importing, cleaning and grouping by trade name 
L <- read.csv(here('clean_databases/landings_SERNAPESCA.csv')) %>% 
  filter(region != 'national') %>% 
  select(year, chl_name = species, tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(landing = sum(tons, na.rm= TRUE))%>% 
  ungroup()
  
# Aquaculture: Importing, cleaning and grouping by trade name
A <- read.csv(here('clean_databases/aquaculture_SERNAPESCA.csv')) %>% 
  filter(region != 'national') %>%
  filter(year !=2012) %>% 
  select(year, chl_name = species, tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(aquaculture = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Imports: Importing, cleaning and grouping by trade name
I <- read.csv(here('clean_databases/imports_ADUANA.csv')) %>% 
  select(year, imports_name = name_sp, tons) %>% 
  filter(year != 2018) %>% 
  left_join(imp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(import = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Exports: Importing, cleaning and grouping by trade name
E <- read.csv(here('clean_databases/exports_ADUANA.csv')) %>% 
  select(year, exports_name = grupo_exportacion, tons) %>% 
  filter(year != 2018) %>% 
  left_join(exp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(export = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Yields: Importing, cleaning and grouping by trade name
Y <- sp_index %>% 
  group_by(trade_name) %>% 
  summarise(mean_yield= mean(yield))%>% 
  ungroup()

# Non-himan consumption: Importing, cleaning and grouping by trade name
NHC <- read.csv(here("clean_databases/NHC_m.csv")) %>% 
  group_by(year, sc_name) %>% 
  summarise(NHC = sum(total_NHC, na.rm = T)) %>% 
  ungroup() %>% 
  left_join(sc_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(NHC = sum(NHC, na.rm= TRUE)) %>% 
  ungroup()
  

```

Combining all datasets by trade name to creat the model

Implementing model Dom_consumption = (L+A)*Y + I - E

```{r combining and calculatind}
  
## full-join
model_trade_name <- L %>% 
  full_join(A, by= c("year", "trade_name")) %>% 
  full_join(NHC, by= c("year", "trade_name")) %>% 
  full_join(E, by= c("year", "trade_name")) %>% 
  full_join(I, by= c("year", "trade_name")) %>% 
  full_join(Y, by= "trade_name") %>% 
  mutate(landing = as.numeric(landing))

#convert all NAs to 0
model_trade_name$landing[is.na(model_trade_name$landing)] <- 0

model_trade_name$aquaculture[is.na(model_trade_name$aquaculture)] <- 0

model_trade_name$NHC[is.na(model_trade_name$NHC)] <- 0

model_trade_name$export[is.na(model_trade_name$export)] <- 0

model_trade_name$import[is.na(model_trade_name$import)] <- 0

model_trade_name$mean_yield[is.na(model_trade_name$mean_yield)] <- 0




###LA HORA DE ELA VERDAD

dom_cons_trade_name <- model_trade_name %>% 
  mutate(HC = ((landing + aquaculture - NHC)*mean_yield) - export + import) %>% 
  mutate(per_cap_kg = format((HC/17762681)*1000000, scientific = FALSE))

per_capita_total <- dom_cons_trade_name %>% 
  group_by(year) %>% 
  summarise(HC_year = sum(HC))

## Save final results
write.csv(dom_cons_trade_name, here("clean_databases/final_model_results.cvs"), row.names = F)


```


top 10 fisheries
```{r}

top_10_hc <- dom_cons_trade_name %>%
  arrange(year, desc(HC)) %>%
  group_by(year) %>%
  top_n(10, HC)

View(top_10_hc)


top_5_hc <- dom_cons_trade_name %>%
  arrange(year, desc(HC)) %>%
  group_by(year) %>%
  top_n(5, HC)

View(top_5_hc)


top_5_per_capita <- dom_cons_trade_name %>%
  arrange(year, desc(per_cap_kg)) %>%
  group_by(year) %>%
  top_n(5, HC)

View(top_5_per_capita)



```


###TODOD CON LANDINGS ANUARIOS

```{r matiching function, echo = FALSE, message=FALSE, warning = FALSE}

##complete specie name list with corresponding trade name
sp_index <- read.csv(here('clean_databases/sp_index.csv'))

# Chilean to trade name
chl_to_trade <- sp_index %>% 
  select(chl_name, trade_name ) %>% 
  unique()

# Imports to trade name 
imp_to_trade <- read.csv(here('clean_databases/imports_index.csv')) %>% 
  rename(imports_name = X...imports_name)

# Exports to trade name 
exp_to_trade <- read.csv(here('clean_databases/exports_index.csv'))

# english to trade
en_to_trade <- sp_index %>% 
  select(en_name, trade_name ) %>% 
  unique()

# scientific to trade
sc_to_trade <- sp_index %>% 
  select(sc_name, trade_name ) %>% 
  unique()

```

Convert each data base into the trade name using the dataframe created above


```{r convertion each datasat to trade name}
# Landings: Importing, cleaning and grouping by trade name 
L_an <- read.csv(here("clean_databases/landings_anuario.csv")) %>% 
  select(year, chl_name = species, sc_name, landing_aqua = tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(landing_aqua = sum(landing_aqua, na.rm= TRUE))%>% 
  ungroup()
  
# Aquaculture: Importing, cleaning and grouping by trade name
# A <- read.csv(here('clean_databases/aquaculture_SERNAPESCA.csv')) %>% 
#   filter(region != 'national') %>%
#   filter(year !=2012) %>% 
#   select(year, chl_name = species, tons) %>% 
#   left_join(chl_to_trade) %>% 
#   group_by(year, trade_name) %>% 
#   summarise(aquaculture = sum(tons, na.rm= TRUE))%>% 
#   ungroup()

# Imports: Importing, cleaning and grouping by trade name
I <- read.csv(here('clean_databases/imports_ADUANA.csv')) %>% 
  select(year, imports_name = name_sp, tons) %>% 
  filter(year != 2018) %>% 
  left_join(imp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(import = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Exports: Importing, cleaning and grouping by trade name
E <- read.csv(here('clean_databases/exports_ADUANA.csv')) %>% 
  select(year, export_name = grupo_exportacion, tons) %>% 
  filter(year != 2018) %>% 
  left_join(exp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(export = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Yields: Importing, cleaning and grouping by trade name
Y <- sp_index %>% 
  group_by(trade_name) %>% 
  summarise(mean_yield= mean(yield))%>% 
  ungroup()

# Non-himan consumption: Importing, cleaning and grouping by trade name
NHC_raw <- read.csv(here("clean_databases/processing.csv")) %>%
  filter(m_p == "M",
         use== "NHC") %>% 
  group_by(year, chl_name) %>% 
  summarise(NHC = sum(tons, na.rm = T)) %>% 
  ungroup() %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(NHC = sum(NHC, na.rm= TRUE)) %>% 
  ungroup()
  
##Alterantive NHC
# raw_NHC <- read.csv(here("clean_databases/raw_NHC.csv")) %>% 
#   rename(chl_name = species) %>% 
#   group_by(chl_name, year) %>% 
#   summarise(NHC = sum(tons, na.rm = T)) %>% 
#   ungroup() %>% 
#   left_join(chl_to_trade) %>% 
#   group_by(year, trade_name) %>% 
#   summarise(NHC = sum(NHC, na.rm= TRUE)) %>% 
#   ungroup()
  
```

Combining all datasets by trade name to creat the model

Implementing model Dom_consumption = (L+A)*Y + I - E

```{r combining and calculatind}
  
## full-join
model_trade_name_Lan <- L_an %>% 
  ##full_join(A, by= c("year", "trade_name")) %>% 
  full_join(NHC_raw, by= c("year", "trade_name")) %>% 
  full_join(E, by= c("year", "trade_name")) %>% 
  full_join(I, by= c("year", "trade_name")) %>% 
  full_join(Y, by= "trade_name")

#convert all NAs to 0
model_trade_name_Lan$landing_aqua[is.na(model_trade_name_Lan$landing_aqua)] <- 0

# model_trade_name_Lan$aquaculture[is.na(model_trade_name_Lan$aquaculture)] <- 0

model_trade_name_Lan$NHC[is.na(model_trade_name_Lan$NHC)] <- 0

model_trade_name_Lan$export[is.na(model_trade_name_Lan$export)] <- 0

model_trade_name_Lan$import[is.na(model_trade_name_Lan$import)] <- 0

model_trade_name_Lan$mean_yield[is.na(model_trade_name_Lan$mean_yield)] <- 0

##intermediate check landing_aqua -NHC
landing_aqua_no_NHC <- model_trade_name_Lan %>% 
  mutate(land_no_NHC = landing_aqua - NHC) %>% 
  arrange(year, desc(land_no_NHC)) %>%
  group_by(year) %>%
  top_n(10, land_no_NHC)

View(landing_aqua_no_NHC)

###L
dom_cons_Lan <- model_trade_name_Lan %>% 
  mutate(HC = ((landing_aqua - NHC)*mean_yield) - export + import) %>%
  mutate(per_cap_kg = format((HC/17762681)*1000000, scientific = FALSE))

per_capita_total_Lan <- dom_cons_Lan %>%
  filter(HC>0) %>% 
  group_by(year) %>% 
  summarise(HC_year = sum(HC)) %>% 
  mutate(per_capita = (HC_year/17762681)*1000)





## Save final results
#write.csv(dom_cons_Lan, here("clean_databases/final_model_results_Lan.cvs"), row.names = F)


```


top 10 fisheries landings anuario
```{r}

top_5_hc_Lan <- dom_cons_Lan %>%
  arrange(year, desc(HC)) %>%
  group_by(year) %>%
  top_n(5, HC)

View(top_5_hc_Lan)


top_5_Land_Aqua <- L_an %>%
  arrange(year, desc(landing_aqua)) %>%
  group_by(year) %>%
  top_n(5, landing_aqua)

View(top_5_Land_Aqua)



top_10_Land_Aqua <- L_an %>%
  arrange(year, desc(landing_aqua)) %>%
  group_by(year) %>%
  top_n(10, landing_aqua)

View(top_10_Land_Aqua)

top_10_NHC <- raw_NHC %>%
  arrange(year, desc(NHC)) %>%
  group_by(year) %>%
  top_n(10, NHC)

View(top_10_NHC)


top_10_I <- I %>%
  arrange(year, desc(import)) %>%
  group_by(year) %>%
  top_n(10, import)

View(top_10_I)

top_10_E <- E %>%
  arrange(year, desc(export)) %>%
  group_by(year) %>%
  top_n(10, export)

View(top_10_E)



```



```{r}


top5_hc_fig2 <- 
  ggplot(data = top_5_hc_Lan, aes(x = year, y = HC, fill = trade_name)) +
    geom_bar(stat= "identity")+
    scale_y_continuous(expand = c(0, 0)) +
    labs(y="Domestic human consumption \n (Tonnes)", x="Year")+
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=11),
        axis.title=element_text(size=12,face="bold"))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1", 5, type = "discrete"),wes_palette("Darjeeling2", 5, type = "discrete")))+
  ggtheme_plot()


```


