---
title: 'Scope 3: Domestic fish consumption'
author: "Camila Vargas e Ignacia Rivera"
date: "August 17, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setup
```{r}
library(rfishbase) #install.packages("rfishbase")
library(tidyverse)
library(here)
```



In order to calculate the domestic food compsumtion we buit a model that considers landings, aquaculture harvest, exports and imports. 

In order to imput the data to the model we first need to combine all our clean datasets into one dataframe that looks like the dummy dataframe (Juliano). This dataframe has the following columns:
Species	
Country	
Region	
Year	
Variable	
Catch	
Aquaculture	
Imports	
Exports	
DHC_Prop

According to the data provided by ADUANA CHILE, exports are in a much detailed levels than imports. For this reseon we decided to create two main dataframe:
1. With landings, aquaculture harvest and exports detailed by species and region
2. With landings, aquaculture harvest, exports and imports by groups of species (specified in celan_databases/species_match.csv) at a national level.

# Creating the first dataframe: Including landings, aquaculture harves and exports at a species and regional level

## Importing datasets

Landings and aquacaltue are then combine with species_match in order to have scientific names


```{r}

##Species list with scientific name 
sp_index <- read_excel(here("clean_databases/sp_index.xlsx")) %>% 
  rename(species = chl_name)

landings_raw <- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(!region %in% c("national", "R.M.")) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  select(-month, -international_waters, -subsector) 

##landings
landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(!region %in% c("national", "R.M.")) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  select(-month, -international_waters, -subsector) %>% 
  group_by(year, region, species) %>% 
  summarise(tons = sum(tons, na.rm = T)) %>% 
  mutate(variable = "tonnes") %>% 
  rename(catch = tons) %>% 
  left_join(sp_index, by= "species")
  
##aquaculture harvest
aquaculture <- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv")) %>% 
  filter(region != "national",
         year!=2012) %>% 
  mutate( tons = as.numeric(tons)) %>% 
  select(-subsector, -month) %>% 
  mutate(variable = "tonnes") %>% 
  rename(aquaculture = tons) %>% 
  left_join(sp_index, by= "species")


## Exports

exports_all <- read_csv(here("clean_databases/exports_ADUANA.csv")) %>% 
  select(year, region, exports= tons, species= especie_SERNAPESCA, sc_name=nombre_cientifico_SERNAPESCA, value) %>%
  filter(year!= 2018,
         region != 20) %>% 
  mutate(region = ifelse(region==0, NA, region)) %>%
  mutate(region = ifelse(region==1, "I", region)) %>% 
  mutate(region = ifelse(region==2, "II", region)) %>% 
  mutate(region = ifelse(region==3, "III", region)) %>% 
  mutate(region = ifelse(region==4, "IV", region)) %>% 
  mutate(region = ifelse(region==5, "V", region)) %>% 
  mutate(region = ifelse(region==6, "VI", region)) %>% 
  mutate(region = ifelse(region==7, "VII", region)) %>% 
  mutate(region = ifelse(region==8, "VIII", region)) %>% 
  mutate(region = ifelse(region==9, "IX", region)) %>% 
  mutate(region = ifelse(region==10, "X", region)) %>% 
  mutate(region = ifelse(region==11, "XI", region)) %>% 
  mutate(region = ifelse(region==12, "XII", region)) %>% 
  mutate(region = ifelse(region==14, "XIV", region)) %>% 
  mutate(region = ifelse(region==15, "XV", region)) %>% 
  mutate(region = ifelse(region==13, "R.M", region)) %>%
  mutate(species = str_to_upper(species)) %>% 
  mutate(sc_name = ifelse(species == "SALMON OTROS", "Salmonidae", sc_name)) %>%
  mutate(sc_name = ifelse(species == "ALGAS MIXTAS", "Algas", sc_name))

exports_tons <- exports_all %>% 
  select(-value) %>% 
  mutate(variable = "tonnes")


exports_value <- exports_all %>% 
  select(-exports) %>%
  rename(exports= value) %>% 
  mutate(variable = "value")


exports <- exports_tons %>% 
  rbind(exports_value)


```




#### Correction for not double accounting algaes
To algaes export we need to dicount all algae products that go out of the facilieties.

```{r}

algaes_product <- read.csv(here("clean_databases/product_algae_NHC.csv"))

exports_algae_discount <- exports_tons %>% 
  left_join(algaes_product, by= c("sc_name", "species", "year", "region")) %>% 
  mutate(tons= ifelse(is.na(tons), 0, tons)) %>% 
  mutate(export_total = exports-tons)
  



```



## Combining all datasets!

```{r}
database1_complete <- landings %>% 
  full_join(aquaculture, by= c("year", "region", "species", "variable", "sc_name", "en_name")) %>%
  full_join(exports, by= c("year", "region", "species", "variable", "sc_name"))

##Save as intermediate file in ordet to compare with yield_DHC
sp_model <- unique(database1_complete[c("species", "sc_name")])

##write.csv(sp_model, here("clean_databases/intermediate/sp_model.csv"), row.names=F)
```


## Calculation a t a national level
```{r}
model_national <- database1_complete %>% 
  group_by(year, sc_name, variable) %>% 
  summarise(catch_nat= sum(catch, na.rm = T),
            aquaculture_nat = sum(aquaculture, na.rm = T),
            exports_nat = round(sum(exports, na.rm = T), 2)) %>% 
  filter(variable == "tonnes") %>% 
  left_join(sp_index, by = "sc_name")
  
```





## Add DHC proportion

```{r}

##read dataset generated in DHC.Rmd
DHC_yield <- read_csv("clean_databases/intermediate/sp_cl_DHC.csv") %>% 
  select(-category, -reference)
  

##Combine both datasets - THIS LOOKS LIKE THE DUMMY DATASET
NSS_model1 <- database1_complete %>% 
  left_join(DHC_yield, by= c("species", "sc_name")) %>% 
  select(species_es= species, species_en, sc_name, year, region, variable, catch, aquaculture, exports, DHC_pop = mean_yield)

##Save


```

## Subtracting the tonnes that are not for direct human consumption

According to the type of product produced from each specie we calculated the amount that does not go to direct human consumption. The first step of the model is to discount those amount form the total landings

```{r}

model1_tonnes <- NSS_model1 %>% 
  filter(variable== "tonnes") %>% 
  select(species_es, species_en, sc_name, year, region, landings=catch, DHC_pop)


##read non-human consumption dataset
NHC_facilities <- read.csv(here("clean_databases/NHC_m.csv")) %>% 
  rename(m_nhc = total_NHC)

##read human consumption raw
HC_facilities_m <- read.csv(here("clean_databases/HC_m.csv")) %>% 
  rename(m_hc = total_HC_m)

##read human consumption products
HC_facilities_p <- read.csv(here("clean_databases/HC_p.csv")) %>% 
  rename(p_hc = total_HC_p)

##join to overall model matrix 
HC_model_landings <- model1_tonnes %>% 
  left_join(NHC_facilities, by = c("sc_name", "year", "region")) %>% 
  left_join(HC_facilities_m, by = c("sc_name", "year", "region")) %>% 
  left_join(HC_facilities_p, by = c("sc_name", "year", "region")) %>%
  mutate(landings = ifelse(is.na(landings), 0, landings),
         m_nhc = ifelse(is.na(m_nhc), 0, m_nhc),
         m_hc = ifelse(is.na(m_hc), 0, m_hc),
         p_hc = ifelse(is.na(p_hc), 0, p_hc)) %>% 
  mutate(L_HC = landings - m_nhc) %>% 
  mutate(L_playa = L_HC - m_hc) %>% 
  mutate(hc_facilities = m_hc - p_hc) %>% 
  mutate(hc_playa = L_playa*DHC_pop) %>% 
  mutate(total_HC = hc_playa + hc_facilities)

##Model sin algas
  
##read NHC sin algas
# NHC_facilities_sin_algas <- read.csv(here("clean_databases/NHC_sin_algas.csv"))%>% 
#   rename(m_nhc = total_NHC)
# 
# HC_model_sin_algas <- model1_tonnes %>% 
#   left_join(NHC_facilities_sin_algas, by = c("sc_name", "year", "region")) %>% 
#   left_join(HC_facilities_m, by = c("sc_name", "year", "region")) %>% 
#   left_join(HC_facilities_p, by = c("sc_name", "year", "region")) %>% 
#   filter(!species %in% c("CHASCA", "CHASCON O HUIRO NEGRO", "CHICOREA DE MAR", "COCHAYUYO", "HUIRO", "HUIRO PALO", "LECHUGUILLA", "LIQUEN GOMOSO", "LUCHE", "LUGA CUCHARA O CORTA", "LUCHA CUCHARA O CORTA", "LUGA NEGRA O CRESPA", "LUGA NEGRA", "LUGA-ROJA", "LUGA ROJA", "PELILLO", "ALGAS MIXTAS", "CAROLA", "HUIRO NEGRO O CHASCON", "HUIRO NEGRO", "HAEMATOCOCCUS", "SPIRULINA", "LUGA LUGA"))

```








Top fisheries for human consumption per region

```{r}

top_5_hc_reg <- HC_model_landings %>% 
  group_by(year, region) %>% 
  top_n(5, total_HC) %>% 
  data.frame()







```





# Creating the second dataframe: Including landings, aquaculture harvest, exports and imports at a species group lavel and at a national scale


## Wrangle importa data

And assing group name to match with database created above

```{r}
##Imports
imports <- read_csv(here("clean_databases/imports_ADUANA.csv")) %>% 
  select(year, tons, name_sp) %>% 
  rename(imports=tons) %>% 
   mutate(variable = "tonnes")

## Import database that matches name_sp with group name
group_match_imports <- read_excel(here("clean_databases/intermediate/species_import_group.xlsx"))

##match imports with group name
imports_group <- imports %>% 
  left_join(group_match_imports, by= "name_sp") %>% 
  select(year, group, variable, imports)

``` 


Database 1 has to be reduce to a nationl level (instead of regional) in order to combine with importa infomation

```{r}
database1_national <- database1_complete %>% 
  group_by(year, species) %>% 
  summarise(catch=sum(catch, na.rm = T),
            aquaculture = sum(aquaculture, na.rm = T),
            exports = sum(exports, na.rm = T))
```

match species with species group

```{r}
database2 <- database1_national %>% 
  left_join(species_match, by= "species")

###MINETRAS TANTO NO ARREGLAMOS LOS NOMBRES DESDE EL PRINCIPIO VOY A AGREGAR COMBRE DE GRUPO A TODAS LAS ESPECIES QUE NO TIENEN A MANO PERO IDELAMENTE HAY QUE COMPLETAR LA BASE DE DATOS EN EL SRCIPT species_group.Rmd
test <- database2 %>% 
  filter(is.na(group)) %>% 
  select(species) %>% 
  unique()
  
##Save to add group name by hand
##write.csv(test,here("clean_databases/intermediate/missing_species_group.csv"), row.names = F)


##Read file with group names
test_complete <- read.csv(here("clean_databases/intermediate/missing_species_group.csv")) %>% 
  select(species, group)

group_name_complete <- species_match %>% 
  select(species, group) %>% 
  rbind(test_complete)

database2_complete_names <- database1_national %>% 
  left_join(group_name_complete, by= "species") %>% 
  mutate(variable= "tonnes")
  
## add to database the products that are not human consumption
no_hc <- read.csv(here("clean_databases/landings_no_HC.csv")) %>% 
  group_by(year, species) %>% 
  summarise(no_hc= sum(tons)) %>%
  ungroup() %>% 
  mutate(species= as.character(species)) %>% 
  mutate(species = ifelse(stringr::str_detect(species,"SARDINA ESPA.*OLA"), "SARDINA ESPANOLA", species))

no_hc_groups <- no_hc %>% 
  left_join(group_name_complete, by= "species")
  

database2_complete <- database2_complete_names %>% 
  left_join(imports_group, by= c("group", "year", "variable")) %>% 
  left_join(no_hc_groups, by= c("group", "year", "species")) %>% 
  select(year, species, group, variable, catch, no_hc, aquaculture, exports, imports) %>% 
  mutate(imports= if_else(is.na(imports), 0, imports)) %>% 
  mutate(no_hc= as.numeric(no_hc)) %>% 
  mutate(no_hc= if_else(is.na(no_hc), 0, no_hc)) %>% 
  group_by(group, year) %>% 
  summarise(catch = sum(catch),
            no_hc = sum(no_hc),
            aquaculture = sum(aquaculture),
            exports= sum(exports),
            imports = sum(imports)) %>% 
  mutate(total = (catch+aquaculture+imports)- no_hc - exports) %>% 
  ungroup()

##Save complete database
write.csv(database2_complete, here("clean_databases/model_df_by_group.csv"), row.names = F)

top_10_hc <- database2_complete %>% 
  arrange(year, desc(total)) %>% 
  group_by(year) %>% 
  top_n(10, total)

top_5_hc <- database2_complete %>% 
  arrange(year, desc(total)) %>% 
  group_by(year) %>% 
  top_n(5, total)

```

## Preliminary figures

```{r}
main_hc <- 
  ggplot(data = top_5_hc, aes(x = year, y = total, fill = group)) + 
    geom_bar(stat= "identity")+
    scale_y_continuous(expand = c(0, 0)) +
    labs(y="Human Consumption (Tons)", x="Year")+
    theme_bw()+
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=11),
        axis.title=element_text(size=12,face="bold"))

plot(main_hc)

```



