---
output: 
  pdf_document: 
    fig_caption: yes
    includes:  
title: "Contributions of marine capture fisheries to the domestic livelihoods and seafood consumption of Chile"
subtitle: "Final Draft"
author: "Camila Vargas e Ignacia Rivera"
bibliography: Oceana.bib
csl: cell.csl
---
\newpage

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(here)
library(magrittr)
library(tidyr)
library(reshape2)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(readxl)
library(cowplot)
library(kableExtra)
library(readr)
library(stringr)
library(purrr)
library(naniar)
library(wesanderson)

ggtheme_plot <- function() {
  theme(
    plot.title = element_text(size = rel(1), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    strip.background = element_blank(),
    # strip.text       = element_text(size = base_size, face = "italic"),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 10,
                               angle = 0,
                               face = "plain"),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.key = element_rect(colour = NA, fill = NA),
    legend.position  = "top",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 18, colour = "darkgrey")
  )
}
```

## Methods

### Data sources

1. The Chilean National Service of Fisheries (SERNAPESCA) is the main entity in charge of keeping track of landings and other fisheries and aquaculture related data in Chile. Most of their data is available on their website but we also obtained databases upon request. We accessed their databases on landings from wild-caught and farmed fisheries, first transaction prices for artisanal fisheries, and landings used in different types of sea products. These databases provided information at the species and regional level. We also downloaded their data on the geographical distribution of production facilities and the types of sea products they produce. We issued data requests on SERNAPESCA's portal to obtain more detailed information. Via this method, they provided us the national record of artisanal fishers (RPA), which lists current registered fishers and boats, and the national record of industrial vessels (RPI), which lists current registered industrial vessels. The RPA informs the region, gender and age for each artisanal fisher as well as the species-specific permits they hold. Together the RPA and RPI provide information of artisanal and industrial vessels like lenght and storing capacity. SERNAPESCA also provided us the number of employees of each gender per processing facilities. 

2. The Chilean Costumes (ADUANAS Chile) is the main entity in charge of keeping records of imported and exported goods in the country. We accessed their exports and imports data per product and region via their website. Each product in the original exports and imports data-sets is identified through an id number that matches the name of the product in a code known as the _Arancel Aduanero_. Most products provided detail of the species or groups of species involved.

3. The Fisheries development institute (IFOP) is a private entity that advices fisheries management based on permanent biological, environmental and economic monitoring, and scientific reasoning. Although most of the data collected by IFOP pertains to stock assessments and biological indicators, they also generate economically relevant statistics. We reviewed their most recent annual report on the economic monitoring of Chilean fisheries for 2015. It includes databases and estimates for employment. We also extracted data from their monthly reports on exports of fisheries products, which are based on ADUANAS's data.  

4. The Chilean Central Bank (Banco Central de Chile) is the entity in charge of estabilizing the Chilean currency. To fullfill its function it monitors the main economic indicators per economic sector. We accessed their data on sectorial GDP for each region per industry. 

5. DataChile (www.datachile.com) is an online plataform that integrates, visualizes and distributes public available data from Chile. They also inform their visualizations with data from the ADUANAS Chile and national surveys. Using their visualization tools we extracted information on fish products imports for different municipalities.

6. We reviewed several reports from other national agencies like the Undersecretary of Fisheries (SUBPESCA), NGOs and scientific publications to complement our analysis.

### Specific methodologies

#### Currency conversion

Any conversion from chilean pesos to dollars was done by first converting the amount in chilean pesos to its value in Chilean Units of account (UF, a monetary unit corrected by inflation) of the respective year. Then, transformed to the equivalent in 2018 Chilean pesos using the 2018 UF value and finally, converting the 2018 value in chilean pesos to its equivalent in 2018 US dollars, based on the conversion for Dec, 2018 (1 US dollars corresponds to 676 Chilean pesos). 

```{r parameters for conversion, echo = FALSE}

# Based on data from the Tax National Service (SII)
mean_uf_2013 <- 22980 
mean_uf_2014 <- 23960
mean_uf_2015 <- 25022
mean_uf_2016 <- 26022
mean_uf_2017 <- 26571
mean_uf_2018 <- 27140 

from_ChP_to_dollar_dec_2018 <- 676 # Based on conversion for Dec, 2018
```

#### Fisheries context

To evaluate the volume contirbution of specific fisheries over the last five years, we used records of landings and aquaculture harvests from SERNAPESCA  (2013-2017). Using monthly data  published by IFOP, we identified the most exported fish products in terms of volume. To identify the participation of different species in exported volumes we generated a new database linking types of products listed in the _Arancel Aduanero_ with the names of the fished and farmed species listed by SERNAPESCA. We searched for all the products ids that refer to the species scientific name, common name or genus name as listed by SERNAPESCA and aggregated them under the species or group name.

#### Estimating National Seafood Supply

We combined data from Chilean Customs and SERNAPESCA to estimate domestic consumption with our model. We accessed their databases between August and September of 2018.

The specific data-sets we used are:

- Catches: annual tonnes per marine species landed in each region (SERNAPESCA, 2013-2017).
- Species contribution to different types of products: annual tons per marine species used in each region to elaborate each type of product (i.e. fish meal, oil, canned products, freeze products, dried products, smoked products, agar, col-agar, carragenina, dry algae) (SERNAPESCA, 2012-2016).
- Aquaculture: annual tonnes per species harvest from aquaculture centers in each region (SERNAPESCA, 2012-2017).
- Exports: quantities of all the products exported from each Chilean region to other countries. (ADUANAS, 2013-2018).
- Imports: quantities of all marine products imported into the country. (ADUANAS, 2013-2018)

Each product in the original exports and imports data-sets is identified through an id number that matches the name of the product in a code known as the _Arancel Aduanero_. This document was only available to us in .pdf format. Thus, to identified the id of products containing domestically fished and harvest marine species we searched for the names of the fished and harvested species listed by SERNAPESCA in the document with the _Arancel Aduanero_. We searched for all the ids that refer to the species scientific name, common name or genus name as listed by SERNAPESCA and aggregated them under the species name. In many cases this index is detailed at a species level, however several products group species. In order to avoid double accounting of a specie, we grouped several species under one group. We did not include the codes of any of the products for non-human consumption to avoid double counting since they were substracted in a previous stage of the model. 

We also calculated all raw material that enters to a facility and is used for non-human consumption products. For this, we assumed that the following categories of products are not for direct human consumption: fish meal, col-agar, agar, carragenina, oil and alginato, dry algae. Hence, the following products are considered for human consumption: fresh, frozen, salty dry, salty wet, smoked, canned, dehydrate. While fish meal, oil, dehydrate, dried algae, and other algae derivatives are use for non human consumption. We did not assume that all aquaculture goes to human-consumption because there are algaes harvest for no- human consumption. Finally, we assume that all algae that goes in to a processing facility is not for human consuption.

It was not possible to do the analysis at a regional level because there are internal transportation of raw and processed material within political regions that is hard to account for. One example of this is that some products are processed in the Metropolitan Region where there are no landings (land lock region).

To identify the species contribute more to human consumption products, we analyzed the main steps of the supply chain. In Chile, landings can either be sold as non-processed products ($NP_{HC}$) or go to processing facilities where they will be transformed in products for human consumption ($P_{HC}$) or products for other uses ($P_{NHC}$). Landing come either from industrial, which includes fabric vessels ($L_i$), or artisanal vessels ($L_a$). We assume that all $L_i$ are send to processing facilities while $L_a$ can either go to processing facilities or be sold as non-processed products, which we assume, are used for human consumption ($NP_{HC}$). Additionally, we also assume all the volumes coming from aquaculture centers enter processing facilities.
In the processesing of both products for HC and for NHC there are discards ($D_{HC}, D_{NHC}$) that can be represented as the differences between what entered the processing facility as raw material and what left it as processed product ($D_{HC} = R_{HC} - P_{HC}$, $D_{NHC} = R_{NHC} - P_{NHC}$ ). 

In Chile, almost all $L_i$ and $Aq$ are send to processing facilities while only some of the $L_a$ is processed. We assume that the rest of the artisanals' catch that is not sent to processing facilities is sold as non-processed products for human consumption ($NP_{HC}$). To get an estimate of $NP_{HC}$ we substracted $Aq$ and $L_i$ to all the raw material entering processing facilities and use the residual as the volume of $L_a$ that is processed. We then, substracted this volume from $L_a$ to get $NP_{HC}$, which is the rest of the artisanals' catch that is not being processed. The assumptions behind this method are (i) that all the raw material in processing facilities that is not being supplied by industrial fishing or aquaculture is supplied by artisanals, and (ii) all the catch that artisanals do not sell to processing facilities is used for human consumption.

We used the R software to clean and format all the available information in long-form databases. All the scripts for replicating our work are available in "scripts/cleaning_data".

#### Estimating National Economic Participation

We used data from Banco Central de Chile to reconstruct the percent of the regional GDP that comes from fisheries and calculate the mean percent between 2013 and 2016. 

To estimate most important fisheries in terms of economic value, we used data on first-transaction prices per species per month and location (requested to SERNAPESCA) and computed the national average price per year. We used two different approaches, one considers the total revenue generated by different fisheries while the other considers how valuable they are based on their price. To estimate total revenue we multiplied price by the landed tons per year. We converted Chilean pesos to 2018 dollars. Data on price at first transaction at the level of species was only available for the artisanal sector since data on industrial transactions is not recorded officially.

We collect information from several sources to reconstruct employment estimates for different stages in the supply chain of marine products. IFOP, 2015 found that 1.57 indirect employments are generated by each direct employment in the manufacturing fishing sector. We apply this multiplier to our most recent estimatation of people employed in the manufacturing fishing center to project indirect employment. We then qualitatively related the geographic distribution of employees in different stages of the supply chain os the most important fisheries in a given region to find the main species in terms of employment contribution.  

We also obtained data from SERNAPESCA detailing the number of people working in each processing facility. However, some of the values were clearly wrong (because they exceeded the entire Chilean population) making it impossible to use this information to generate reliable estimates. Thus, we relied of information from (@ifop_monitoreo_2015) to study the manufacturing sector. 

## Results

### Fisheries context

#### Main wild-caught fisheries in terms of volume

According to landing records from SERNAPESCA over the last five years, the artisanal sector lands on average 1.6 ± 0.58 million annual tons of marine species while the industrial sector lands around 0.9 ± 0.16 million tons per year. Is worth noting that the artisanal sector includes landings of algae and benthic species that are not targeted by the industrial sector. Both sectors have among their most landed species the Peruvian anchoveta (*Engraulis ringens*) and the Araucanian herring (*Strangomera bentincki*) two species that are mostly used as inputs for fish meal and oil (**Figure 1**). Other important species in artisanal landings are the Giant grey kelp (*Lessonia nigrescens*), exported mostly as dried algae, and the Jumbo squid (*Dosidicus gigas*), mostly commercialized fresh or frozen for HC. The Chilean jack mackerel (*Trachurus murphyi*) is also an important fishery in terms of volume for the industrial sector, which is also mostly use in the production of fish meal and oil and to a lesser extent use for direct HC smoked or canned. Other important species are the mote sculpin and different algaes for the artisanal sector (*Table App.1*). The South pacific hake (*Merluccius gayi*) and the Chub Mackarel (*Scomber japonicus*) are also important for industrial landings (*Table App.2*).

Over the last years the catch composition of the artisanal sector has presented no major changes except from one species of algae (Giant brown kelp, *Lessonia trabeculata*) replacing other (Red algae gracilaria, *Gracilaria spp*) within the top 5 landed species as for 2013. In the case of the industrial sector, composition of the main three species seem to be stable over the last five years. Yet, the South pacific hake (*Merlucius gayi*) stoped being one of the specied most fished by the industrial sector after 2013. This can be explined by the reduction of the total allowable catch for this species given its risk of collapse in those years (@subpesca_cuota_2013). After 2013, the industrial sector introduced the Jumbo squid (*Dosidicus gigas*) among its top 5 landed species. As for 2016 the Chub mackerel (*Scomber japonicus*) has entered the top 5 for the indutrial sector replacing Patagonian grenadier (*Macruronus magellanicus*).

```{r Most landed species in Chile, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= 'Top 5 species landed in Chile by the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. Data source: Landing records from SERNAPESCA, 2018.' }

# Importing landings and species names 

landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv"))
sp_index <- read_csv(here("clean_databases/sp_index.csv")) %>% 
  mutate(chl_name = str_to_upper(chl_name)) 

# Calculating annual landings per sector 

annual_landings_sector <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)) %>% 
  group_by(subsector) %>% 
  summarise(annual_tons = mean(tons), sd = sd(tons) )

# Top 5 landed per sector

# to avoid repeated observations

 landings_top_5 <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  rename(chl_name = species) %>% 
  left_join(sp_index) %>% 
  group_by(year, sc_name, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)/1000000)%>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, subsector) %>% 
  top_n(5, tons) %>%
  left_join(sp_index) %>% 
  select(year, subsector, tons, species = en_name) %>% 
  ggplot(aes(x = year, y = tons, fill = species))+
  geom_bar(stat = 'identity') +
  facet_wrap(~subsector)+
  labs(y="Landings (Million tons)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black'))+
  ggtheme_plot()
 
 landings_top_5
 
 # Top 10 landed per sector (Appendix)

landings_top_10 <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  rename(chl_name = species) %>% 
  left_join(sp_index) %>% 
  group_by(year, sc_name, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)/1000000)%>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, subsector) %>% 
  top_n(10, tons)

# to avoid double obsv
sc_name_to_en <- sp_index %>% 
  select(sc_name, en_name) %>% 
  unique()

#  top 10 artisanal
top_10_art <- landings_top_10 %>% 
  filter(subsector == 'artisanal') %>%
  left_join(sc_name_to_en) %>% 
  group_by(year) %>%
  mutate(rank = rank(desc(tons))) %>% 
  select(-sc_name, -subsector, -tons) %>% 
  spread(key= year, value =en_name) 
  
top_10_ind <- landings_top_10 %>% 
  filter(subsector == 'industrial') %>%
  left_join(sc_name_to_en) %>% 
  group_by(year) %>%
  mutate(rank = rank(desc(tons))) %>% 
  select(-sc_name, -subsector, -tons) %>% 
  spread(key= year, value =en_name) 
```

#### Regional Diferences in volumes of wild-caught fisheries 

Landed tons for 2017 shows that the amount of landings varies widely along the coast of Chile for both artisanal and industrial sectors.  **Figure 2** show the mean total landing per sector and region. The color of the bars represents the most landed species in 2017. We observe that the VIII is by far the region with most landings from both sectors. However, the industrial sector also concentrate landings in the north of the country while artisanals are more spread. The most landed species in 2017 also varies by region but has changed little over the last five years (**Table App.3** and **Table App.4**). 

```{r Landings per region, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= 'Mean annual total landings per region for the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. The color represents the most landed species in 2017. Error bars represent one standard deviation. Note that scales are different between pannels. Data source: Landing records from SERNAPESCA, 2018.'}

# Main species in 2017 per region and sector to add color to the barplot

  main_species_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, region, subsector) %>% 
  top_n(1, tons) %>% 
  filter(year == 2017)
  
 # Landings per region 
  
  landings_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, subsector, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(subsector, region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000)%>% 
  left_join(main_species_per_region, by = c('region', 'subsector')) %>% 
  select(subsector, region, annual_tons, sd, chl_name = species) %>% 
    left_join(sp_index) %>% 
    select(-chl_name, -sc_name, species_2017= en_name)
  
# Reordering regions
 landings_per_region$region <- factor(landings_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

 landings_per_region <- landings_per_region %>% 
 ggplot(aes(x = region, y = annual_tons, fill=species_2017))+
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(~ subsector, scales = 'free')+
  labs(y="Landings (Thousand Tons)", x="Region")+
  coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette('Royal1')))+
  ggtheme_plot()
 
 
# Main species per region artisanal sector overlast 5 years (Appendix)

 main_species_reg_art <- landings %>% 
  filter(region != "national", international_waters == 0, subsector == 'artisanal') %>% 
  mutate(tons = as.numeric(tons)) %>%
  rename(chl_name = species) %>% 
  left_join(sp_index) %>%
  group_by(year, region, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  left_join(sc_name_to_en) %>% 
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
 
main_species_reg_art$region <- factor(main_species_reg_art$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))
 
# Main species per region industrial sector overlast 5 years (Appendix)

 main_species_reg_ind <- landings %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  filter(region != "national", international_waters == 0, subsector == 'industrial') %>%
  rename(chl_name = species) %>% 
  left_join(sp_index) %>%
  group_by(year, region, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  filter(tons !=0 ) %>% 
  left_join(sc_name_to_en) %>% 
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
  
main_species_reg_ind$region <- factor(main_species_reg_ind$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

   
 landings_per_region

```

#### Main farmed fisheries in terms of volume

The main species harvested from aquaculture centers in the last 5 years have been the Atlantic salmon (*Salmo salar*), Coho salmon (*Oncorhynchus kisutch*), mussels (*Mytilus chilensis*) and the Red algae gracilaria (*Gracilaria spp.*) in descending order (**Figure 3**). Overall, the composition of the most farmed species and its volumes have been stable over the last 5 years. However, the Raibow trout (*Oncorhynchus mykiss*) seems to be becoming less important over recent years.  

```{r Most important farmed species in terms of volume in Chile, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= 'Top 5 species farmed in Chile between 2013 and 2017. Data source: Harvest records from SERNAPESCA, 2018.' }

# Importing aquaculture database
aquaculture<- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv"))

# Top aquaculture
  aquaculture_top_5 <-aquaculture %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons), chl_name = species) %>%
  left_join(sp_index) %>% 
  group_by(year, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>% 
  group_by(year) %>% 
  top_n(5, tons) %>%
  left_join(sp_index) %>% 
  rename(species= en_name) %>% 
  ggplot(aes(x = as.factor(year), y = tons/1000, fill = species))+
  geom_bar(stat = 'identity') +
  labs(y="Harvest (Thousand Tons)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  ggtheme_plot()
  
   aquaculture_top_5
  
```

#### Regional Diferences in farmed fisheries 

Aquaculture production is highly concentrated in the South of the country, between the X and the XII regions (**Figure 4**). Nevertheless, from the I to the IV region the most farmed resources are algaes and scallops (*Argopecten purpuratus*) (**Table App.5**). The red abalone (*Haliotis rufescens*) became the most farmed in 2013 in the V region, replacing the Turbot (*Scophtalmus maximus*). Towards the south salmons and trouts are the most farmed species along with some algaes and mussels (*Mytilus chilensis*). The most farmed species in the austral south are the Atlantic salmon (*Salmo salar*) and the Chilean mussel (*Mytilus chilensis*).

```{r Harvest from aquaculture centers per region, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= 'Mean annual harvest per region from aquaculture centers between 2012 and 2017. Error bars represent one standard deviation. Data source: Harvests records from SERNAPESCA, 2018.' }

# Main species in 2017 per region in case we want to add color to the barplot

  main_species_per_region_aq <- aquaculture %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  filter(year == 2017) %>% 
  select(region, chl_name = species, region)


 # Aquaculture per region 
  
  aquaculture_per_region <- aquaculture %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  group_by(year, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000) %>% 
  left_join(main_species_per_region_aq) %>% 
  left_join(sp_index) %>% 
  select(annual_tons, sd, species =en_name, region) 
  
  
# Reordering regions
  aquaculture_per_region$region <- factor(aquaculture_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

  aquaculture_per_region <- aquaculture_per_region %>% 
  ggplot(aes(x = region, y = annual_tons, fill = species))+
  geom_bar(stat = 'identity', fill= wes_palette('Darjeeling1')[2]) +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  labs(y="Harvest from aquaculture (Thousand Tons)", x="Region")+
  coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  #scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  ggtheme_plot()
  
  aquaculture_per_region

  # Most important farmed species per region (Appendix)
  
  aquaculture_species_per_region <-aquaculture %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons), chl_name = species) %>%
  left_join(sp_index) %>% 
  filter(tons != 'NA') %>% 
  group_by(year, sc_name, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>%
  left_join(sc_name_to_en) %>% 
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
  
aquaculture_species_per_region$region <- factor(aquaculture_species_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

```

#### Main Traded marine species 
 
Chile is a major exporter of seafood products. According to SUBPESCA, Chile has been exporting at a rate of 1.3 ± 0.09 million tons per year since 2012 (@subpesca_informe_2017). Around 72% of the landings from fisheries and aquaculture are exported outside the country (@ifop_monitoreo_2015). **Figure 5** shows the six most exported groups of species in terms of volume, which are Jack mackerels, Salmons, Mussels, Mollusks, Trouts and Algae. For most of these groups of species exports are much larger than imports excpet from the Jack mackerel for which the net trade is almost zero. Salmon exports are on the rise. 

```{r International trade per species, echo = FALSE, warning=FALSE, message=FALSE}

# Importing export and import data 
sp_exports<- read_csv(here("clean_databases/exports_ADUANA.csv"))
sp_imports<- read_csv(here("clean_databases/imports_ADUANA.csv"))

# Matching database
sp_match_trade <- read.csv(here("clean_databases/intermediate/export_groups_index.csv")) %>% 
  select(-species, group = grupo_exportacion)

# 6 most exported tons species
six_exports <- sp_exports%>% 
  filter(year != 2018) %>% 
  #mutate(chl_name= str_to_upper(grupo_exportacion))%>% 
  #left_join(sp_match_trade) %>% 
  group_by(year, grupo_exportacion) %>% 
  summarise(tons = round(sum(tons)/1000,2), value = sum(value)) %>% 
  select(-value) %>% 
  group_by(year) %>%
  mutate(rank = rank(desc(tons))) %>% 
  filter(rank < 5) %>% 
  select(-rank)

# Imported tons per species
six_imports <- sp_imports%>% 
  filter(year != 2018) %>% 
  left_join(species_match_import, by = 'name_sp') %>% 
  select(year, tons, group) %>%
  group_by(year, group) %>% 
  summarise(tons = sum(tons))


# What are the sp with more imported volumes for the 10 most exported products

net_trade <- ten_exports %>% 
  left_join(sp_imports, by = c('group', 'year')) %>% 
  rename( exports = tons.x, imports = tons.y) 

net_trade$imports[is.na(net_trade$imports)] <- 0

# Names to english

sp_index_trade <- sp_index %>% 
  select(group, en_group) %>% 
  unique()

# Plot for 6 most exported species 

net_trade = net_trade %>%
  left_join(sp_index_trade) %>% 
  select(year, en_group, exports, imports) %>% 
  gather(key = operation, value= tons, 3:4) %>% 
  ggplot(aes(x = as.factor(year), y = tons, fill = operation))+
  geom_bar(stat = 'identity')+
  facet_wrap(~en_group, scales = 'free')+
  labs(y="Thousand tons", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black'))

net_trade


```

### National Seafood Supply

#### Species contribution to Human and Non-human consumption

The main species 

```{r Species in HC products, echo = FALSE, warning=FALSE, message=FALSE}

processing <- read.csv(here("clean_databases/processing.csv"))

## Species matching english scientific
sc_to_en <- sp_index %>% 
  select(sc_name, en_name) %>% 
  unique()

## Most important species based on  contribution to HC with Aquaculture

sp_HC_with_Aq <- processing %>%
  filter(m_p == 'M') %>% # Puedes comparar M y P si encuentras q es interesante
  left_join(chl_to_sc) %>%
  group_by(year, sc_name, use) %>%
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, use) %>% 
  top_n(5, tons) %>%
  left_join(sc_to_en) %>% 
  rename(species = en_name) %>% 
  ggplot(aes(x= as.factor(year), y = tons/1000, fill = species)) +
  geom_bar(stat= 'identity') +
  labs(y="Tons (Thousand tons)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Royal1")))+
  facet_wrap(~use, scale = 'free')+
  ggtheme_plot()
  
## Without Aquaculture (this assumed that all Aq goes to processing facilities). Fr this one we use raw material no?

sp_HC_no_Aq <- processing %>%
  filter(m_p == 'M') %>% # Puedes comparar M y P si encuentras q es interesante
  left_join(chl_to_sc) %>%
  group_by(year, sc_name, use) %>%
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  left_join(Aq, by = c('sc_name', 'year')) %>% 
  rename(tons_raw = tons.x, tons_Aq = tons.y) 

sp_HC_no_Aq$tons_Aq[is.na(sp_HC_no_Aq$tons_Aq)] <- 0
  
sp_HC_no_Aq <- sp_HC_no_Aq %>%
  mutate(tons = tons_raw - tons_Aq) %>% 
  filter(use == "HC") %>% 
  filter(!sc_name %in% c("Mytilus chilensis", "Salmo salar", "Oncorhynchus mykiss", "Oncorhynchus kisutch")) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year) %>% 
  top_n(5, tons) %>% 
  left_join(sc_to_en) %>% 
  rename(species = en_name) %>% 
  ggplot(aes(x= as.factor(year), y = tons/1000, fill = species)) +
  geom_bar(stat= 'identity') +
  labs(y="Tons (Thousand tons)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Royal1")))+
  #facet_wrap(~use, scale = 'free')+
  ggtheme_plot()
```

Considering aquiculture, results show that the main species that contribute to human consumption are Atlantic Slamon (*Salmo salar*) and the Chilean Mussel (*Mytilus chilensis*), two of the main aquaculture products from Chile (Figure sp_HC_with_Aq). On the other hand species that are used for non-human consumptions are Peruvian Anchoveta (*Engraulis ringens*) Araucarian Herring (*Strangomera bentincki*), the two main wild-caught fisheries of the country (Figure Figure sp_HC_with_Aq). The contribution of each specie to human consumption and non human consumption is relativly stable through time.

```{r figure  }
sp_HC_with_Aq
```

Within the wild-cought species, the one that have historically contributed more to human consumption products are: Chilean Jack Mackerel (*Trachurus murphyi*), Jumbo flying squid (*Dosidicus gigas*), and in a minor contribution Sea Urchins (*Loxechinus albus*) (Figure sp_HC_no_Aq). These results also show that the amount of with caught fiheries that go to non-human consumption is more than the double and some years the triple of that that goes to human consumption.


```{r figure species for HC w_no aq echo= FALSE, message=FALSE, warning=FALSE}
sp_HC_no_Aq
```


In Chile, almost all $L_i$ and $Aq$ are send to processing facilities while only some of the $L_a$ is processed. We assume that the rest of the artisanals' catch that is not sent to processing facilities is sold as non-processed products for human consumption ($NP_{HC}$).


```{r Non-processed Human consumption, echo= FALSE, message=FALSE, warning=FALSE}

## Raw material in processing plants with scientific name per region and year

processing_minus_ind_Aq <- processing %>% 
  filter(m_p== 'M', tons !=0) %>% 
  select(-use, -m_p, -product) %>% 
  left_join(chl_to_sc) %>% 
  select(-chl_name) %>% 
  group_by(year, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  left_join(l.i, by= c('year', 'sc_name')) %>% #How much of each sp comes from Li? 
  mutate(tons_raw = as.numeric(tons.x), tons_ind = as.numeric(tons.y)) %>% 
  select(-tons.x, -tons.y) %>% 
  left_join(Aq, by=c('sc_name', 'year')) %>% #How much of each sp comes from Aq?
  rename(tons_Aq = tons)

processing_minus_ind_Aq$tons_ind[is.na(processing_minus_ind_Aq$tons_ind)] <- 0
processing_minus_ind_Aq$tons_Aq[is.na(processing_minus_ind_Aq$tons_Aq)] <- 0

processing_minus_ind_Aq <- processing_minus_ind_Aq %>% 
  mutate(tons_art= tons_raw - tons_ind - tons_Aq) 

## Substracting artisanal catch that goes to faiclities to artisanal landings to get the rest (non-processed human consumption)

NP_HC <- l.a %>%
  left_join(processing_minus_ind_Aq, by= c('sc_name', 'year')) %>% 
  select(year, sc_name, tons_landed = tons, tons_processed= tons_art) %>% 
  mutate(tons_processed = ifelse(tons_processed < 0, 0, tons_processed)) %>% 
  mutate(tons_NP_HC = tons_landed - tons_processed)

## Plot

NP_HC_plot <-NP_HC %>% 
  arrange(year, desc(tons_NP_HC)) %>%
  group_by(year) %>% 
  top_n(5, tons_NP_HC) %>% 
  left_join(sc_to_en) %>% 
  select(year, species= en_name, tons = tons_NP_HC) %>% 
  filter(year != 2015) %>%  # There is no data for oil
  ggplot(aes(x= as.factor(year), y = tons/1000, fill = species)) +
  geom_bar(stat= 'identity') +
  labs(y="Tons of non-processed HC (Thousand tons)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Moonrise1")))+
  ggtheme_plot()
  
  
```

```{r}
NP_HC_plot
```




** Main fisheries for food supply **
** Total and per capita food consumption in the country **
** Regional differences in per capita and total consumption **
We where not able to do a regional analysis. Even though most of the data is provided at a regional level. We realized that is was not possible to assume that what was landed in a region was processed in that same area. According to a study done along the Chilean coast, 62% of the population form the north declares to eat fish once a week. in the north centre region that numbers decreases to 54% and it reaches its lowest point in the Metropolitan Region (42%). In the center south, 48% each fish at least once a week while in the south 48% (@fundacion_chile_chile_2016)

#### National Economic Participation

Fisheries, as an economic sector in Chile, considers activities of extraction and farming of marine and freshwater products. This sector contributes to around 0.4% of the national GDP although it is a highly variable sector (@central_cuentas_2017). An estimate of 2014, indicates that 90% of the sales of the sector came from farmed products (@sence_reporte_2015). What reflects the relevance of aquaculture in economic terms relative to wild fisheries. Yet, the extractive sector has a higher number of firms, 58% of the total (@sence_reporte_2015). Fishing manufacturing is closely relates to fisheries. Based on the last input-output matrix for the country's economy, for every 100 US dollars spent in the fisheries sector, another 52.2 dollars are generated in the manufacturing sector. Other sectors affected by fisheries are the financing sector (24.8 dollars generated every 100) and transportation and comunication (13.4 dollars generated every 100). 

The contribution of fisheries to regional GDP suggests that the economic importance of the sector to regional economies varies widely. Ranging from almost 30% in the XI region to 0% in the metropolitan (RM) and IX region (*Figure 5*).  

```{r Fisheries contirbution to regional and national PIB, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= 'Percent of the regional GDP contirbuted by the fisheries sector averaging data between 2013 and 2016. Data source: Banco Central de Chile.'}

PIB_region <- read.csv(here("information/raw_databases/Banco_Central/PIB_region.csv")) %>% 
  mutate(perc_contr = (PIB_fish * 100)/ PIB_total) %>% 
  group_by(region) %>% 
  summarise(mean_perc_contr = mean(perc_contr), sd= sd(perc_contr))

# Reordering regions
PIB_region$region <- factor(PIB_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

PIB_region_plot <- ggplot(PIB_region, aes(x= as.factor(region), y= mean_perc_contr)) +
  geom_bar(stat = 'identity', fill = wes_palette("Darjeeling1")[2]) +
  coord_flip()+
  xlab('Region')+
  ylab('Contribution of fisheries industry to regional GDP (%)')+
  scale_y_continuous(expand = c(0, 0), breaks = c(seq(0,40, by = 5)))+
  geom_errorbar(aes(ymin=mean_perc_contr-sd, ymax=mean_perc_contr+sd), width=.2,
                 position=position_dodge(.9)) +
  #geom_label(aes(label = paste(as.character(round(mean_perc_contr,1)))), position = )+
  ggtheme_plot() 
PIB_region_plot 

```

Seafood products are one of the most important exports of the country. IFOP valued the exportation of fisheries-related products in $6,280 million dollars in 2017 (@ifop_boletines_2017). This corresponded to ~ 9 % of the total value of national exportations. The most important exported species are the Atlantic salmon (*Salmo salar*), the Rainbow trout (*Oncorhynchus mykiss*) and the Chinook salmon (*Oncorhynchus tshawytscha*)(*Figure 6*). All three produced via aquaculture. Mussels, specifically the Chilean Ribbed Mussel (*Aulacomya ater*) and the Chorus mussels (*Choromytilus chorus*) are also among the most valuable exports. These are landed by artisanal fishers and to a lesser extent produced in aquaculture centers. Other wild caught species that are valuable in terms of epxports are the Jumbo squid and the Chilean jack mackerel (@subpesca_informe_2017). 

The most valuable sea products in terms of exports are frozen and fresh followed by fish meal and dried algae (@ifop_boletines_2017). In 2016, fish fillets were the fourth most exported product by Chile. This commodity generates around 2 billion US dollar per year and is mainly made of farmed salmon (@datachile_fish_2016). 

```{r exports per species, echo = FALSE, warning=FALSE, message=FALSE, fig.cap='Main exported species in terms of econmic value. Source: Export data from Aduanas Chile, 2018'}
 
# Importing exports data and calculating FOB per year and species
exports_ADUANA <- read_csv(here("clean_databases/exports_ADUANA.csv"))

main_sp_exports <- exports_ADUANA%>% 
  filter(year != 2018) %>% 
  mutate(chl_name= str_to_upper(especie_SERNAPESCA))%>% 
  left_join(sp_index) %>% 
  group_by(year, sc_name) %>% 
  summarise(FOB = sum(value)/1000000) %>% 
  arrange(year, desc(FOB)) %>% 
  group_by(year) %>% 
  top_n(5, FOB) %>% 
  left_join(sp_index) %>% 
  select(year, species= en_name, FOB) %>% 
  ggplot(aes(x = year, y = FOB, fill = species))+
  geom_bar(stat = 'identity') +
  labs(y="FOB (Million US dollar)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black'))+
  ggtheme_plot()

main_sp_exports
```

Based on our estimates of species-specific revenue the most important species for the artisanal sector are the Southern king crab (*Lithodes santolla*) and the Chilean sea urchin (*Loxechinus albus*) (*Figure 7*). Nontheless, this varies across years. In 2017 the Patagonian toothfish (*Dissostichus eleginoides*) and the Southern rays bream or Corvina (*Brama australis*) were the species that provided more revenue followed by the Pink cusk-eel (*Genypterus blacodes*) in the artisanal sector.

Some of these species are among the most expenssive (*Table App.6*). The Patagonian toothfish (*Dissostichus eleginoides*) has the highest mean ex-vessel price in the artisanal sector what explains in part why is one of the top 5 in term of revenues. We should notice that the fishing costs for different types of species could vary widely and we are not accounting for this in our analysis. Yet, we can qualitatively consider the costs of different types of fishing. For example, some of these species are fished by diving near the shore like the Chilean abalone (*Concholepas concholepas*) and the Chilean sea urchin (*Loxechinus albus*) while others like the Patagonian toothfish (*Dissostichus eleginoides*), the Southern king crab (*Lithodes santolla*) or the Swordfish (*Xiphias gladius*) involve offshore fishing, which tends to be much more expensive. The gathering of algae from the shore is arguably one of the less costly extractive activities for the artisanal sector but it is not among the ones that generate more revenue. We could not generate similar estimates for the industrial and aquaculture subsectors due to lack of data. 

```{r Species-specific revenues artisanal, echo = FALSE, message= FALSE, warning= FALSE, fig.cap= 'Species that generated the most revenue for the artisanal sector over the past five years based on records on landings and ex-vessel prices. Source: records from SERNAPESCA, 2018.'}

# importing cleaned prices dataset and aggregating at the national level
prices_art <- read_csv(here("clean_databases/prices_art_SERNAPESCA.csv")) %>% 
  filter(!(species == 'NA'))
prices_art[is.na(prices_art)] <- as.numeric(0) 
  
# Converting to 2018 dollars
prices_art <- prices_art %>% 
  mutate(price_dollar = ifelse(year == 2013, ((price/mean_uf_2013)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2014, ((price/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2015, ((price/mean_uf_2015)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2016, ((price/mean_uf_2016)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ((price/mean_uf_2017)*mean_uf_2018)/from_ChP_to_dollar_dec_2018)))))
  
# Most important species artisanal based on national revenue
most_revenue_sp_art <- prices_art %>% 
  mutate(revenue= (price_dollar * tons), chl_name = str_to_upper(species)) %>%
  left_join(sp_index) %>%
  group_by(sc_name, year) %>% 
  summarise(national_revenue = round(sum(revenue),2)) %>% 
  arrange(year, desc(national_revenue)) %>% 
  group_by(year) %>% 
  top_n(4, national_revenue) %>%
  left_join(sp_index) %>%
  select(year, national_revenue, species= en_name) %>% 
  ggplot(aes(x = as.factor(year), y = national_revenue, fill= species)) + 
  geom_bar(stat = 'identity')+
  labs(y="Revenues (2018 US dollars)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling2"), wes_palette("Darjeeling1")))+
  ggtheme_plot()

most_revenue_sp_art

# Most expensive species artisanal (Appendix)
most_expensive <- prices_art %>%
  mutate(chl_name = str_to_upper(species)) %>% 
  left_join(sp_index) %>% 
  group_by(sc_name) %>% 
  summarise(mean_price = round(mean(price_dollar),2), sd= round(sd(price_dollar), 2))%>%
  arrange(desc(mean_price)) %>% 
  top_n(10, mean_price) %>% 
  left_join(sp_index) %>% 
  select(species = en_name, mean_price, sd) %>% 
  unique()

```

Fisheries and aquaculture are considered to be the activities that provide less direct employment in the country after agriculture (@sense_reporte_2015). Yet, the number of people employed in the sector has been on the rise since 2010 (@sence_reporte_2015). We estimated a total of ~ 278,000 job positions generated by fisheries and aquaculture. This estimate considers artisanal and industrial fishers, workers in processing facilities and aquaculture centers and indirect jobs generated by the manufacturing sector (*Table 2*). It does not include other indirect jobs generated by the extractive activities and its commercialization. Thus, the number of total people directly and indirectly employed by fisheries and aquaculture in Chile could be higher. 

```{r Employment in fisheries from different sources, echo = FALSE, message= FALSE, warning= FALSE}

# Number of employees

Artisanal_extraction <- c('88,968', '2018', 'RPA (SERNAPESCA, 2018)')
Industrial_extraction <- c('3,525', '2018', 'RPI (SERNAPESCA, 2018)')
Aquaculture <- c('17,631', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Manufacturing <- c('65,451', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Indirect <- c('102,758', '2018', 'Own estimate based on IFOP,2015 multipliers' )

table_employment <- rbind(Artisanal_extraction, Industrial_extraction, Aquaculture, Manufacturing, Indirect) %>% 
  data.frame() %>% 
  mutate(Stage_supply_chain = c('Artisanal Extraction', 'Industrial Extraction', 'Aquaculture Centers', 'Manufacturing', 'Indirect employment'), Employees= X1, Year_estimate= X2, Source= X3)
table_employment <- table_employment [4:7]

knitr::kable(table_employment, col.names = c('Stage of the supply chain', 'Employees', 'Year of the estimate', 'Source'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most recent estimates of people employed in different stages of the supply chain of sea products in Chile. Note that the estimate of indirect employment only considers indirect employment generated by the manufacturing stage.') %>% 
kable_styling(latex_options = 'hold_position')

```

People employed in the fisheries and aquaculture activities has an average monthly income of 579 US dollars, the lowest mean income in the country after the one generated by the agricultural sector (@sence_reporte_2015). This value ranges from 258 US dollars for people working independently and part-time to 1,160 US dollars for full-time employers (*Table App. 7*).

```{r transforming average income from chilean pesos 2014 to dollars 2018, echo =FALSE}

mean_income_sector_chp <- 345541 # Based on report from SENSE 2015
mean_income_sector_dollar_2018 <- ((mean_income_sector_chp/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018
```

```{r data on incomes, echo = FALSE, message=FALSE, warning= FALSE}

full_time <- c(692786, 210046, 528532, 227808, 402353)
part_time <-c(422099, 154040, 277600, 172445, 180434)

table_incomes_sector <- rbind(full_time, part_time) %>% 
  data.frame() %>% 
  mutate(Type = c('Full time', 'Part time'), Employer= round(((X1/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), Independent= round(((X2/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), Dependent_Contract= round( ((X3/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, 2), Dependent_no_contract= round( ((X4/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), mean_income_fisheries= round( ((X5/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018),2)

table_incomes_sector <- table_incomes_sector[6:11]

```

On average, people employed in the fisheries and aquaculture sector are a younger and less educated relative to the national average and are mostly men (*Table App. 8*). However, there is high variation within subsectors, for example, artisanal fishers tend to be older @tam_gone_2018. Based on estimates from the New national survey of employment (NENE, 2015) only 57.7% of the people employed in the sector is formaly employed by an employer, the rest work independently (@sense_reporte_2015). 

```{r demographics employees, echo = FALSE, warnings = FALSE, message= FALSE}
# Demographics of employees

Average_age <- c('42.1', '43.2')
Education_level <- c('9.5', '11.9')
Women_perc <- c('8.2','40')

table_demo_employment <- rbind(Average_age, Education_level, Women_perc) %>% 
  data.frame() %>% 
  mutate(Characteristic = c('Average age', 'Years of formal education', 'Percentage of women'), Fisheries_and_Aquaculture= X1, National= X2)
table_demo_employment <- table_demo_employment [3:5]


```

The artisanal sector corresponds to fishing activities performed by vessels equal to or smaller than 18 meters in length and by collectors in the inter-tidal zone (@subpesca_panorama_2018). Based on the current RPA, there are around 89,000 artisanal fishers along with 12,700 vessels in Chile. *Figure 6*, shows the number of artisanal fishers per region and gender. They are concentrated in the X and VIII regions and are mostly men. The main activity in the artisanal sector is harvester or collector, this mostly corresponds to algae collection on the shore. All regions present similar composition of fishing activities (*Figure 7*). However, the X region seems to rely more on benthic resources for employment than the others since they have a greater share of divers and algae collectors.

```{r Artisanal employment distribution and gender, echo= FALSE, warning= FALSE, message= FALSE, fig.cap= 'Number of artisanal fishers in each region by gender. Source: National record of artisanal fishers (SERNAPESCA, 2018)'}

# Importing national records of artisanal fisheres
rpa <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "RPA PESCADORES", skip = 2, n_max = 138294) %>% 
  rename(region = "Regi\u00f3n", year = "A\u00f1o Inscripci\u00f3n", gender = Genero)

art_reg_gender<- rpa %>% 
  group_by(`RPA Pescador`, region, gender) %>% 
  summarise(n = n()) %>%
  group_by(region, gender) %>% 
  summarise(count= n()) %>%
  mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV')))))))))))))))) %>% 
  select(gender, count, Region) 

# Reordering regions
art_reg_gender$Region <- factor(art_reg_gender$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

art_reg_gender <-  ggplot(art_reg_gender, aes(x= as.factor(Region), y= count, fill = gender)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Region", x="Number of people", title = 'Artisanal')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  ggtheme_plot()

art_reg_gender

```

```{r Artisanal distirbution by activity, echo = FALSE, warning= FALSE, message= FALSE, fig.cap= 'Number of permits issued in each region for different activities. One person can be registered in multiple activities. Source: National record of artisanal fishers (SERNAPESCA, 2018) '}


# Number of boat owners, divers, fishers and gatherers in each region.

rpa_activity <- rpa %>% 
  select(region, year, ARMADOR, BUZO, RECOLECTOR, PESCADOR) %>% 
  gather(activity, practice, ARMADOR:PESCADOR) %>% 
  filter(practice == 'X') %>% 
  group_by(region, activity) %>% 
  summarize(count = n()) %>% 
  mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV'))))))))))))))), activity = ifelse(activity == 'ARMADOR', 'Boat owner', ifelse(activity == 'BUZO', 'Diver', ifelse(activity == 'RECOLECTOR', 'Collector', 'Fisher')))) 

rpa_activity$Region <- factor(rpa_activity$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

rpa_activity <- rpa_activity %>%  
  ggplot(aes(as.factor(Region), count, fill = activity)) +   
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Region", x="Number of permits", title= 'Artisanal')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  ggtheme_plot()

rpa_activity
```

The industrial sector comprises activities performed by vessels larger than 18 meters. Currently, there are 475 industrial vessels and 164 vessel owners according to national industrial records (RPI), from which only 25 are individual owners and the rest are firms (@maturana_mujeres_2017). Besides, 3,500 jobs are generated through vessels operations (@subpesca_pesca_2018). We could only find information on the distirbution of vessels (*Figure 8*). We observe that most industrial  extractive activities occur in the VIII region, as suggested by landings estimates discussed above. Still, vessels are registered along the etire coast. Some of the firms that own industrial vessels also own processing facilities whose employees are considered as part of the manufacturing stage. 

```{r Distribution of industrial vessels, echo= FALSE, warning= FALSE, message=FALSE, fig.cap= 'Number of industrial vessels per region. Source: National record of industrial fisheries (SERNAPESCA, 2018)'}

# Importing national records of industrial fisheres
rpi <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "RPI", skip = 2, n_max = 473) %>% 
  rename(region = "Regi\u00f3n") %>% 
  group_by(region) %>% 
  summarise(count = n()) %>% 
    mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV')))))))))))))))) 

# Reordering regions
rpi$Region <- factor(rpi$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

ind_vessels_reg <-  ggplot(rpi, aes(x= as.factor(Region), y= count)) +
  geom_bar(stat = 'identity', fill = wes_palette("Darjeeling1")[2]) +
  coord_flip()+
  labs(y="Region", x="Number of vessels", title = 'Industrial')+
  scale_y_continuous(expand = c(0, 0))+
  ggtheme_plot()

ind_vessels_reg

```


A rigourus study by IFOP estimated a total of ~50,000 job positions in the manufacturing of marine products in processing facilities in 2014 (@ifop_monitoring_2015). However, a more recent report form SERNAPESCA reports a total of 65,451 jobs (@maturana_mujeres_2017). *Figure 9* displays the geographic distribution of people employed in the manufacturing sector per gender and type of product for 2015. Employment in the processing facilities is dominated by human consumption (HC) products, with ~ 40,000 job positions. Products for animal consumption (AC) generated around 5,000 positions while algae products employed less than 3,000 people. Most people employed by the manufacturing sector are located in the southern part of the country and the majority are men. 

```{r Employment in facilities, echo= FALSE, warning=FALSE, message=FALSE, fig.cap= 'Number of employees in the fisheries manufacturing sector in each region by gender in 2015. AC refers to products for animal consumption and HC products for human consumption. Source: Report by IFOP based on facilities survey (IFOP, 2015)'}

# Importing IFOP, 2015 estimates of employees per type of facility

manufacturing_employment <- read.csv(here("information/raw_databases/IFOP/employment_per_processing_type_area.csv")) %>% 
  rename(M = ï..man_employee, F = women_employee, Product = product, Zone = region) %>% 
  gather(key = Gender, value = Employees, 1:2)

# Reordering regions
manufacturing_employment$Zone <- factor(manufacturing_employment$Zone , levels = c("South", "Mid South","North" ))

manufacturing_employment <- ggplot(manufacturing_employment, aes(x= as.factor(Zone), y= Employees, fill = Gender)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Employees", x="Zone", title = 'Manufacturing')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  facet_wrap(~Product, scales = 'free')+
  ggtheme_plot()
  
manufacturing_employment 

```

There are 17,631 people employed among the 3,683 aquaculture centers currently registered in Chile (@maturana_mujeres_2017). From these, ~4,500 are women. *Figure 10* shows the distribution of aquaculture centers per region and type of farmed species. The main type of product being farmed is fish, mostly salmon, followed by molluscs and algae. The X region hosts more than half of the aquaculture centers in the country. 

```{r Aquaculture centers per region and type of species, echo= FALSE, message= FALSE, warning= FALSE, fig.cap= 'Number of aquaculture centers in each region by type of species in 2017. Source: Report Subsector Acuicultura 2017 by SERNAPESCA (SERNAPESCA website, 2017)'}

# Importing number of aquac. centers, SERNAPESCA

centers_aquaculture <- read.csv(here("information/raw_databases/SERNAPESCA/employment/centers_per_species_aquaculture.csv")) %>% 
  rename(Region = ï..Region) %>% 
  gather(key = Species, value = 'Centers', 2:7)

# Reordering regions
centers_aquaculture$Region <- factor(centers_aquaculture$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

centers_aquaculture <- ggplot(centers_aquaculture, aes(x= as.factor(Region), y= Centers, fill = Species)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Centers", x="Regions", title = 'Aquaculture centers')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  #facet_wrap(~Species, scales = 'free', ncol = 2)+
  ggtheme_plot()
  
centers_aquaculture 

```

** Species-specific estimates of economic contribution **

## Critical Analysis

### Data uncertainty 

Most of our analisis relied on publically available data collected by govermental institutions. Although data collection has improved over the years and there have been important national efforts towards transparency, there are still reasons to think that data may be biased. Particularly concerning is landing data since it is mainly generated via landings declarations. This requiere fishers to announce their intention to land their catch to the authority in charge in their macrozone who must weight and certify the catch. Without the certification of their catch fishers cannot sell it in legal establishments and they risk their permits and penalty fees. Still, the incomplete enforcement may generate incentives to undereport. Illegal unreported and unregulated (IUU) fishing has been evidenced in Chile, even in systems where fishers are expected to have high incentives against it like Territorial users' rights for fishing (@oyanedel_illegal_2018). 

The lack of consistency in data collection between and within agencies brings another source of uncertaity for an analysis like ours. A major challenge was to match trade and landings data. SERNAPESCA collects data under the common names while Aduanas Chile record exported and imported products under the international Harmonized System nomenclature. To overcome this issue we use the information provided by Aduanas Chile on the species contained in some types of products. This allowed us to aggregate products per species and match volumes and values to landing record by SERNAPESCA. In addition, SERNAPESCA applies several variations of a species common name between databases. This difficults the calculation of aggregated estimates. We solve this by generating a dataset that categorize all the names currently used by SERNAPESCA under their respective scientific name to be able to aggregate data on a particular species. Although we did it as carefully as we could and we double checked the code several times, this kind of steps are always prone to human error bringing more uncertainty to the analisis. As for 2013, SERNAPESCA is using an online plataform that facilitates the reporting of catches to fishers and improves data automatization. Yet, there are still an important number of landings that are reported and recorded on paper. We expect that this new technology will bring more consistency in data collection facilitating this types of analsis in the future. 

A major information gap we identified is the lack of publically available data on ex-vessel prices and employment for the industrial sector. To our knowledge neither SERNAPESCA, IFOP nor SUBPESCA collect data on this regard. Thus, identifying what were the most valuable species for the industrial sector was not possible. An approach we were unable to apply due to time constraints is to derive ex-vessel prices from trade data (@melnychuk_reconstruction_2017). Contrary to what we expected, the fisheries sector for which we found most available information was the artisanal sector. We were able to get a reliable count of the direct employment generated by the artisanal sector thanks to the National Record of Artisanal Fishers (RPA). Still, we cannot ensure we are not underestimating the number of people that rely on fishing. We know enforcement is weak specially in remote areas of the country therefore there is room for illigal activity. We were also able to estimate species-specific revenues by using landings and prices of first transactions. Finally, information on employees in processing facilities was also available. However, some values were unrealistic since they were larger than the entire population of the country. Thus, we prefered relied on a survey study performed by IFOP to fullfill this estimate.   

We also lacked data on the transportation of landings and products between Chilean regions and final destination. This prevented us from applying our model of food supply at the regional level. The problem is that the catch landed in one region could be processed in a different one and then the final product sold in a third region. Running the regional analysis without accounting for this issue would have led to incorrect estimates of food consumption. 

### Estimating National Seafood Supply

Chile is considered one of the main countries in fishery production worldwide (@fao_estado_2016), however several reports and studies suggest the Chilean population does not consume as many marine products (@gonzales_solo_2018). On average, people consume 10-13 kg a year, compared to the 20 kg a year recommended by Food and Agriculture organization (2016). Looking into what type of seafood the Chilean population consumes, a study done in 2012 shows that 60% of the seafood consume are pelagic fishes, 12% are mollusk and 2% are crustaceans (@villena_marcelo_diagnostico_2012).

According to the National Health Survey (Encuata Nacional de Salud, ENS), only 9.2% of the chilean population eats fish at least two times a week as recommended. This see value seemed to be decreasing since the early 2000s where this percentage was 10.7% (@ENS). World data (@ritchie_our_nodate) also shows a decrease in the amount of seafood consumed the Chilean population from ~16kg consumed in 2004 to 12.5kg in 2013.

The frequency of seafood consumption varies along socio-economic groups in Chile, according to a study done in 2016 by Fundación Chile and Adimark (@fundacion_chile_chile_2016), 55% of the high-income group consumes seafood at least once a week compared to the 38% of the lower-income group. It has also been identified (@ENE) that people with less years of schooling tend to eat less fish (only 6.3 percent eats seafood twice a week) that people that have compleated high shool education, where 11.2% eats as frequntly. 

Within the Chilean population there is a strong belief that seafood is expensive. However, there is wide range of fish prices. There are expensive products sucha as salmon or lobsters and afordable alternatives such as Chilean Jack Mackerel (*Trachurus murphyi*), Jumbo Flying Squid (*Dosidicus gigas*) and Southern Rays Bream (*Brama australis*).

FUNDACION CHILE
En el mismo estudio, cuando se pregunta a los chilenos cuál es su formato de consumo en un mes normal, el 84% de las personas contesta que pre ere el pescado enlatado, un 80% lo consume fresco y sólo un 37% de las personas se inclinan por el producto congelado. Esta gran diferencia respecto del consumo de pescado congelado representa un desafío a
la hora de derribar mitos que puedan existir en la población sobre la inocuidad, contenido nutricional, entre otros factores.


### Estimating National Economic Participation

Aquaculuture is the most important activity in the fisheries sector in terms of economic value. Their profits come mainly from the exportation of salmon prodcts. This positions salmons as the group of marine species that contribute the most to the country's economy. The Chilean jack mackerel (*Trachurus murphyi*) and the Jumbo squid (*Dosidicus gigas*) are the wild-caught species that have contributed the most in exports over the pasts years. The latter is use for DHC and is targeted by both the artisanal an industrial sector. Yet, current discussion on closing access to the industrials. 

Although the direct employment in aquaculture centers is low relative to the number of artisanal fishers, it is very likely that an important share of the job positions in the manufacturing sector are being supported by aquaculture. Especially in the South where most job positions for manufacturing of human consumption products are concetrated. The role of Algaes in terms of employment for the artisanal sector is also worth notice. The revenues of most these species are not particularly high, however the lower costs of extraction relative to other types of fishing make algae collection an attractive form of livelihood for coastal communities. 



## Appendix

```{r Appendix, echo = FALSE, message= FALSE, warning= FALSE}

# Top 10 landed species artisanal (for each year between 2013-2017)
knitr::kable(top_10_art, col.names = c('Ranking', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Top 10 landed species per year in the artisanal sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Top 10 landed species industrial (for each year between 2013-2017)
knitr::kable(top_10_ind, col.names = c('Ranking', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Top 10 landed species per year in the industrial sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most landed species per region artisanal (for each year between 2013-2017)
knitr::kable(main_species_reg_art, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per and year region in the artisanal sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most landed species per region industrial (for each year between 2013-2017)
knitr::kable(main_species_reg_ind, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per region and year in the industrial sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most harvested species per region (for each year between 2012-2017)
knitr::kable(aquaculture_species_per_region, col.names = c('Region', '2012', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most harvested species per region and year. When no data is available means there were no aquaculure that year in that region. Source: Harvest records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most expensive species in the artisanal sector
knitr::kable(most_expensive, col.names = c('Species', 'Mean price (US$/Kg)', 'SD Price'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most expensive species in Chile based on their mean ex-vessel price between 2013 and 2017. Source: Data on ex-vessel prices for artisanal sector, SERNAPESCA, 2018.') %>% 
  kable_styling(latex_options = 'hold_position')

# Mean income for different times of employees in the fisheries and aquaculture sectors
knitr::kable(table_incomes_sector , col.names = c('Type', 'Employer', 'Independent', 'Dependent with contract', 'Dependent without contract', 'Mean income'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Monthly incomes of fisheries sector in 2018 US dollars. Source: Elaborated based on data from New national employment survey (NENE) in 2015 and published in SENSE, 2015') %>% 
  kable_styling(latex_options = 'hold_position')

# Demographics of people employed in the fisheries and aquculuture sector
knitr::kable(table_demo_employment, format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Average demographic characteristics of people employed in the fisheries and aquaculture sector in Chile. Source: Elaborated based on data from New supplementary income survey (NESI) in 2015 and published in SENSE, 2015') %>% 
kable_styling(latex_options = 'hold_position')

```


## References

