---
title: Contributions of marine capture fisheries to the domestic livelihoods and seafood
  consumption of Brazil Chile and Peru
author: Juliano Palacios Abrantes, Camila Vargas, Daniel Viana, Maria I. Rivera, Renato Gozzer, Rocío López, Santiago de la Puente, Andrés M. Cisneros-Montemayor
date: "December 21, 2018"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
header-includes:
  - \usepackage{float}
  - \floatplacement{figure}{H}
subtitle: Report prepared for Oceana
---

<!-- # Comments -->
<!-- Chunk, description -->

<!-- ## Overall -->
<!-- - Should we use FAO and SAU or just one? FAO has aqua. trade and catch data while SAU has also value and industrial vs subsistance... -->
<!-- - Subsistance fisheries discussion for countries -->
<!-- - Comment on NSS, Does it make sense? -->
<!-- - Do they eat Kelp in Chile? O_o -->
<!-- - Graphs names to common name (SAU) -->
<!-- - Tons, Tones or Tonnes? -->

<!-- ## Brazil -->
<!-- - Produção em São Paulo 1 milhão ou 100? Sao_Paulo_Production_G -->

<!-- ## Chile -->

```{r setup, eval = T, echo=F, warning=F,message=F,results='hide'}

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R
# Please include any package used in your analysis

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

#### Library ####
packages <- c(
  "here",
  "readxl", # Read dataframe
  "readr",
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  "cowplot",
  "ggpubr",
  # "plotly",
  "wesanderson",
  "png",
  "grid",
  "stringr",
  "purrr", #for loading multiple datasets
  # "rgdal",
  # "RColorBrewer",
  "knitr",
  "kableExtra",
  "data.table",
  "ggrepel",
  "gridExtra",
  # "ggmap",
  # "rgeos",
  "sf",
  "rgdal", #Spatial analysis
  "tools" #Spatial analysis
)

suppressWarnings(
  suppressMessages(
    suppressPackageStartupMessages(
      ipak(packages)
    )
  )
)

# Remove scientific numbers
options("scipen"=100, "digits"=4) 

```

```{r Data_Fun, eval = T, echo = F, warning = F, message = F, results = 'hide'}

#### Functions ####

# Note that the main function is on a separate file (corefx.r)
#source('Functions/corefx.R')

#### For all plots look alike ###

ggtheme_plot <- function() {
  theme(
    plot.title = element_text(size = rel(1), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    strip.background = element_blank(),
    # strip.text       = element_text(size = base_size, face = "italic"),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 10,
                               angle = 0,
                               face = "plain"),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.key = element_rect(colour = NA, fill = NA),
    legend.position  = "top",
    legend.justification = "center",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    strip.text.x = element_text(size = 16, colour = "black")
  )
}

ggtheme_map <- function(base_size = 9, Region = "NA") {
  
  theme(text             = element_text(#family = "Helvetica",
    color = "gray30", size = base_size),
    plot.title       = element_text(size = rel(1.25), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "transparent"),
    strip.background =element_rect(fill = "transparent"),
    axis.line        = element_blank(), # element_line(colour = "grey30", size = .5))
    axis.ticks       = element_blank(),
    axis.text        = element_blank(),
    axis.title       = element_blank(),
    # legend.key       = element_rect(colour = NA, fill = NA, size = 4),
    legend.position = "bottom",
    legend.justification = "center"
    # legend.key.width =unit(6,"line")
  )
}



####

# Color codes:
B_Col <- "#046C9A"
C_Col <- "#F21A00"
P_Col<- "#0B775E"

```

\clearpage

### Country-specific data, Chile

1. *SERNAPESCA* -SERNAPESCA - The Chilean National Service of Fisheries is the main entity in charge of keeping track of landings and other fisheries and aquaculture related data in Chile. Most of their data is available on their website but we also obtained additional databases upon request[C^1]. We accessed their databases on landings from wild-caught and farmed fisheries, first transaction prices (ex-vessel price) for artisanal fisheries, and landings used in different types of sea products. These databases provided information at the species and regional level. We also downloaded their data on the geographical distribution of production facilities and the types of sea products they produce. We issued data requests on SERNAPESCA’s portal to obtain more detailed information. Via this method, they provided us the national record of artisanal fishers (RPA), which lists current registered fishers and boats, and the national record of industrial vessels (RPI), which lists current registered industrial vessels. The RPA informs the region, gender and age for each artisanal fisher as well as the species-specific permits they hold. Together the RPA and RPI provide information on artisanal and industrial vessels, including length and storing capacity. SERNAPESCA also provided the number of employees of each gender per processing facility.

[^C1]: Chilean National Service of Fisheries. Data available athttp://www.sernapesca.cl/informes/estadisticas

2. *Aduanas Chile (Chilean Customes)*. is the main entity in charge of keeping records of imported and exported goods in the country. We accessed their exports and imports data by product and region via their website. Each product in the original exports and imports datasets is identified through an ID number that matches the name of the product in a code known as the Arancel Aduanero. Most products provided details on the species or groups of species involved.

3. *IFOP* - The Fisheries Development Institute is a private entity that advises fisheries management based on permanent biological, environmental and economic monitoring, and scientific reasoning. Although most of the data collected by IFOP pertains to stock assessments and biological indicators, they also generate economically relevant statistics. We reviewed their most recent annual report on the economic monitoring of Chilean fisheries for 2015. It includes databases and estimates for employment. We also extracted data from their monthly reports on exports of fisheries products, which are based on ADUANAS's data.  

4. *CCB* - The Chilean Central Bank (*Banco Central de Chile*) is the entity in charge of establishing the Chilean currency. To fulfill its function it monitors the main economic indicators per economic sector. We accessed their data on sectorial GDP (including fisheries) for each region. 

5. We reviewed several reports from other national agencies like the Undersecretary of Fisheries (SUBPESCA), NGOs and scientific publications to complement our analysis.

```{r C_Regions_Image, fig.width=6, fig.height=4, echo=FALSE, fig.align= "center", fig.cap="Regional distribution of Chile. Image from Bórquez, A.S. and Hernández A.J. 2009"}

## https://marine.rutgers.edu/~cfree/mapping-south-american-fish-stock-boundaries/
#
#img <- readPNG("Report_Figures/Chile/Chile_Regions.png")
# grid.raster(img)
# 
```


```{r parameters for conversion, eval=T, echo = FALSE}
# Based on data from the Tax National Service (SII)
mean_uf_2013 <- 22980 
mean_uf_2014 <- 23960
mean_uf_2015 <- 25022
mean_uf_2016 <- 26022
mean_uf_2017 <- 26571
mean_uf_2018 <- 27140 
from_ChP_to_dollar_dec_2018 <- 676 # Based on conversion for Dec, 2018
```

\clearpage

## Results Chile

### Fisheries Context

To evaluate the volume contribution of specific fisheries over the last five years, we used records of landings and aquaculture harvests from SERNAPESCA  (2013-2017). Using monthly data published by IFOP, we identified the most exported fish products in terms of volume. To identify the participation of different species in exported volumes we generated a new database linking types of products listed in the *Arancel Aduanero* with the names of the fished and farmed species listed by SERNAPESCA. We searched for all the products ids that refer to the species scientific name, common name or genus name as listed by SERNAPESCA and aggregated them under the species name.

#### Wild Caught Fish

##### Main wild-caught fisheries in terms of volume 

According to landing records from SERNAPESCA over the last five years, the artisanal sector lands on average 1.6 ± 0.58 million annual tonnes of marine species while the industrial sector lands around 0.9 ± 0.16 million tonnes per year. Is worth noting that large part of the artisanal landings are of algae and benthic species that are not targeted by the industrial sector. Both sectors have among their most landed species the Peruvian anchoveta (*Engraulis ringens*) and the Araucanian herring (*Strangomera bentincki*), two species that are mostly used as inputs for fish meal and oil (Figure 19). Other important species in artisanal landings are the Giant grey kelp (*Lessonia nigrescens*), exported mostly as dried algae, and the Jumbo squid (*Dosidicus gigas*), mostly commercialized fresh or frozen for human consumption (HC). The Chilean jack mackerel (*Trachurus murphyi*) is also an important fishery in terms of volume for the industrial sector, which is also mostly used in the production of fish meal and oil and to a lesser extent for direct HC. 

Over the last years the catch composition of the artisanal sector has presented no major changes except for one species of algae (Giant brown kelp, *Lessonia trabeculata*) replacing other (Red algae gracilaria, *Gracilaria spp*) within the top 5 landed species as for 2013. In the case of the industrial sector, composition of the main species is also stable over the last years. Yet, the landings of Peruvian anchoveta (*Engraulis ringens*) started decreasing in 2014 relative to previous years, this incresed the relevance of other species like the Chilean jack mackerel (*Trachurus murphyi*). Other important species for the artisanal sector are the Mote sculpin (*Normanichthys crockeri*) and the Chilean sea urchin (*Loxechinus albus*) (Table 1).

```{r Top_landed_Chile_G, echo = FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=9, fig.cap= 'Top 5 species landed in Chile by the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. Data source: Landing records from SERNAPESCA, 2018.' }

# Sepcies index
sp_index <- read_csv(here("clean_databases/sp_index.csv"))%>% 
  mutate(chl_name = str_to_upper(chl_name)) 

# Importing landings and species names 
landings <- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  select(-region, - month, - international_waters, chl_name = species) %>% 
  left_join(sp_index) %>% 
  select(subsector, year, tons, en_name)

landings$tons[is.na(landings$tons)] <- 0 

# Calculating annual landings per sector 
annual_landings_sector <- landings %>%
  group_by(year, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)) %>% 
  group_by(subsector) %>% 
  summarise(annual_tons = mean(tons), sd = sd(tons) )

# Top 5 landed per sector

# Artisanal
most_landed_art <- landings %>% 
  filter(subsector == 'artisanal') %>% 
  group_by(en_name) %>% 
  summarise(tons = sum(tons, na.rm= T)) %>% 
  arrange(desc(tons)) %>% 
  top_n(5, tons) 
most_landed_art <- most_landed_art$en_name

# Industrial
most_landed_ind <- landings %>% 
  filter(subsector == 'industrial') %>% 
  group_by(en_name) %>% 
  summarise(tons = sum(tons, na.rm= T)) %>% 
  arrange(desc(tons)) %>% 
  top_n(5, tons) 
most_landed_ind <- most_landed_ind$en_name

## Landings per year more landed and others

## Artisanal
landings_year_art <- landings %>%
  filter(subsector == 'artisanal') %>% 
  mutate(species = ifelse(en_name %in% most_landed_art, en_name, "Others")) 

##Industrials
landings_year<- landings %>%
  filter(subsector == 'industrial') %>% 
  mutate(species = ifelse(en_name %in% most_landed_ind, en_name, "Others")) %>% 
  rbind(landings_year_art) %>% 
  group_by(year, subsector, species) %>% 
  summarise(tons = sum(tons, na.rm= TRUE))


## Plot

landings_plot <- landings_year %>%
  ggplot(aes(x = as.numeric(year),
             y = as.numeric(tons)/1000000, 
             fill = species)) +
  geom_area() +
  facet_wrap(~subsector)+
  labs(y="Landings (Million tonnes)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black'))+
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 3, byrow = TRUE))

landings_plot

# Table 

Total_Top_10 <- landings %>% 
  group_by(year, en_name, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)/1000) %>%  #thousand
  arrange(year, desc(tons)) %>% 
  group_by(subsector,en_name) %>% 
  summarise(
    Mean = mean(tons,na.rm = T),
    s.d. = sd(tons,na.rm = T),
    Total = sum(tons,na.rm = T)
  ) %>% 
  group_by(subsector) %>% 
  top_n(10, Total) %>% 
  ungroup() %>% 
  select(-subsector) %>% 
  rename(Species = en_name) %>% 
  kable("latex",
        caption = "Top 10 landed (tons) species by sector in Chile from 2013-2017.",
        booktabs = T,
        digits = 0,
        format.args = list(big.mark = ",")
        ) %>% 
  group_rows("Artisanal", 1, 10, latex_gap_space = "0.5em") %>% 
  group_rows("Industrial", 11, 20, latex_gap_space = "1em")

```

These resuts are in line with those observed using international databases. SAU and FAO databases also identify Peruvian anchoveta (*Engraulis ringens*), the Araucarian herring (*Strangomera bentincki*) and the Chilean jack mackerel (*Trachurus murphyi*) as the main species. 

##### Regional Diferences in volumes of wild-caught fisheries

Landings varies widely along the coast of Chile for both artisanal and industrial sectors. Figure 20 shows the mean total landing per sector and region. The color of the bars represents the most landed species in 2017. We observe that the VIII region is by far the region with most landings from both sectors. However, the industrial sector also concentrates landings in the north of the country, while artisanals are spread throughout the country. The most landed species in 2017 also vary by region but have changed little over the last five years (Table XX Appendix and Table XX Appendix). 

```{r Landings_per_region, echo = FALSE, fig.height=5, fig.width=10, warning=FALSE, message=FALSE, fig.cap= 'Mean annual total landings per region for the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. The color represents the most landed species in 2017 and regions are ordered vrom north to south. Error bars represent one standard deviation. Note that scales are different between pannels. Data source: Landing records from SERNAPESCA, 2018.'}

# Main species in 2017 per region and sector to add color to the barplot
# Importing landings and species names 
landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv"))

  main_species_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, region, subsector) %>% 
  top_n(1, tons) %>% 
  filter(year == 2017)
  
 # Landings per region 
  
  landings_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, subsector, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(subsector, region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000)%>% 
  left_join(main_species_per_region, by = c('region', 'subsector')) %>% 
  select(subsector, region, annual_tons, sd, chl_name = species) %>% 
    left_join(sp_index) %>% 
    select(-chl_name, -sc_name, species_2017= en_name) %>% 
    filter(species_2017 != "NA")
  
# Reordering regions

landings_per_region$region <- factor(landings_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

 landings_per_region %>% 
 ggplot(aes(x = region,
            y = annual_tons, 
            fill=species_2017))+
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(~ subsector, scales = 'free')+
  labs(y="Landings (Thousand Tons)", x="Region")+
  coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual("Species",values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette('Royal1')))+
  ggtheme_plot() +
   guides(fill = guide_legend(nrow = 4, byrow = TRUE))
 
```

#### Aquaculture production

##### Main farmed fisheries in terms of volume

The main species harvested from aquaculture centers in the last 5 years have been the Atlantic salmon (*Salmo salar*), the Chilean mussels (*Mytilus chilensis*), the Coho salmon (*Oncorhynchus kisutch*), and the Red algae gracilaria (*Gracilaria spp.*) in descending order (Figure 21). These species correspond to most of the harvest. Overall, the composition of the most farmed species and its volumes have been stable over the last 5 years. However, the Rainbow trout (*Oncorhynchus mykiss*) seems to be becoming less important over recent years.  

```{r Top_Aquaculture_Chile_G, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=7, fig.cap= 'Top 5 species farmed in Chile between 2013 and 2017. Data source: Harvest records from SERNAPESCA, 2018.' }

# Importing aquaculture database
aquaculture<- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv")) %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons), chl_name = species) %>%
  left_join(sp_index) %>% 
  select(year, region, tons, en_name)

# Overall, top 5 

most_farmed <- aquaculture %>% 
  group_by(en_name) %>% 
  summarise(tons = sum(tons, na.rm= T)) %>% 
  arrange(desc(tons)) %>% 
  top_n(5, tons) 

most_farmed <- most_farmed$en_name

# Top aquaculture
aquaculture_plot <- aquaculture %>% 
  mutate(species = ifelse(en_name %in% most_farmed, en_name, 'Others')) %>% 
  group_by(year, species) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  ggplot(aes(x = year, y = tons/1000000, fill = species))+
  geom_area() +
  labs(y="Harvest (Million Tonnes)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"))) +
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

aquaculture_plot
```


##### Regional diferences in farmed fisheries

Aquaculture production is highly concentrated in the South of the country, between regions X and XII (Figure 22). Nevertheless, from the I to the IV region the most farmed resources are algaes and scallops (*Argopecten purpuratus*) (**Table App.5**). The red abalone (*Haliotis rufescens*) became the most farmed in 2013 in the V region, replacing the turbot (*Scophtalmus maximus*). Towards the south, salmons and trouts are the most farmed species along with some algaes and mussels (*Mytilus chilensis*). The most farmed species in the austral south are the Atlantic salmon (*Salmo salar*) and the Chilean mussel (*Mytilus chilensis*).

```{r Harvest from aquaculture centers per region, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4,fig.width=6, fig.cap= 'Mean annual harvest per region from aquaculture centers between 2012 and 2017. Error bars represent one standard deviation. Only showing regions with production. Database SERNAPESCA'}

# Main species in 2017 per region in case we want to add color to the barplot

  main_species_per_region_aq <- aquaculture %>% 
  mutate(tons = as.numeric(tons)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  filter(year == 2017) %>% 
  select(region, en_name, region)


 # Aquaculture per region 
  
  aquaculture_per_region <- aquaculture %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  group_by(year, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000) %>% 
  left_join(main_species_per_region_aq) %>% 
  left_join(sp_index) %>% 
  select(annual_tons, sd, species =en_name, region) 
  
  
# Reordering regions
  aquaculture_per_region$region <- factor(aquaculture_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

  aquaculture_per_region %>% 
    filter(annual_tons >=1) %>% #Remove regions without production
  ggplot(aes(x = region, y = annual_tons, fill = species))+
  geom_bar(stat = 'identity', fill= C_Col) +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  labs(y="Aquaculture production \n (Thousand Tons)", x="Region")+
  # coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  #scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  ggtheme_plot()
  
  
```


#### Fish Trade
 
Around 72% of the landings from fisheries and aquaculture are exported outside the country [@ifop_monitoreo_2015]. In Chile, exported volumes of marine products are larger than imported volumes. The main species exported is Atlantic salmon produced in aquaculture, followed by far by jack mackerel (Figure 23-A). Tuna leads the import activities followed, also by far, by jack mackerel (Figure 23-B)

```{r Trade_Chile, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=8, fig.cap="Main species exported and imported in Chile between 2013 and 2017. Thousand Tons"}

# Importing export and import data 
sp_exports<- read_csv(here("clean_databases/exports_ADUANA.csv"))
sp_imports<- read_csv(here("clean_databases/imports_ADUANA.csv"))
# Matching database
exports_index <- read.csv(here("clean_databases/exports_index.csv"))
imports_index <- read_csv(here("clean_databases/imports_index.csv"))

#### New plot 

# Determine Top 10 Exported species between 2013 and 2017 (sum imports)
Top_Exports <- sp_exports%>% 
  filter(year != 2018) %>% 
  mutate(export_name= grupo_exportacion)%>% 
  left_join(exports_index) %>% 
  group_by(trade_name) %>% 
  summarise(Tons = round(sum(tons)/1000,2), value = sum(value)) %>% #tons
  filter(!is.na(trade_name)) %>% 
  mutate(trade_name = as.character(trade_name)) %>% 
  select(-value) %>% 
  arrange(desc(Tons)) %>% 
  top_n(5, Tons) 

Five_Exports <- sp_exports%>% 
  filter(year != 2018) %>% 
  mutate(export_name= grupo_exportacion)%>% 
  left_join(exports_index) %>% 
  group_by(trade_name, year) %>% 
  summarise(Tons = round(sum(tons)/1000,2), value = sum(value)) %>% #tons
  filter(!is.na(trade_name)) %>% 
  ungroup() %>% 
  mutate(trade_name = as.character(trade_name)) %>% 
  select(-value) %>% 
  mutate(species = ifelse(trade_name %in% Top_Exports$trade_name, trade_name, "Others"), Operation = "Exports") 

  # Determine Top 10 Imported species between 2013 and 2017 (sum imports)
Top_Imports <- sp_imports%>% 
  filter(year != 2018) %>% 
  mutate(imports_name= name_sp)%>% 
  left_join(imports_index) %>% 
  group_by(trade_name) %>% 
  summarise(Tons = round(sum(tons)/1000,2)) %>% #tons
  filter(!is.na(trade_name)) %>% 
  mutate(trade_name = as.character(trade_name)) %>% 
  arrange(desc(Tons)) %>% 
  top_n(5, Tons) 

Five_Imports <- sp_imports%>% 
  filter(year != 2018) %>% 
  mutate(imports_name= name_sp)%>% 
  left_join(imports_index) %>% 
  group_by(trade_name, year) %>% 
  summarise(Tons = round(sum(tons)/1000,2), value = sum(value_us)) %>% #tons
  filter(!is.na(trade_name)) %>% 
  ungroup() %>% 
  mutate(trade_name = as.character(trade_name)) %>% 
  select(-value) %>% 
  mutate(species = ifelse(trade_name %in% Top_Imports$trade_name, trade_name, "Others"), Operation = "Imports") 

# Final plot
Five_Exports %>% 
  bind_rows(Five_Imports) %>% 
  ggplot(.,
         aes(
           x = as.numeric(year),
           y = Tons,
           fill = species
         )
         ) +
  geom_bar(stat = 'identity')+
  facet_wrap(~Operation,
             scales = 'free',
             nrow = 1#,
             # strip.position =  "right"
             ) +
  ggtheme_plot() +
  scale_fill_manual("Operation", values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black')) +
  ggtheme_plot() +
  xlab("Year") +
  ylab("Tonnes (Thousand)")

Five_Exports
```

```{r Trade_Chile_Table, echo = FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=4}

# Importing export and import data 
sp_exports<- read_csv(here("clean_databases/exports_ADUANA.csv"))
sp_imports<- read_csv(here("clean_databases/imports_ADUANA.csv"))

# Matching database
exports_index <- read.csv(here("clean_databases/exports_index.csv"))
imports_index <- read_csv(here("clean_databases/imports_index.csv"))

# Exported species
Exports_tbl <- sp_exports%>% 
  filter(year != 2018) %>% 
  mutate(export_name= grupo_exportacion)%>% 
  left_join(exports_index) %>% 
  group_by(year, trade_name) %>% 
  summarise(Exports = round(sum(tons)/1000,2), value = sum(value)) %>% #tons
  select(-value)

# Top 10 imported species
ten_imports <- sp_imports%>% 
  filter(year != 2018) %>% 
  mutate(imports_name = name_sp) %>% 
  left_join(imports_index) %>% 
  select(year, tons, trade_name) %>%
  group_by(year, trade_name) %>% 
  summarise(Imports = sum(tons)) %>% #tons
  group_by(year) %>%
  mutate(rank = rank(desc(Imports))) %>% 
  filter(rank < 10) %>% 
  select(-rank)

# What are the exports volumes for the 10 most exported products
Imports_Exports <- ten_imports %>% 
  left_join(Exports_tbl, by = c('trade_name', 'year')) %>% 
  replace(., is.na(.), 0)
  
# Totals table
Totals <- Imports_Exports %>%
  mutate("Net trade" = Imports-Exports) %>% 
    gather("Operation","Tons",3:5) %>% 
    group_by(trade_name,Operation) %>% 
    summarise(
      Total = sum(Tons),
      Mean = mean(Tons),
      s.d. = sd(Tons)
    ) 
  
# Final table to print
Imports_Exports %>%
  mutate("Net trade" = Imports-Exports) %>% 
  gather("Operation","Tons",3:5) %>% 
  spread(year,Tons) %>% 
  left_join(Totals,
            by =c("trade_name","Operation")
  ) %>% 
  arrange(Operation,desc(Total)) %>% 
  ungroup() %>% 
  select(-Operation) %>%
  filter(!is.na(trade_name)) %>% 
  rename(Species = trade_name) %>% 
  kable("latex",
    caption = "Main Imports, exports and net trade (Imports-Exports) for Chile", 
    booktabs = T,
    format.args = list(big.mark = ",")
  ) %>% 
  kable_styling(latex_options = c("repeat_header","scale_down")) %>% 
  group_rows("Exports", 1, 10, latex_gap_space = "0.5em") %>% 
  group_rows("Imports", 11, 20, latex_gap_space = "1em") %>% 
  group_rows("Net Trade", 21, 30, latex_gap_space = "1em")

```

### National Seafood Supply

#### Species contribution to Human and Non-human consumption

As a first approach to analyzed main fisheries that countribute to human conssumption in Chile, we looked into landings and aquaculture harvest that goes to processing facilities as raw material for products for human consumpotion (HC) and for non-human consumption (NHC). Our results show that the main species that enter supply chains for human consumption are the Atlantic Slamon (*Salmo salar*) and the Chilean Mussel (*Mytilus chilensis*), two of the main aquaculture products from Chile (Figure sp_HC_with_Aq). On the other hand species that are used for non-human consumption products are Peruvian Anchoveta (*Engraulis ringens*) Araucarian Herring (*Strangomera bentincki*), the two main wild-caught fisheries of the country (Figure HC_with_Aq). A third importatn specie for NHC is the Giant gray kelp (*Lessonia nigrescens*) and the Chilean Jack Mackerel (*Trachurus murphyi*) which contributs similarly to both HC and NHC products. The relative proportion of contribution of each specie to human consumption and non human consumption is relativly stable through time.

```{r Species in HC products, echo = FALSE, eval=F, warning=FALSE, message=FALSE, fig.cap= "Tonnes of main species landed and haverst in aquaculture combined designated for human and non human consumption"}

processing <- read.csv(here("clean_databases/processing.csv"))

Aq <- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv")) %>% 
  filter(region != 'national') %>% 
  select(-subsector, -month) %>% 
  rename(chl_name = species) %>% 
  left_join(chl_to_sc) %>% 
  group_by(year, sc_name) %>% 
  summarise(tons=sum(tons, na.rm=TRUE))
  
Aq$tons[is.na(Aq$tons)] <- 0

# Landings artisanal
l.a <- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(region != 'national', subsector == 'artisanal') %>% 
  select(-international_waters, -month) %>% 
  rename(chl_name = species) %>% 
  left_join(chl_to_sc) %>% 
  select(-chl_name, -subsector) %>% 
  group_by(sc_name, year) %>% 
  summarise(tons= sum(tons, na.rm=TRUE))

l.a$tons[is.na(l.a$tons)] <- 0  

# Landings industrial

l.i <- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
    filter(region != 'national', subsector == 'industrial') %>% 
    select(-international_waters, -month, -subsector)%>% 
    rename(chl_name = species) %>% 
    left_join(chl_to_sc) %>% 
    select(-chl_name) %>% 
    group_by(sc_name, year) %>% 
    summarise(tons= sum(tons, na.rm = TRUE))

l.i$tons[is.na(l.i$tons)] <- 0

## Species matching english scientific
sc_to_en <- sp_index %>% 
  select(sc_name, en_name) %>% 
  unique()

## Most important species based on  contribution to HC with Aquaculture

sp_HC_with_Aq <- processing %>%
  filter(m_p == 'M') %>% 
  left_join(chl_to_sc) %>%
  group_by(year, sc_name, use) %>%
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  ungroup() %>% 
  left_join(sc_to_en)
  
top_5_HC_with_Aq <- sp_HC_with_Aq %>% 
  group_by(en_name, year) %>% 
  summarise(tons = sum(tons)) %>% 
  arrange(year, desc(tons)) %>%
  group_by(year) %>% 
  top_n(5, tons)
  
HC_with_Aq <- sp_HC_with_Aq  %>% 
  mutate(Species = ifelse(en_name %in% top_5_HC_with_Aq$en_name, en_name, "Others")) %>% 
  ggplot(aes(x= as.factor(year), y = tons/1000, fill = Species)) +
  #geom_area(position = position_stack(reverse = TRUE))+
  geom_bar(stat= 'identity') +
  labs(y="Thousand tonnes", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Royal1")))+
  facet_wrap(~use, scale = 'free')+
  ggtheme_plot()

```

We then looked into only wild-caught species only and identified that the species that have historically contributed more to human consumption products are: Chilean Jack Mackerel (*Trachurus murphyi*), Jumbo flying squid (*Dosidicus gigas*), and in a minor contribution Sea Urchins (*Loxechinus albus*) (HC_no_Aq). These results (Figure HC_with_Aq NHC and HC_no_Aq ) also shows that the amount of wild caught fiheries that go to non-human consumption is generally more than the double than what that goes to human consumption. It is necessary to remark that we identified that data per sector for 2015, prodived by SERNAPESCA and used for this analysis is not accurate. However we stil can see the trendes comparing data for years 2013, 2014, 2016 and 2017.


```{r, echo = FALSE, eval=F, warning=FALSE, message=FALSE, fig.cap= " Raw tonnes of main wild-cought species that are used for human consumption"}  

## Without Aquaculture (this assumed that all Aq goes to processing facilities). 

sp_HC_no_Aq <- processing %>%
  filter(m_p == 'M') %>% 
  left_join(chl_to_sc) %>%
  group_by(year, sc_name, use) %>%
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  left_join(Aq, by = c('sc_name', 'year')) %>% 
  rename(tons_raw = tons.x, tons_Aq = tons.y) %>% 
  ungroup()

sp_HC_no_Aq$tons_Aq[is.na(sp_HC_no_Aq$tons_Aq)] <- 0

sp_HC_no_Aq <- sp_HC_no_Aq %>%
  mutate(tons = tons_raw - tons_Aq) %>%
  left_join(sc_to_en) %>% 
  filter(use == "HC") %>% 
  filter(!sc_name %in% c("Mytilus chilensis", "Salmo salar", "Oncorhynchus mykiss", "Oncorhynchus kisutch")) 
  
top_5_HC_no_Aq <-  sp_HC_no_Aq %>% 
  group_by(en_name, year) %>% 
  summarise(tons = sum(tons)) %>% 
  arrange(year, desc(tons)) %>%
  group_by(year) %>% 
  top_n(5, tons)


HC_no_Aq <- sp_HC_no_Aq %>% 
  mutate(Species = ifelse(en_name %in% top_5_HC_no_Aq$en_name, en_name, "Others")) %>%
  ggplot(aes(x= as.factor(year), y = tons/1000, fill = Species)) +
  geom_bar(stat= 'identity') +
  labs(y="Thousand tonnes", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),
                               wes_palette("Royal1")))+
  ggtheme_plot()

```


In Chile, almost all industrial landings ($L_i$) and aquaculture harvest ($Aq$) go to processing facilities while only some of the artisanal landing ($L_a$) is processed. We assume that the the artisanals' catch that is not sent to processing facilities is sold as non-processed products for human consumption ($NP_{HC}$). These approach allowed us to analyze main species sold by artisanal fisher assuming it is all for human consumption. Results show that species vary across years, but the Southern rays bream (*Brama australis*) is the only specie that is common for every year of data (Figure NP_HC_plot). Indicating that historically has been an importat specie for direct human consumption. This aligns with results obtained in a survey implemented along Chile where 51% answered that Southern rays bream was their first preference (@fundacion_chile_chile_2016). Within 2013, 2014 and 2016, we can see the presence of anchoveta (*Engraulis ringens*), while in 2017 there is high increase in Jumbo flying squid (*Dosidicus gigas*). 


```{r Non-processed Human consumption,eval=F, echo= FALSE, message=FALSE, warning=FALSE, fig.cap= "Main artisanal landings that are not processe in facilities"}

## Raw material in processing plants with scientific name per region and year

processing_minus_ind_Aq <- processing %>% 
  filter(m_p== 'M', tons !=0) %>% 
  select(-use, -m_p, -product) %>% 
  left_join(chl_to_sc) %>% 
  select(-chl_name) %>% 
  group_by(year, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = TRUE)) %>% 
  left_join(l.i, by= c('year', 'sc_name')) %>% #How much of each sp comes from Li? 
  mutate(tons_raw = as.numeric(tons.x), tons_ind = as.numeric(tons.y)) %>% 
  select(-tons.x, -tons.y) %>% 
  left_join(Aq, by=c('sc_name', 'year')) %>% #How much of each sp comes from Aq?
  rename(tons_Aq = tons)

processing_minus_ind_Aq$tons_ind[is.na(processing_minus_ind_Aq$tons_ind)] <- 0
processing_minus_ind_Aq$tons_Aq[is.na(processing_minus_ind_Aq$tons_Aq)] <- 0

processing_minus_ind_Aq <- processing_minus_ind_Aq %>% 
  mutate(tons_art= tons_raw - tons_ind - tons_Aq) 

## Substracting artisanal catch that goes to faiclities to artisanal landings to get the rest (non-processed human consumption)

NP_HC <- l.a %>%
  left_join(processing_minus_ind_Aq, by= c('sc_name', 'year')) %>% 
  select(year, sc_name, tons_landed = tons, tons_processed= tons_art) %>% 
  mutate(tons_processed = ifelse(tons_processed < 0, 0, tons_processed)) %>% 
  mutate(tons_NP_HC = tons_landed - tons_processed) %>% 
  mutate(tons_NP_HC = ifelse(tons_NP_HC < 0, 0, tons_NP_HC)) %>%
  left_join(sc_to_en)

top_5_NP_HC <-NP_HC %>% 
  filter(year != 2015) %>%# data not accurate
  group_by(year, en_name) %>% 
  summarise(tons_NP_HC = sum(tons_NP_HC, na.rm = T)) %>% 
  ungroup() %>% 
  arrange(year, desc(tons_NP_HC)) %>%
  group_by(year) %>% 
  top_n(5, tons_NP_HC) %>% 
  filter(!is.na(tons_NP_HC)) %>% 
  select(year, tons_NP_HC, Species = en_name)
  
  
## Plot
NP_HC %>% 
  filter(!en_name %in% top_5_NP_HC$Species,
         year != 2015) %>% 
  group_by(year) %>% 
  summarise(tons_NP_HC = sum(tons_NP_HC, na.rm = T)) %>% 
  mutate(Species = "Others") %>% 
  bind_rows(top_5_NP_HC) %>%
  filter(!Species %in% c("Giant kelp", "Red algae gracilaria")) %>% 
  ggplot(aes(x= as.factor(year), y = tons_NP_HC/1000, fill = Species)) +
  geom_bar(stat= 'identity') +
  labs(y="Tons of non-processed HC (Thousand tons)", x="Year")+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"), wes_palette("Moonrise1")))+
  ggtheme_plot()
 

```



#### Main fisheries for food supply

To identify the main fisheries for human food supply in Chile we combined datasets from landings, aquaculture, raw material that goes into facilities for non-human consumption, imports and exports. Assuming that all local landings and harvest that is not processed as non-human consumption products nor exported is comsumed locally. Eventhough landings and aquaculture databases are detailed at a species level, we agragated the data in larger groups (See Suplementary Material) to be able to combine them with the Chilean Customs datasets. Our results show, that the main fishery for local consumption across the years has been squid. Within this group we considered: Patagonian squid (*Loligo gahi*), Jumbo flying squid (*Dosidicus gigas*), Argentine shortfin squid (*Illex argentinus*) and Japanese flying squid (*Todarodes Pacificus*). These species predominate local consumption from 2013 to 2017, however, we data shows that in 2017 there was a significant increase in the exports of this group of species, therefore explaining their relatively less contribution to domestic consumption. Other local fisheries important for domestic human consumption in Chile are Hakes (Southern hake *Merluccius australis*, South Pacific hake *Merluccius gayi gayi*, Patagonian grenadier *Macruronus magellanicus*, Southern blue whiting *Micromesistius australis*) and Southern rays bream (*Brama australis*) (Figure line 952). It is important to mention that the Southern rays bream is classified on its own group, therefore its relative contribution to domestic consumption is higher compared to other aggregated groups. Another caviet to make here is that squids have a the highest yield of all species, therfore there contribution to domestic consumption per landed tonnes is higher, partly explining its high predominance in our results. These resutls differe from the outcomes obtained with the international databases, where the model concludes that the main fish group destined for food in Chile is salmon. This difference can be explained due to the details in export data. In the model done with regional data, we grouped a larger groups of species under one group in order to match all databases which overall drived differences in our analysis. We also know that our regional analysis is sensitive to the yield, therefore results show that all salmon produced in Chile is exported. But even though most of the salmon is exported, according to SalmonChile [@asociacion_de_la_industria_de_salmon_de_chile_salmonchile_2018],  10-20% of the production stays in the country.

Our models show the chilean population also consumes high volumens of Tuna, however this is mainly imported in a canned format. This results aliged with the chilean population preferences, based on a study along the country, where 84% of the respondee aswered that their normal format for eating fish within a month was canned fish, follwed by freshed fish (80%) and frozen fish (37%) [@fundacion_chile_chile_2016]. The ten most importat fisheries according to our model are provided in the Appendix (Table XX).


```{r chl domestic consumption importa and wrangling databases, eval=F, echo = FALSE, message=FALSE, warning = FALSE}

## 1. functiosn to standarize names 

##complete specie name list with corresponding trade name
sp_index <- read.csv(here('clean_databases/sp_index.csv'))

# Chilean to trade name
chl_to_trade <- sp_index %>% 
  select(chl_name, trade_name ) %>% 
  unique()

# Imports to trade name 
imp_to_trade <- read.csv(here('clean_databases/imports_index.csv')) %>% 
  rename(imports_name = X...imports_name)

# Exports to trade name 
exp_to_trade <- read.csv(here('clean_databases/exports_index.csv'))

# english to trade
en_to_trade <- sp_index %>% 
  select(en_name, trade_name ) %>% 
  unique()

# scientific to trade
sc_to_trade <- sp_index %>% 
  select(sc_name, trade_name ) %>% 
  unique()

##2. Importing, cleaning, formating and grouping by trade name, databases necesary for the model 

# Landings + aquaculture form the annual reports of SERNAPESCA: 
L_an <- read.csv(here("clean_databases/landings_anuario.csv")) %>% 
  select(year, chl_name = species, sc_name, landing_aqua = tons) %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(landing_aqua = sum(landing_aqua, na.rm= TRUE))%>% 
  ungroup()

# Imports
I <- read.csv(here('clean_databases/imports_ADUANA.csv')) %>% 
  select(year, imports_name = name_sp, tons) %>% 
  filter(year != 2018) %>% 
  left_join(imp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(import = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Exports
E <- read.csv(here('clean_databases/exports_ADUANA.csv')) %>% 
  select(year, export_name = grupo_exportacion, tons) %>% 
  filter(year != 2018) %>% 
  left_join(exp_to_trade) %>% 
  filter(trade_name != 'NA') %>% 
  group_by(year, trade_name) %>% 
  summarise(export = sum(tons, na.rm= TRUE))%>% 
  ungroup()

# Yields
Y <- sp_index %>% 
  group_by(trade_name) %>% 
  summarise(mean_yield= mean(yield))%>% 
  ungroup()

# Non-himan consumption
NHC_raw <- read.csv(here("clean_databases/processing.csv")) %>%
  filter(m_p == "M",
         use== "NHC") %>% 
  group_by(year, chl_name) %>% 
  summarise(NHC = sum(tons, na.rm = T)) %>% 
  ungroup() %>% 
  left_join(chl_to_trade) %>% 
  group_by(year, trade_name) %>% 
  summarise(NHC = sum(NHC, na.rm= TRUE)) %>% 
  ungroup()

## 3. COmbaining all datasets in one dataframe
model_trade_name_Lan <- L_an %>% 
  ##full_join(A, by= c("year", "trade_name")) %>% 
  full_join(NHC_raw, by= c("year", "trade_name")) %>% 
  full_join(E, by= c("year", "trade_name")) %>% 
  full_join(I, by= c("year", "trade_name")) %>% 
  full_join(Y, by= "trade_name")

#convert all NAs to 0
model_trade_name_Lan$landing_aqua[is.na(model_trade_name_Lan$landing_aqua)] <- 0
model_trade_name_Lan$NHC[is.na(model_trade_name_Lan$NHC)] <- 0
model_trade_name_Lan$export[is.na(model_trade_name_Lan$export)] <- 0
model_trade_name_Lan$import[is.na(model_trade_name_Lan$import)] <- 0
model_trade_name_Lan$mean_yield[is.na(model_trade_name_Lan$mean_yield)] <- 0


## 4. Model
##Domestic consumption pero species
dom_cons_Lan <- model_trade_name_Lan %>% 
  mutate(HC = ((landing_aqua - NHC)*mean_yield) - export + import) %>%
  mutate(per_cap_kg = format((HC/17762681)*1000000, scientific = FALSE)) %>% 
  filter(trade_name != "Algae")

##Annul per-capita consumption
per_capita_total_Lan <- dom_cons_Lan %>%
  filter(HC>0) %>% 
  group_by(year) %>% 
  summarise(HC_year = sum(HC)) %>% 
  mutate(per_capita = (HC_year/17762681)*1000)

 model_results <- dom_cons_Lan
```



```{r top 5 fisheries for hc, eval=F, echo = FALSE, message=FALSE, warning = FALSE, fig.cap= 'Main  groups of species consumed by the chilean population in the last 5 years'}

##Top 5 fisheries for domestic HC
top_5_hc_cl <- model_results%>%
  filter(!is.na(year)) %>% 
  group_by(year, trade_name) %>%
  summarise(HC = sum(HC)) %>% 
  top_n(n = 5, wt = HC)
  
model_results %>% 
  filter(
    !is.na(year),
    !trade_name %in% top_5_hc_cl$trade_name,
    HC > 0) %>% 
  group_by(year) %>% 
  summarise(HC = sum(HC)) %>% 
  mutate(trade_name = "Others") %>% 
  bind_rows(top_5_hc_cl) %>%
  ggplot(.,
         aes(x = year, 
             y = HC/1000, 
             fill = trade_name)) + 
  geom_area(position = position_stack(reverse = TRUE))+
  # geom_bar(stat= "identity")+
  scale_y_continuous(expand = c(0, 0)) +
  labs(y= "Domestic human consumption \n (Thausands Tonnes)", x="Year") +
  scale_fill_manual(#trade_name = "Species",
                    values = c(wes_palette("Darjeeling1", 5, type = "discrete"),wes_palette("Darjeeling2", 3, type = "discrete"), "#000000")) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
  ggtheme_plot()

```


#### Total and per capita food consumption in the country
Our models is highly sensitive to species yield, therefore in many cases results of our calculations return a negative number implaying that exports are higher than the product of the yield times the local production (landings and aquaculture) minus all tons that go for non-human consumption products. Assuming that all negative numbers do not contribut to domestic consumption  because they are mainly exported or used for non human consumption (eg:algaes) and non is imported, we calculated a percapita consumption per year. We summed all the human consumption per year (Table XX, chunk below) and devided by the chilean population of 17,574,033 [@instituto_nacional_de_estadisticas_-_chile_sintesis_2018]. We alculated an average of 11.5 kg of seafood compsumption a year, which is similar to what has been estimated by the FAO (2016) for Chile of 10-13kg of fish a year.

```{r percapita year consumption table eval=F}
##Percapita consumption table

kable(per_capita_total_Lan, col.names = c('Year', 'Total Human Consumption (tonnes)', 'Per-capita consumption/year (kg)'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Total tones of seafood for humanc consumption and per-capita consumption per year (kg) considering a popilation of 17,762,681 individuals') %>% 
kable_styling(latex_options = 'hold_position') #%>% 
  #column_spec(4, width = "6cm")


```



#### Regional differences in per capita and total consumption
We where not able to do a regional analysis for domestic consuption. Even though most of the data is provided at a regional level. We realized that is was not possible to assume that what was landed in a region was processed and then exported in that same area. Nor that the imports per area accounted for the actual imported food consumed in the regions because there is internal transport within regions. On this regards, there is a study done in 2016, along the Chilean coast where people were asked how much fish they eat in a week. Regionla results say that 62% of the population from the north declares to eat fish once a week. In the north centre region that numbers decreases to 54% and it reaches its lowest point in the Metropolitan Region (42%). In the center south, 48% each fish at least once a week while in the southern most area 45% (@fundacion_chile_chile_2016)


### National Economic Participation

Fisheries, as an economic sector in Chile, considers activities of extraction and farming of marine and freshwater products. This sector contributes to around 0.4% of the national GDP although it is a highly variable sector (@central_cuentas_2017). An estimate of 2014, indicates that 90% of the sales of the sector came from farmed products (@sence_reporte_2015). What reflects the relevance of aquaculture in economic terms relative to wild fisheries. Yet, the extractive sector has a higher number of firms, 58% of the total (@sence_reporte_2015). Fishing manufacturing is closely relates to fisheries. Based on the last input-output matrix for the country's economy, for every 100 US dollars spent in the fisheries sector, another 52.2 dollars are generated in the manufacturing sector. Other sectors affected by fisheries are the financing sector (24.8 dollars generated every 100) and transportation and communication (13.4 dollars generated every 100). 

#### Wild-caught Fish Economic Participation

The contribution of fisheries to regional GDP suggests that the economic importance of the sector to regional economies varies widely. Ranging from almost 30% in the XI region to 0% in the metropolitan (RM) and IX region (Figure 23).  

```{r Fisheries_PIB_Chile_G, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6, fig.cap= 'Percent of the regional GDP contirbuted by the fisheries sector averaging data between 2013 and 2016. Data source: Banco Central de Chile.'}

PIB_region <- read.csv("Chile/PIB_region.csv") %>% 
  mutate(perc_contr = (PIB_fish * 100)/ PIB_total) %>% 
  group_by(region) %>% 
  summarise(mean_perc_contr = mean(perc_contr), sd= sd(perc_contr))

# Reordering regions
PIB_region$region <- factor(PIB_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

PIB_region %>% 
  filter(!is.na(mean_perc_contr)) %>% 
ggplot(., 
       aes(x= as.factor(region),
           y= mean_perc_contr)) +
  geom_bar(stat = 'identity', 
           fill = C_Col) +
  # coord_flip()+
  xlab('Region')+
  ylab('Contribution to GDP (%)')+
  scale_y_continuous(expand = c(0, 0), breaks = c(seq(0,40, by = 5)))+
  geom_errorbar(aes(ymin=mean_perc_contr-sd, ymax=mean_perc_contr+sd), width=.2,
                 position=position_dodge(.9)) +
  #geom_label(aes(label = paste(as.character(round(mean_perc_contr,1)))), position = )+
  ggtheme_plot() 

```

Based on our estimates of species-specific revenue the most important species for the artisanal sector are the Southern king crab (*Lithodes santolla*) and the Chilean sea urchin (*Loxechinus albus*) (Figure 23). Nonetheless, this varies across years. In 2017 the Patagonian toothfish (*Dissostichus eleginoides*) and the Southern rays bream or Corvina (*Brama australis*) were the species that provided more revenue followed by the Pink cusk-eel (*Genypterus blacodes*) in the artisanal sector.

Some of these species are among the most expensive (*Table App.6*). The Patagonian toothfish (*Dissostichus eleginoides*) has the highest mean ex-vessel price in the artisanal sector what explains in part why is one of the top 5 in term of revenues. We should notice that the fishing costs for different types of species could vary widely and we are not accounting for this in our analysis. Yet, we can qualitatively consider the costs of different types of fishing. For example, some of these species are fished by diving near the shore like the Chilean abalone (*Concholepas concholepas*) and the Chilean sea urchin (*Loxechinus albus*) while others like the Patagonian toothfish (*Dissostichus eleginoides*), the Southern king crab (*Lithodes santolla*) or the Swordfish (*Xiphias gladius*) involve offshore fishing, which tends to be much more expensive. The gathering of algae from the shore is arguably one of the less costly extractive activities for the artisanal sector but it is not among the ones that generate more revenue. We could not generate similar estimates for the industrial and aquaculture subsectors due to lack of data. 

```{r Revenues_artisanal_Chile_G, echo = FALSE, message= FALSE, warning= FALSE, fig.height=4, fig.width=8, fig.cap= 'Species that generated the most revenue for the artisanal sector over the past five years based on records on landings and ex-vessel prices. Source: records from SERNAPESCA, 2018.'}

# importing cleaned prices dataset and aggregating at the national level
prices_art <- read_csv("clean_databases/Chile/prices_art_SERNAPESCA.csv") %>% 
  filter(!(species == 'NA'))
prices_art[is.na(prices_art)] <- as.numeric(0) 
  
# Converting to 2018 dollars
prices_art <- prices_art %>% 
  mutate(price_dollar = ifelse(year == 2013, ((price/mean_uf_2013)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2014, ((price/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2015, ((price/mean_uf_2015)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ifelse(year == 2016, ((price/mean_uf_2016)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, ((price/mean_uf_2017)*mean_uf_2018)/from_ChP_to_dollar_dec_2018)))))
  
# Most important species artisanal based on national revenue
Top_5 <- prices_art %>% 
  mutate(revenue= (price_dollar * tons), chl_name = str_to_upper(species)) %>%
  left_join(sp_index) %>%
  group_by(sc_name, year) %>% 
  summarise(national_revenue = round(sum(revenue),2)) %>% 
  group_by(sc_name) %>% 
  summarise(total=sum(national_revenue)) %>% 
  arrange(desc(total)) %>% 
  slice(1:5)
  
prices_art %>% 
  mutate(revenue= (price_dollar * tons), chl_name = str_to_upper(species)) %>%
  left_join(sp_index) %>%
  group_by(sc_name, year) %>% 
  summarise(national_revenue = round(sum(revenue),2)) %>% 
  arrange(year, desc(national_revenue)) %>% 
  filter(sc_name %in% Top_5$sc_name) %>% 
  left_join(sp_index) %>%
  select(year, national_revenue, species= en_name) %>% 
  ggplot(aes(x = as.factor(year), y = national_revenue, fill= species)) + 
  geom_bar(stat = 'identity')+
  labs(y="Revenue\n(2018 US dollars)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual("Species",values = c(wes_palette("Darjeeling1")))+
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

```

#### Trade Economic Participation

Seafood products are one of the most important exports of the country. IFOP valued the exportation of fisheries-related products in $6,280 million dollars in 2017 (@ifop_boletines_2017). This corresponded to ~ 9 % of the total value of national exportation. The most important exported species are the Atlantic salmon (*Salmo salar*), the Rainbow trout (*Oncorhynchus mykiss*) and the Chinook salmon (*Oncorhynchus tshawytscha*)(*Figure 6*). All three produced via aquaculture. Mussels, specifically the Chilean Ribbed Mussel (*Aulacomya ater*) and the Chorus mussels (*Choromytilus chorus*) are also among the most valuable exports. These are landed by artisanal fishers and to a lesser extent produced in aquaculture centers. 

The most valuable sea products in terms of exports are frozen and fresh followed by fish meal and dried algae (@ifop_boletines_2017). In 2016, fish fillets were the fourth most exported product by Chile. This commodity generates around 2 billion US dollar per year and is mainly made of farmed salmon (@datachile_fish_2016). 

```{r Exports_Chile_G, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=8, fig.cap='Main exported species in terms of econmic value. Dataset Aduanas Chile'}
 
# Importing exports data and calculating FOB per year and species
exports_ADUANA <- read_csv("clean_databases/Chile/exports_ADUANA.csv")


exports_ADUANA%>% 
  filter(year != 2018) %>% 
  mutate(chl_name= str_to_upper(especie_SERNAPESCA))%>% 
  left_join(sp_index) %>% 
  group_by(year, sc_name) %>% 
  summarise(FOB = sum(value)/1000000) %>% 
  arrange(year, desc(FOB)) %>% 
  group_by(year) %>% 
  top_n(5, FOB) %>% 
  left_join(sp_index) %>% 
  select(year, species= en_name, FOB) %>% 
  ggplot(aes(x = year, y = FOB, fill = species))+
  geom_bar(stat = 'identity') +
  labs(y="FOB\n(Million US dollar)", x="Year")+ #JEPA_Chile, qué es FOB?
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual("Species",values = c(wes_palette("Darjeeling1"))) +
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

```


#### Fish and Aquaculture Employment

Fisheries and aquaculture are considered to be the activities that provide less direct employment in the country after agriculture (@sense_reporte_2015). Yet, the number of people employed in the sector has been on the rise since 2010 (@sence_reporte_2015). We estimated a total of ~ 278,000 job positions generated by fisheries and aquaculture. This estimate considers artisanal and industrial fishers, workers in processing facilities and aquaculture centers and indirect jobs generated by the manufacturing sector (*Table 2*). It does not include other indirect jobs generated by the extractive activities and its commercialization. Thus, the number of total people directly and indirectly employed by fisheries and aquaculture in Chile could be higher. 

```{r Employment_Chile_T, echo = FALSE, message= FALSE, warning= FALSE}

# Number of employees

Artisanal_extraction <- c('88,968', '2018', 'RPA (SERNAPESCA, 2018)')
Industrial_extraction <- c('3,525', '2018', 'RPI (SERNAPESCA, 2018) and SUBPESCA website (2018)')
Aquaculture <- c('17,631', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Manufacturing <- c('65,451', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Indirect <- c('102,758', '2018', 'Own estimate based on IFOP,2015 multipliers' )

table_employment <- rbind(Artisanal_extraction, Industrial_extraction, Aquaculture, Manufacturing, Indirect) %>% 
  data.frame() %>% 
  mutate(Stage_supply_chain = c('Artisanal Extraction', 'Industrial Extraction', 'Aquaculture Centers', 'Manufacturing', 'Indirect employment'), Employees= X1, Year_estimate= X2, Source= X3)
table_employment <- table_employment [4:7]

kable(table_employment, col.names = c('Stage of the supply chain', 'Employees', 'Year of the estimate', 'Source'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most recent estimates of people employed in different stages of the supply chain of sea products in Chile. Note that the estimate of indirect employment only considers indirect employment generated by the manufacturing stage.') %>% 
kable_styling(latex_options = 'hold_position') %>% 
  column_spec(4, width = "6cm") #%>% #wraps text in column2

```

People employed in the fisheries and aquaculture activities have an average monthly income of 579 US dollars, the lowest mean income in the country after the one generated by the agricultural sector (@sence_reporte_2015). This value ranges from 258 US dollars for people working independently and part-time to 1,160 US dollars for full-time employers (*Table App. 7*).

On average, people employed in the fisheries and aquaculture sector are a younger and less educated relative to the national average and are mostly men (*Table App. 8*). However, there is high variation within subsectors, for example, artisanal fishers tend to be older @tam_gone_2018. Based on estimates from the New national survey of employment (NENE, 2015) only 57.7% of the people employed in the sector is formally employed by an employer, the rest work independently (@sense_reporte_2015). 

```{r demographics employees, eval =F, echo = FALSE, warnings = FALSE, message= FALSE}
# Demographics of employees

Average_age <- c('42.1', '43.2')
Education_level <- c('9.5', '11.9')
Women_perc <- c('8.2','40')

table_demo_employment <- rbind(Average_age, Education_level, Women_perc) %>% 
  data.frame() %>% 
  mutate(Characteristic = c('Average age', 'Years of formal education', 'Percentage of women'), Fisheries_and_Aquaculture= X1, National= X2)
table_demo_employment <- table_demo_employment [3:5]


```
 
#### Employment in the Artisanal Sector

The artisanal sector corresponds to fishing activities performed by vessels equal to or smaller than 18 meters in length and by collectors in the inter-tidal zone (@subpesca_panorama_2018). Based on the current RPA, there are around 89,000 artisanal fishers along with 12,700 vessels in Chile. *Figure X-A*, shows the number of artisanal fishers per region and gender. They are concentrated in the X and VIII regions and are mostly men. The main activity in the artisanal sector is harvester or collector, this mostly corresponds to algae collection on the shore. All regions present similar composition of fishing activities (*Figure X-B*). However, the X region seems to rely more on benthic resources for employment than the others since they have a greater share of divers and algae collectors.

```{r Artisanal_distirbution_G, echo = FALSE, warning= FALSE, message= FALSE, fig.height=4, fig.width=9, fig.cap= 'Total number of fishers (A) and permits (B) in each region. One person can be registered in multiple activities. Dataset SERNAPESCA'}

# Importing national records of artisanal fisheres
rpa <- read_excel("Chile/RPA_RPI_mano_obra.xlsx", sheet = "RPA PESCADORES", skip = 2, n_max = 138294) %>% #JEPA_Chile, are we not giving them this database?
  rename(region = "Regi\u00f3n", year = "A\u00f1o Inscripci\u00f3n", gender = Genero)

art_reg_gender<- rpa %>% 
  group_by(`RPA Pescador`, region, gender) %>% 
  summarise(n = n()) %>%
  group_by(region, gender) %>% 
  summarise(count= n()) %>%
  mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV')))))))))))))))) %>% 
  select(gender, count, Region) 

# Reordering regions
art_reg_gender$Region <- factor(art_reg_gender$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

P1 <- ggplot(art_reg_gender, 
       aes(x= as.factor(Region),
           y= count/1000,  #thousand
           fill = gender)) +
  geom_bar(stat = 'identity') +
  # coord_flip()+
  labs(y="Number of people \n(Thousand)", x="", title = '')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual("Gender",values = wes_palette("Darjeeling1"))+
  ggtheme_plot()

# Number of boat owners, divers, fishers and gatherers in each region.

rpa_activity <- rpa %>% 
  select(region, year, ARMADOR, BUZO, RECOLECTOR, PESCADOR) %>% 
  gather(activity, practice, ARMADOR:PESCADOR) %>% 
  filter(practice == 'X') %>% 
  group_by(region, activity) %>% 
  summarize(count = n()) %>% 
  mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV'))))))))))))))), activity = ifelse(activity == 'ARMADOR', 'Boat owner', ifelse(activity == 'BUZO', 'Diver', ifelse(activity == 'RECOLECTOR', 'Collector', 'Fisher')))) 

rpa_activity$Region <- factor(rpa_activity$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

P2 <- rpa_activity %>%  
  ggplot(aes(as.factor(Region), count/1000,
             fill = activity)) +   
  geom_bar(stat = 'identity') +
  # coord_flip()+
  labs(y="Number of permits\n(Thousand)", x="", title= '')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual("Activity",values = wes_palette("Darjeeling2"))+
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

gt <- arrangeGrob(P1,
                  P2,
                  ncol = 2,
                  nrow = 1,
                  bottom = "Region"
                  )

as_ggplot(gt) + # transform to a ggplot
  draw_plot_label(
    label = c("B", "A"),
    size = 14,
    x = c(0.5, 0),
    y = c(1, 1)
    )

```

#### Employment in the Industrial Sector

The industrial sector comprises activities performed by vessels larger than 18 meters. Currently, there are 475 industrial vessels and 164 vessel owners according to national industrial records (RPI), from which only 25 are individual owners and the rest are firms (@maturana_mujeres_2017). Besides, 3,500 jobs are generated through vessels operations (@subpesca_pesca_2018). We could only find information on the distribution of vessels (*Figure 8*). We observe that most industrial  extractive activities occur in the VIII region, as suggested by landings estimates discussed above. Still, vessels are registered along the entire coast. Some of the firms that own industrial vessels also own processing facilities whose employees are considered as part of the manufacturing stage. 

```{r Distribution of industrial vessels, echo= FALSE, warning= FALSE, message=FALSE, fig.height=4, fig.width=6, fig.cap= 'Number of industrial vessels per region. Dataset SERNAPESCA.'}

# Importing national records of industrial fisheres
rpi <- read_excel("Chile/RPA_RPI_mano_obra.xlsx", sheet = "RPI", skip = 2, n_max = 473) %>% 
  rename(region = "Regi\u00f3n") %>% #JEPA_Chile, are we not giving them this dataset?
  group_by(region) %>% 
  summarise(count = n()) %>% 
    mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV')))))))))))))))) 

# Reordering regions
rpi$Region <- factor(rpi$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

ggplot(rpi, aes(x= as.factor(Region), y= count)) +
  geom_bar(stat = 'identity', fill = C_Col) +
  # coord_flip() +
  labs(y="Number of vessels\n(Thousand)", x="Region", title = '')+
  scale_y_continuous(expand = c(0, 0))+
  ggtheme_plot()

```

#### Employment in the Industry

A rigorous study by IFOP estimated a total of ~50,000 job positions in the manufacturing of marine products in processing facilities in 2014 (@ifop_monitoring_2015). However, a more recent report form SERNAPESCA reports a total of 65,451 jobs (@maturana_mujeres_2017). *Figure 9* displays the geographic distribution of people employed in the manufacturing sector per gender and type of product for 2015. Employment in the processing facilities is dominated by human consumption (HC) products, with ~ 40,000 job positions. Products for animal consumption (AC) generated around 5,000 positions while algae products employed less than 3,000 people. Most people employed by the manufacturing sector are located in the southern part of the country and the majority are men. 

```{r Employment in facilities, echo= FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=9, fig.cap= 'Number of employees in the fisheries manufacturing sector in each region by gender in 2015. AC refers to products for animal consumption and HC products for human consumption. Source: Report by IFOP based on facilities survey (IFOP, 2015)'}

# Importing IFOP, 2015 estimates of employees per type of facility

manufacturing_employment <- read.csv("Chile/employment_per_processing_type_area.csv") %>% 
  rename(M = man_employee, F = women_employee, Product = product, Zone = region) %>% 
  gather(key = Gender, value = Employees, 1:2)

# Reordering regions
manufacturing_employment$Zone <- factor(manufacturing_employment$Zone , levels = c("South", "Mid South","North" ))

ggplot(manufacturing_employment,
       aes(x= as.factor(Zone),
           y= Employees/1000,  #thousand
           fill = Gender)) +
  geom_bar(stat = 'identity') +
  # coord_flip()+
  labs(y="Employees (Thousand)",
       x="Zone"
       )+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  facet_wrap(~Product, scales = 'free')+
  ggtheme_plot()
  
```

#### Employment in the Aquaculture Sector

There are 17,631 people employed among the 3,683 aquaculture centers currently registered in Chile (@maturana_mujeres_2017). From these, ~4,500 are women. *Figure 10* shows the distribution of aquaculture centers per region and type of farmed species. The main type of product being farmed is fish, mostly salmon, followed by mollusks and algae. The X region hosts more than half of the aquaculture centers in the country. 

```{r Aquaculture centers per region and type of species, echo= FALSE, message= FALSE, fig.height=4,fig.width=8, warning= FALSE, fig.cap= 'Number of aquaculture centers in each region by type of species in 2017. Source: Report Subsector Acuicultura 2017 by SERNAPESCA (SERNAPESCA website, 2017)'}

# Importing number of aquac. centers, SERNAPESCA

centers_aquaculture <- read.csv("Chile/centers_per_species_aquaculture.csv") %>% 
  # rename(Region = ï..Region) %>% 
  gather(key = Species, value = 'Centers', 2:7)

# Reordering regions
centers_aquaculture$Region <- factor(centers_aquaculture$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

ggplot(centers_aquaculture, aes(x= as.factor(Region), y= Centers, fill = Species)) +
  geom_bar(stat = 'identity') +
  # coord_flip()+
  labs(y="Number of Centers", x="Regions")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  #facet_wrap(~Species, scales = 'free', ncol = 2)+
  ggtheme_plot() +
  guides(fill = guide_legend(nrow = 1, byrow = TRUE))

```


** Species-specific estimates of economic contribution **

\clearpage


## Critical Analysis

### Data Uncertainty and Gaps

Most of our analysis relied on publically available data collected by governmental institutions. Although national efforts towards transparency and better data collection have increased over the years, there are still sources of uncertainty and error. Clear evidence of the presence of error in our analysis are the negative estimates of domestic consumption we got for some species after substracting exports to all what was produced in the country. Negative estimates are not logic since a country cannot export more than what it has produced. We think that these estimates can be mainly explain by the presence of Illegal unreported and unregulated fishing (IUU), non-perfect matching between databases, and lack of information on species-specific yields and ingredients composition of marine products. 

Particularly concerning is landing data since it is mainly generated via landings declarations. Fishers should announce their intention to land catch to the authority in charge, who must weight and certify the catch. Commercialization centers should not sell any kind of catch that is not certified. Thus, fishers have an incentive to report their landings. In addition, those fishers that do not declare their landings risk their permits and penalty fees. Still, these rules are poorly enforced generating incentives to underreport. IUU fishing occurs in Chile. In 2017, the five most seized species due to non-compliance with fishing regulations were the Araucanian herring (*Strangomera bentincki*), the Peruvian anchoveta (*Engraulis ringens*), the Giant grey kelp (*Lessonia nigrescens*), the Jumbo Flying Squid (*Dosidicus gigas*), and the South pacific hake (*Merluccius gayi*). All of them are among the most landed species in the country. There has been recent efforts to generate estimates for IUU for some Chilean fisheries like Loco (*Concholepas concholepas*) (@oyanedel_illegal_2018) and the South pacific hake (@wwfchile_estimacion_2017). However, we did not include them in our analysis because they are not available for all species and including them only for the species that have been studied could have bias our results in a systematic way. IUU is not recorded in the landings database but products coming from IUU that are exported are likely to be recorded in the exports database. This could be making us to underestimate domestic consumption. 

The lack of consistency in data collection between and within agencies brings another source of uncertainty to our analysis. We found inconsistencies between two databases of landings published by SERNAPESCA in their website. They differed in the level at which data was aggregated. One data base was clasified per secto (artisanal, industria, fabric vessels) and the other was the summary for lanfings and aquaculture for each year pero region. When we aggregated the most detailed database to the same level as the anual summary database we observed differences in landing volumes for the year 2015. We considered this to be part of the measurement error involved in any process of data collection or data entry. We used the summary annual data dor our domestic human consumption model because it alinged more with amounts reported all other years.

A major challenge was to match trade and landings data. SERNAPESCA collects data under the common names while Aduanas Chile record exported and imported products under the international Harmonized System nomenclature. To overcome this issue we use the information provided by Aduanas Chile on the species contained in some types of products. This allowed us to aggregate products per groups of species and match volumes and values to landing record by SERNAPESCA. In addition, SERNAPESCA applies several variations of a species common name between databases. This difficults the calculation of aggregated estimates. We solve this by generating a dataset that categorize all the names currently used by SERNAPESCA under a common group among the import and export data (See appendix for group details). We designated the groups and did all the corresponfing matching, this step is prone to human error.

It is possible that we have overestimated exports volumes for some species since some products combine the species of interest with other ingredients. We could not account for this since we did not have data on the products composition on different ingredients. However, for the most exported products this is not a major issue since fish fillets and fresh products are usually almost 100% made by fish meat or shellfish. Similarly, we lacked data on species-specific yields to estimate how many Kg of a species translates into different marine products. To fill this gap, we use information from Peru on production yields and apply it to Chilean species. Some of them were excatly the same species but for those that did not present a match in Peru we use the yield of the most similar species. We found our analysis to be particularly sensitive to the values of species-specific yields, highlighting the importance of improving this aspect of the model.

A major information gap we identified is the lack of publically available data on ex-vessel prices and employment for the industrial sector. To our knowledge neither SERNAPESCA, IFOP nor SUBPESCA collect data on this regard. Thus, identifying what were the most valuable species for the industrial sector was not possible. An approach we were unable to apply due to time constraints is to derive ex-vessel prices from trade data (@melnychuk_reconstruction_2017). Contrary to what we expected, the fisheries sector for which we found most available information was the artisanal sector. We were able to get a count of the direct employment generated by the artisanal sector thanks to the National Record of Artisanal Fishers (RPA). Still, we cannot ensure we are not underestimating the number of people that rely on fishing. We know enforcement is weak specially in remote areas of the country therefore there is room for illigal activity. We were also able to estimate species-specific revenues by using landings and prices of first transactions. Finally, information on employees in processing facilities was also available. However, some values were unrealistic since they were larger than the entire population of the country. Thus, we prefered relied on a survey study performed by IFOP to fullfill this estimate.   

We also lacked data on the transportation of landings and products between Chilean regions and final destination. This prevented us from applying our model of food supply at the regional level. The problem is that the catch landed in one region could be processed in a different one and then the final product sold in a third region. Running the regional analysis without accounting for this issue would have led to incorrect estimates of food consumption. 

### Fisheries context

-  The most important fisheries in Chile in terms of landings are Anchovy (*Engraulis ringens*), the Araucanian herring (*Strangomera bentincki*), which are fished by both the artisanal and industrial sectors, and the Chilean jack mackerel (*Trachurus murphyi*), fished mostly by artisanals. 

### Seafood Consumption

**Key messages**:

- Most importants species in terms of food consumption in Chile are Jumbo Squid (*Dosidicus gigas*) and Southern rays bream (*Brama australis*) followed by hakes and mackerel. 

- Trends in food consumption show that in general the species that contrubut to domestic food consumption has been constant through time exept for Hake, that has been constanlly decreasing.

- Tuna is one of the main groups of species consumed in the country but is mainly imported.

- Most of wild-caught fish is processed for non-human consumption products.

- Despite its massive production, aquaculutre is not a major source of food supply for the Chilean population. Most of its production is exported or are too expenssive to be accessed by lower income groups which is the population of interest in the contxt of food security.


Chile is considered one of the main countries in fishery production worldwide [@fao_estado_2016], however several reports and studies suggest the Chilean population does not consume as many marine products [@gonzales_solo_2018]. We found that on average between 2013 to 2017 the chilean population eats ~11kg of seafood a year. This is lower than the 20 kg a year recommended by Food and Agriculture organization (2016). According to the National Health Survey [@ministerio_de_salud_gobierno_de_chile_encuesta_2018], only 9.2% of the Chilean population eats fish at least two times a week as recommended. This value has decreased since the early 2000s where this percentage was 10.7% [@ministerio_de_salud_gobierno_de_chile_encuesta_2018]. World Data [@ritchie_our_nodate] also shows a decrease in the amount of percapita seafood consumed per year by the Chilean population from ~16kg consumed in 2004 to 12.5kg in 2013.

Looking into what type of seafood the Chilean population consumes, a study done in 2012 shows that 60% of the seafood consume are pelagic fishes, 12% are mollusk and 2% are crustaceans [@villena_marcelo_diagnostico_2012]. With our model we also identifyed that Chileans moslty rely on pelagic fisheries as their primary source for seafood. The most important groups of species for domestic food consumption are squids, tuna, hake, and the Southern ray beam. However, it is likely that we have overestimated squid consumption. It is worth noting that Tuna is mostly imported into the country while most of the species that are use in Chile for Human Consumption products like salmon, trouts and mussels are exported. This reflects the importance of market forces on world-wide seafood distirbution and supply. Salmon and trouts get better prices a broad and are too expensive for Chileans in general. 

Seafood consumption in Chile has been shown to be correlated with income, as in other parts of the world. However, the data is spotty and is provably missing some of the consumption that occurs in coastal communities. According to a study done in 2016 by Fundación Chile and Adimark [@fundacion_chile_chile_2016], the frequency of seafood consumption varies along socio-economic groups in Chile, 55% of the high-income group consumes seafood at least once a week compared to the 38% of the lower-income group. It has also been identified [@ministerio_de_salud_gobierno_de_chile_encuesta_2018] that people with less years of schooling tend to eat less fish (only 6.3 percent eats seafood twice a week) than people that have completed high school education, where 11.2% eats as frequently. 

Within the Chilean population there is a strong belief that seafood is expensive. However, there is wide range of fish prices. There are expensive products such as salmon and affordable alternatives such as Chilean Jack Mackerel (*Trachurus murphyi*), Jumbo Flying Squid (*Dosidicus gigas*) and Southern Rays Bream (*Brama australis*) [@ministerio_de_salud_gobierno_de_chile_encuesta_2018].


### Economic participation

Aquaculuture is the most important activity in the fisheries sector in terms of economic value. Their profits come mainly from the exportation of salmon products. This positions salmons as the group of marine species that contribute the most to the country's economy. The Chilean jack mackerel (*Trachurus murphyi*) and the Jumbo squid (*Dosidicus gigas*) are the wild-caught species that have contributed the most in exports over the pasts years. The latter is use for DHC and is targeted by both the artisanal an industrial sector. Yet, current discussion on closing access to the industrials to this fishery could bring consequences for both income and food supply. 

Among wild-caught fisheries the ones that seems to generate more revenues for the artisanal sector are the Southern king crab (*Lithodes santolla*) and the Chilean sea urchin (*Loxechinus albus*). While the first fishery occurs only in the South of the country, the latter is ditributed along the coast of Chile.

Although the direct employment in aquaculture centers is low relative to the number of artisanal fishers, it is very likely that an important share of the job positions in the manufacturing sector are being supported by aquaculture. Especially in the South where most job positions for manufacturing of human consumption products are concetrated. The role of Algaes in terms of employment for the artisanal sector is also worth notice. The revenues of most these species are not particularly high, however the lower costs of extraction relative to other types of fishing make algae collection an attractive form of livelihood for coastal communities. 

\clearpage

## Oceana Next Steps 

We learned that there is poor undertanding on Chile regarding the main marine species in terms of food security and economic participation. Although people have tried to answered these kind of questions before there is not a constant track of what species are being eaten and how much. We think that an approach like the one presented here could be easily automatized to keep track of the trends on domestic seafood consumption. Nonetheless, it would requiere some improvements to get more reliable estimates by full-filling some of the datagaps. 

A mayor challenge we faced was to combine all datasets without losing species resolution. We think there is potential to imporve the methods we used for scraping data from the imports and exports databases as for example, the Chilean customs recently uploaded the export information for 2018 as an excel file combining type of products, aduana code, exported tones and values. When we ask for this sort of document back in October, the Chilean customes told us they were working on it so hopefully the publish historic data in that fomat. This could increase the species resolution, reduce human error and facilitate automatization. Othe two things that would improve the estimates we obtained here are species-specific IUU estimates and yields per type of product for Chilean species. IUU estimates are becoming more popular and different methods have been developed for estimate it from elliciation tasks to more model driven approaches. This kinds of estimate could be also of interested to management authorities. Thus, Oceana could cooperate with them to ellaborate them. 

Chile is increasingly pushing for an increase in seafood consumption. Currently, there are programs that promote seafood consumption like _Del Mar a mia Mesa_ (From the sea to my table). We think Oceana could contribute adding to the discussion the whole picture of the state of the fisheries and the economic implication for the country of transitioning to a more fish based diet. For this an integrated analysis about the state of the fishery and its economic importance for the country needs to be done. More than just promoting the increase in domestic seafood consumption we think that finding out why is it important and who will benefit and loose it is key for the success of any project. 

In our analysis we did not explore the role of the commercialization chain on food supply and distribution. Investing in better understanding from where can people access seafood and at what prices is crucial to improve food security. Thus, there is a need to better characterize the distirbution of seafood in Chile and explore the possibility of improving seafood access by better managing the supply chain. 

Finally, it would be interesting to study the preferences an attitudes of Chileans towards products that have not been historically consumed but that are becoming important in terms of landings like algae and squid. Understanding the potential for this species to succed as food products within the country may be important to improve access to marine food at lower prices. 


# Conclusions

The main takeaway of our analysis regarding the contribution of domestic fisheries for human consumption is that most of the wild-caught fish currently does not go to human consumption, but to produce no human consumption products like fish meal and oil or algae products. On the other hand, most of the aquiculture produced in the country is for human consumption but it is exported elsewhere. Only XX% of the local production stays in the country for human consumption. Additionally, one of the most relevant species for human consumption in Chile is tuna, which mainly in imported and canned.

In economic terms, aquaculture is more important than wild-caught fisheries producing ~ 90% of the sales of the economic sector. Nontheless, wild fisheries is more important in terms of employment if we consider both the artisanal and industrial sector. 



## Chile

# References

PRODUCE. (2010). Manual de administración de las infraestructuras pesqueras artesanales. Ministerio de la Producción, Viceministerio de Pesquería, Dirección General de Pesca Artesanal. Lima, Perú. 121p.



# Supplemental Material List

## Chile
### Methods


Any conversion from Chilean pesos to dollars was done by first converting the amount in Chilean pesos to its value in Chilean Units of account (UF, a monetary unit corrected by inflation) of the respective year. Then, transformed to the equivalent in 2018 Chilean pesos using the 2018 UF value and finally, converting the 2018 value in Chilean pesos to its equivalent in 2018 US dollars, based on the conversion for Dec, 2018 (1 US dollars corresponds to 676 Chilean pesos).

### Results

```{r Table 1 Appendix main species per region and year, eval=F,echo=F}

# Main species per region artisanal sector overlast 5 years (Appendix)
# to avoid double obsv

 main_species_reg_art <- landings %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  select(- month, - international_waters, chl_name = species) %>% 
  left_join(sp_index) %>% 
  select(subsector, region, year, tons, en_name)
  group_by(year, region, en_name, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>%
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
 
main_species_reg_art$region <- factor(main_species_reg_art$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))
 
# Main species per region industrial sector overlast 5 years (Appendix)

 main_species_reg_ind <- landings %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  filter(region != "national", international_waters == 0, subsector == 'industrial') %>%
  rename(chl_name = species) %>% 
  left_join(sp_index) %>%
  group_by(year, region, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  filter(tons !=0 ) %>% 
  left_join(sc_name_to_en) %>% 
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
  
main_species_reg_ind$region <- factor(main_species_reg_ind$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

   
 landings_per_region
 
 
 
```


```{r App1, eval=F}

# Most important farmed species per region (Appendix)
  
  aquaculture_species_per_region <-aquaculture %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons), chl_name = species) %>%
  left_join(sp_index) %>% 
  filter(tons != 'NA') %>% 
  group_by(year, sc_name, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>%
  left_join(sc_name_to_en) %>% 
  select(-tons, -sc_name) %>%
  unique() %>% 
  spread(year, en_name)
  
aquaculture_species_per_region$region <- factor(aquaculture_species_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))
```


```{r App6, eval=F}

# Most expensive species artisanal (Appendix)
most_expensive <- prices_art %>%
  mutate(chl_name = str_to_upper(species)) %>% 
  left_join(sp_index) %>% 
  group_by(sc_name) %>% 
  summarise(mean_price = round(mean(price_dollar),2), sd= round(sd(price_dollar), 2))%>%
  arrange(desc(mean_price)) %>% 
  top_n(10, mean_price) %>% 
  left_join(sp_index) %>% 
  select(species = en_name, mean_price, sd) %>% 
  unique()
```
  
```{r App7, eval =F}
  
```  

```{r transforming average income from chilean pesos 2014 to dollars 2018, eval=F, echo =FALSE}

mean_income_sector_chp <- 345541 # Based on report from SENSE 2015
mean_income_sector_dollar_2018 <- ((mean_income_sector_chp/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018

```

```{r data on incomes, eval=F, echo = FALSE, message=FALSE, warning= FALSE}

full_time <- c(692786, 210046, 528532, 227808, 402353)
part_time <-c(422099, 154040, 277600, 172445, 180434)

table_incomes_sector <- rbind(full_time, part_time) %>% 
  data.frame() %>% 
  mutate(Type = c('Full time', 'Part time'), Employer= round(((X1/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), Independent= round(((X2/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), Dependent_Contract= round( ((X3/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018, 2), Dependent_no_contract= round( ((X4/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018,2), mean_income_fisheries= round( ((X5/mean_uf_2014)*mean_uf_2018)/from_ChP_to_dollar_dec_2018),2)

table_incomes_sector <- table_incomes_sector[6:11]

```
  
```{r Appendix, eval=F, echo = FALSE, message= FALSE, warning= FALSE}

# Top 10 landed species artisanal (for each year between 2013-2017)
knitr::kable(top_10_art, col.names = c('Ranking', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Top 10 landed species per year in the artisanal sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Top 10 landed species industrial (for each year between 2013-2017)
knitr::kable(top_10_ind, col.names = c('Ranking', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Top 10 landed species per year in the industrial sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most landed species per region artisanal (for each year between 2013-2017)
knitr::kable(main_species_reg_art, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per and year region in the artisanal sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most landed species per region industrial (for each year between 2013-2017)
knitr::kable(main_species_reg_ind, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per region and year in the industrial sector. Source: Landing records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Most harvested species per region (for each year between 2012-2017)
knitr::kable(aquaculture_species_per_region, col.names = c('Region', '2012', '2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most harvested species per region and year. When no data is available means there were no aquaculure that year in that region. Source: Harvest records from SERNAPESCA, 2018') %>% 
kable_styling(latex_options = 'hold_position')

# Appendix table top 10 species human consumption
 top_10_hc <- model_results%>%
   arrange(year, desc(HC)) %>%
   group_by(year) %>%
   top_n(10, HC)
 
 top_10_hc_table <- top_10_hc %>% 
   select(year, trade_name, HC) %>% 
   filter(!is.na(year)) %>% 
   group_by(year) %>%
   mutate(rank = rank(desc(HC))) %>%
   select(-HC) %>% 
   spread(key= year, value =trade_name)
 
 knitr::kable(top_10_hc_table, col.names = c('2013', '2014', '2015', '2016', '2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Main groups of species for human consumption') %>% 
 kable_styling(latex_options = 'hold_position')


# Most expensive species in the artisanal sector
knitr::kable(most_expensive, col.names = c('Species', 'Mean price (US$/Kg)', 'SD Price'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most expensive species in Chile based on their mean ex-vessel price between 2013 and 2017. Source: Data on ex-vessel prices for artisanal sector, SERNAPESCA, 2018.') %>% 
  kable_styling(latex_options = 'hold_position')

# Mean income for different times of employees in the fisheries and aquaculture sectors
knitr::kable(table_incomes_sector , col.names = c('Type', 'Employer', 'Independent', 'Dependent with contract', 'Dependent without contract', 'Mean income'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Monthly incomes of fisheries sector in 2018 US dollars. Source: Elaborated based on data from New national employment survey (NENE) in 2015 and published in SENSE, 2015') %>% 
  kable_styling(latex_options = 'hold_position')

# Demographics of people employed in the fisheries and aquculuture sector
knitr::kable(table_demo_employment, format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Average demographic characteristics of people employed in the fisheries and aquaculture sector in Chile. Source: Elaborated based on data from New supplementary income survey (NESI) in 2015 and published in SENSE, 2015') %>% 
kable_styling(latex_options = 'hold_position')

```
  
