---
output: 
  pdf_document: 
    fig_caption: yes
    includes:  
title: "Contributions of marine capture fisheries to the domestic livelihoods and seafood consumption of Chile"
subtitle: "Final Draft"
author: "Camila Vargas e Ignacia Rivera"
bibliography: main-fisheries-chile.bib
csl: cell.csl
---
\newpage

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(here)
library(magrittr)
library(tidyr)
library(reshape2)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(readxl)
library(cowplot)
library(kableExtra)
library(readr)
library(stringr)
library(purrr)
library(naniar)
library(wesanderson)

ggtheme_plot <- function() {
  theme(
    plot.title = element_text(size = rel(1), hjust = 0, face = "bold"),
    panel.background = element_blank(),
    strip.background = element_blank(),
    # strip.text       = element_text(size = base_size, face = "italic"),
    panel.border     = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 10,
                               angle = 0,
                               face = "plain"),
    axis.text.y = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.key = element_rect(colour = NA, fill = NA),
    legend.position  = "top",
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text.x = element_text(size = 18, colour = "darkgrey")
  )
}
```

## Methods

### Data sources

1. The Chilean National Service of Fisheries (SERNAPESCA) is the main entity in charge of keeping track of landings and other fisheries and aquaculture related data in Chile. Most of their data is available on their website. We accessed their databases on landings from wild-caught and farmed fisheries, first transaction prices for artisanal fisheries, and landings used in different types of sea products. These databases provided information at the species and regional level. We also downloaded their data on the geographical distribution of production facilities and the types of sea products they produce. 

We issued data requests on SERNAPESCA's portal to obtain more detailed information. Via this method, they provided us the national record of artisanal fishers (RPA), which lists current registered fishers and boats, and the national record of industrial vessels (RPI), which lists current registered industrial vessels. The RPA informs the region, gender and age for each artisanal fisher as well as the species-specific permits they hold. Together the RPA and RPI provide information of artisanal and industrial vessels like lenght and storing capacity. SERNAPESCA also provided to us the number of employees of each gender per processing facilities as well as species-specific landings per vessel, and artisanal fisher. 

2. The Chilean Costumes (ADUANAS Chile) is the main entity in charge of keeping records of imported and exported goods in the country. We accessed their exports and imports data per product and region via their website.

3. The Fisheries development institute (IFOP) is a private entity that advices fisheries management based on permanent biological, environmental and economic monitoring, and scientific reasoning. Although most of the data collected by IFOP pertains to stock assessments and biological indicators, they also generate economically relevant statistics and landing estimations. We reviewed their most recent annual report on the economic monitoring of Chilean fisheries for 2015. It includes databases and estimates for employment and the economic value of different fisheries. 

4. The Chilean Central Bank (Banco Central de Chile) is the entity in charge of estabilizing the Chilean currency. To fullfill its function it monitors the main economic indicators per economic sector. We accessed their data on sectorial GDP for each region per industry. 

5. DataChile (www.datachile.com) is an online plataform that integrates, visualizes and distributes public available data from Chile. We accessed some of their databases for convience since they are already cleaned and organize for analyses. We downloaded from their website data from the New National Employment Surevy (NENE) and data from the Suplementary National Survey of Income (NESI). We also accessed their datasets on export and import records (originally from ADUANAS Chile), which where organized at the municipality level.

6. The Undersecretary of Fisheries (SUBPESCA) takes cares of designing and imlpementing fisheries and aquaculture regulations. They generate monthly reports on changes in landings for the most important species and marine products exportations. The data in these reports comes from SERNAPESCA and ADUANAS. Thus, extracting data from these reports would be redundat but we use them to reference particular information of interest. 

7. The Research Office of Agrarian Policies (ODEPA) as well as scientific publications. 

### Specific methodologies

#### Fisheries context

We used records of landings and aquaculture farming from SERNAPESCA to evaluate the main species in terms of volume over the last five years (2013-2017). <!--- We compare these results with information from different sources -->  

#### Estimating National Seafood Supply

We combined data from ADUANAS Chile and SERNAPESCA to generate two main databases on the number of tons that are catches, harvested, exported and imported for all marine species in Chile. 

The specific data-sets we used are:

- Catches: annual tonnes per marine species landed in each region (SERNAPESCA, 2013-2017).
- Species contribution to different types of products: annual tons per marine species used in each region to elaborate each type of product (i.e. fish meal, oil, canned products, freeze products, dried products, smoked products, agar, col-agar, carragenina) (SERNAPESCA, 2012-2016).
- Aquaculture: annual tonnes per species harvest from aquaculture centers in each region (SERNAPESCA, 2012-2017).
- Exports: quantities of all the products exported from each Chilean region to other countries. (ADUANAS, 2013-2018).
- Imports: quantities of all marine products imported into the country. (ADUANAS, 2013-2018)

 
Each product in the original exports and imports data-sets is identified through an id number that matches the name of the product in a code known as the _Arancel Aduanero_. This document was only available to us in .pdf format. Thus, to identified the id of products containing domestically fished and harvest marine species we searched for the names of the fished and harvested species listed by SERNAPESCA in the document with the _Arancel Aduanero_. We searched for all the ids that refer to the species scientific name, common name or genus name as listed by SERNAPESCA and aggregated them under the species name.

We included a column that indicates the number of tonnes of each species that were used in products that are not directly ingested by humans. For this, we assumed that the following categories of products are not for direct human consumption: fish meal, col-agar, agar, carragenina, oil and alginato. Finally, to aggregate all the data at comparable levels we generated broad categories of species based on the imports-data set, which is the one that has less detail at the species level.

Because imports-data is not available at the regional level we generate two databases, (i) one describing tons landed, harvested, exported and imported at the national level per groups of species and (ii) one containing tons landed, harvested and exported at the regional level per species.

We used the R software to clean and format all the available information in long-form databases. All the scripts for replicating our work are available in "scripts/cleaning_data".

We determine the domestic human consumption of marine products through this general model: (Catch - Products not directly for human consumption) + Aquaculture + Imports - Exports. We assumed that all the products that are no exported or are not used in other elaboration process are consumed by the local population. For this report we explore database i mention above, because it has import quantities.

#### Estimating National Economic Participation


** General estimates of fisheries economic contribution ** 

** Income **
We use data from Banco Central de Chile to reconstruct the percent of the regional GDP that comes from fisheries and calculate the mean percent between 2013 and 2016. 

To estimate most important fisheries in terms of economic value, we used data on first-transaction prices per species per month and location (requested to SERNAPESCA) and computed the national average price per year. We used two different approaches, one considers the total revenue generated by different fisheries while the other considers how valuable they are based on their price. To estimate total revenue we multiplied price by the landed tons per year. We converted Chilean pesos to 2018 dollars. Data on price at first transaction at the level of species was only available for the artisanal sector since data on industrial transactions is not recorded officially.



** Employment **

We collect information from several sources to reconstruct employment estimates for different stages in the supply chain of marine products. IFOP, 2015 found that 1.57 indirect employments are generated by each direct employment in the manufacturing fishing sector. We apply this multiplier to our most recent estimate ofnumber of people employed in the manufacturing fishing center to find an estimate of indirect employment. We then qualitatively related the geographic distribution of employees in different stages of the supply chain with the most important fisheries in a given region to find the most important specie sin terms of employment contribution.  

We also obtained data from SERNAPESCA detailing the number of people working in each processing facility. However, some of the values were clearly wrong (because they exceeded the entire Chilean population) making it impossible to use this information to generate reliable estimates. Thus, we relied of information from (@ifop_monitoreo_2015) to study the manufacturing sector. 

## Results

#### Fisheries context

** Main wild-caught fisheries in terms of volume**

According to landing records from SERNAPESCA over the last five years, the artisanal sector lands on average 1.6 ± 0.58 million tons per year of marine species while the industrial sector lands around 0.9 ± 0.16 million tons per year. The most landed species by the artisanal sector have been the Araucanian herring (*Strangomera bentincki*), the Giant grey weed (*Lessonia nigrescens*) and the Peruvian anchoveta (*Engraulis ringens*) and the most landed species by the industrial sector are the Peruvian anchoveta (*Engraulis ringens*), the Chilean jack mackerel (*Trachurus murphyi*) and the Araucanian herring (*Strangomera bentincki*) (**Figure 1**).

Over the last years the catch composition of the artisanal sector has presented no major changes except from one species of algae replacing other within the top 5 landed species. In the case of the industrial sector, composition of the main three species seem to be stable over the last five years. Yet, the South pacific hake (*Merlucius gayi*) stoped being one of the specied most fished by the industrial sector after 2013. This can be explined by the reduction of the total allowable catch for this species given its risk of collapse in those years (@subpesca_cuota_2013). After 2013, the industrial sector introduced the Jumbo squid (*Dosidicus gigas*) among its top 5 landed species. As for 2016 the Chub mackerel (*Scomber japonicus*) has entered the top 5 for the indutrial sector replacing Patagonian grenadier (*Macruronus magellanicus*).


```{r Most landed species in Chile, echo = FALSE, fig.cap= 'Top 5 species landed in Chile by the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. Data source: Landing records from SERNAPESCA, 2018.' }

# Importing landings farmed and species names 

landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv"))
species_list <- read.csv(here("clean_databases/intermediate/sp_list.csv")) %>% 
  rename(species= especie_SERNAPESCA, sc_name=nombre_cientifico_SERNAPESCA) %>% 
  mutate( species = str_to_upper(species)) 

# Calculating annual landings per sector 

annual_landings_sector <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T)) %>% 
  group_by(subsector) %>% 
  summarise(annual_tons = mean(tons), sd = sd(tons) )

# Top 5 landed per sector

 landings_top_5 <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, species, subsector) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, subsector) %>% 
  top_n(5, tons) %>% 
  left_join(species_list) %>% 
  select(year, subsector, tons, species = sc_name) %>% 
  ggplot(aes(x = year, y = tons, fill = species))+
  geom_bar(stat = 'identity') +
  facet_wrap(~subsector)+
  labs(y="Landings (Tons)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2"),  'black'))+
  ggtheme_plot()
 
 landings_top_5
```

 ** Regional Diferences in wild-caught fisheries **

The amount of landed tons (Figure X) and the main species landed vary widely along the coast of Chile for both sectors. The most landed species in each region has change little over the last five years (Tables SX). 

In the north of Chile, from the XV to the III region, the most landed species of the artisanal sector over the last five years have been the Peruvian anchoveta (*Engraulis ringens*) and the algae Giant grey weed (**Lessonia nigrescens**). The species most landed by the industrial sector in this area is also the Peruvian anchoveta (*Engraulis ringens*). Between the IV and VI regions, artisanals' most landed species are algaes and the Jumbo squid (*Dosidicus gigas*) while for industrials the msot important species in these regions are the Chilean jack mackerel (*Trachurus murphyi*), Squat lobsters (*Cervimunida johni*), Shrimps (*Heterocarpus reedi*) and the South pacific hake (*Merlucius gayi*), which is also the most landed species by artisanal fishers in the VII region. The VIII is by far the region with most landings from both the artisanal and industrial sector. In this region artisanals' most important species is the Araucanian herring (*Strangomera bentincki*) whereas for the industrial sector is the Chilean jack mackarel (*Trachurus murphyi*). In the IX region the Southern rays bream is the (*Brama australis*) most landed species by the artisanals and towards the south, from X to the XII region artisanals' most landed species are sea urchins (*Loxechinus albus*) and different algaes. The most important species for the industrial sector in the austral south are the Southern king crab (*Lithodes santolla*) and the Patagonian toothfish (*Dissostichus eleginoides*) in the X region, and the Patagonian grenadier (*Macruronus magellanicus*) in the XI region. 
 

```{r Landings per region, echo = FALSE, fig.cap= 'Mean annual landings per region for the artisanal (left pannel) and industrial sector (right pannel) between 2013 and 2017. Error bars represent one standard deviation. Note that scales are different between pannels. Data source: Landing records from SERNAPESCA, 2018.'}

# Main species in 2017 per region and sector in case we want to add color to the barplot

  main_species_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year, region, subsector) %>% 
  top_n(1, tons) %>% 
  filter(year ==2017)
  
 # Landings per region 
  
  landings_per_region <- landings %>% 
  filter(region != "national", international_waters == 0) %>% 
  mutate(tons = as.numeric(tons), subsector = ifelse(subsector =='fabric_vessel', 'industrial', subsector)) %>% 
  group_by(year, subsector, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(subsector, region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000)%>% 
  left_join(main_species_per_region, by = c('region', 'subsector')) %>% 
  select(subsector, region, annual_tons, sd, species)
  
# Reordering regions
 landings_per_region$region <- factor(landings_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

 landings_per_region <- landings_per_region %>% 
 ggplot(aes(x = region, y = annual_tons))+
  geom_bar(stat = 'identity', fill = wes_palette("Darjeeling1")[2] ) +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  facet_wrap(~ subsector, scales = 'free')+
  labs(y="Landings (Thousand Tons)", x="Region")+
  coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  ggtheme_plot()
 
 
# Main species per region artisanal sector overlast 5 years (Appendix)

 main_species_reg_art <- landings %>% 
  filter(region != "national", international_waters == 0, subsector == 'artisanal') %>% 
  mutate(tons = as.numeric(tons)) %>% 
  left_join(species_list) %>%
  group_by(year, region, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  filter(tons!= 0) %>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  select(-tons) %>% 
  spread(year, sc_name)
 
main_species_reg_art$region <- factor(main_species_reg_art$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))
 
# Main species per region industrial sector overlast 5 years (Appendix)

  main_species_reg_ind <- landings %>% 
  filter(region != "national", international_waters == 0, subsector == 'industrial') %>% 
  mutate(tons = as.numeric(tons)) %>% 
  left_join(species_list) %>%
  group_by(year, region, sc_name) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  filter(tons!= 0) %>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>% 
  select(-tons) %>% 
  spread(year, sc_name)
  
main_species_reg_ind$region <- factor(main_species_reg_ind$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

   
 landings_per_region

```

** Main farmed fisheries in terms of volume**

The main species harvested from aquaculture centers in the last 5 years have been the Atlantic salmon (*Salmo salar*), mussels (*Mytilus chilensis*), Coho salmon (*Oncorhynchus kisutch*) in descending order (Figure x), among others (S Table x). Overall, the composition of the most farmed species and its volumes have been stable over the last 5 years. 

```{r Most important farmed species in terms of volume in Chile, echo = FALSE, fig.cap= 'Top 5 species farmed in Chile between 2013 and 2017. Data source: Harvest records from SERNAPESCA, 2018.' }

# Importing aquaculture database
aquaculture<- read_csv(here("clean_databases/aquaculture_SERNAPESCA.csv"))

# Top aquaculture
  aquaculture_top_5 <-aquaculture %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  group_by(year, species) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>% 
  group_by(year) %>% 
  top_n(5, tons) %>% 
  left_join(species_list) %>% 
  select(-species) %>% 
  rename(species = sc_name) %>% 
  ggplot(aes(x = as.factor(year), y = tons/1000, fill = species))+
  geom_bar(stat = 'identity') +
  labs(y="Harvest (Thousand Tons)", x="Year")+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  ggtheme_plot()
  
# Most important farmed species per region (Appendix)
  
  aquaculture_species_per_region <-aquaculture %>% 
  filter(month == "anual") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  group_by(year, species, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  arrange(year, desc(tons)) %>%
  group_by(year, region) %>% 
  top_n(1, tons) %>%
  filter(tons != 0) %>% 
  left_join(species_list) %>% 
  select(-species, -tons) %>% 
  rename(species = sc_name) %>% 
  spread(year, species)
  
aquaculture_species_per_region$region <- factor(aquaculture_species_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

```

 ** Regional Diferences in farmed fisheries **

Aquaculture production is dramaticlly concentrated in the South of the country, between the X and the XII regions (Figure X). From the I to the IV region the most farmed resources are algaes and scallops (*Argopecten purpuratus*). The red abalone (*Haliotis rufescens*) became the most farmed in 2013 in the V region, replacing the Turbot (*Scophtalmus maximus*). Towards the south salmons are the most farmed species along with some algaes and mussels (*Mytilus chilensis*) (Table SX).

```{r Harvest from aquaculture centers per region, echo = FALSE, fig.cap= 'Mean annual harvest per region from aquaculture centers between 2012 and 2017. Error bars represent one standard deviation. Data source: Harvests records from SERNAPESCA, 2018.' }

 # Aquaculture per region 
  
  aquaculture_per_region <- aquaculture %>% 
  filter(region != "national") %>% 
  mutate(tons = as.numeric(tons)) %>% 
  group_by(year, region) %>% 
  summarise(tons = sum(tons, na.rm = T))%>% 
  group_by(region) %>% 
  summarise(annual_tons = mean(tons, na.rm = T)/1000, sd = sd(tons, na.rm = T)/1000)
  
  
# Reordering regions
  aquaculture_per_region$region <- factor(aquaculture_per_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

  aquaculture_per_region <- aquaculture_per_region %>% 
  ggplot(aes(x = region, y = annual_tons))+
  geom_bar(stat = 'identity', fill = wes_palette("Darjeeling1")[2] ) +
  geom_errorbar(aes(ymin=annual_tons-sd, ymax=annual_tons+sd), width=.2,
                 position=position_dodge(.9)) +
  labs(y="Harvest from aquaculture (Thousand Tons)", x="Region")+
  coord_flip()+
  scale_y_continuous(expand = c(0, 0))+
  ggtheme_plot()
  
  aquaculture_per_region

```

 ** Main Traded marine products **
 Imports
 Exports
 Net trade
 
Around 72% of the landings from fisheries and aquaculture are exported outside the country (@ifop_monitoreo_2015).

#### National Seafood Supply

<!--- Here, about the imports of animal products between regions https://es.datachile.io/products/productos-de-origen-animal-01 
Validar resultados con Informe subsectorial SUBPESCA
-->
** Main fisheries for food supply **
** Total and per capita food consumption in the country **
** Regional differences in per capita and total consumption **



#### National Economic Participation

##### Income  

The fisheries industry contributes to around 0.4% of the national GDP although it is a higly variable sector. During 2017 there was an increase in the sectors' GDP attributed to the wild caught of sardines and anchovies and to a lesser extent to farmed salmon (@central_cuentas_2017). This increase in fisheries production echoed in the manufacturing industry mainly via fish meal production (@central_cuentas_2017). The contribution of fisheries to regional GDP varies widely going for almost 30% in the XI region to 0% in the metropolitan (RM) and IX region (Fig. X). 

The XI is by far the one with more GDP coming from the fisheries sector. Supresingly this region is neither the one with higher landings nor the one with most aquaculture  



```{r Fisheries contirbution to regional and national PIB, echo = FALSE, fig.cap= 'Percent of the regional GDP contirbuted by the fisheries sector averaging data between 2013 and 2016. Data source: Banco Central de Chile.'}

PIB_region <- read.csv(here("information/raw_databases/Banco_Central/PIB_region.csv")) %>% 
  mutate(perc_contr = (PIB_fish * 100)/ PIB_total) %>% 
  group_by(region) %>% 
  summarise(mean_perc_contr = mean(perc_contr))

# Reordering regions
PIB_region$region <- factor(PIB_region$region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

PIB_region_plot <- ggplot(PIB_region, aes(x= as.factor(region), y= mean_perc_contr)) +
  geom_bar(stat = 'identity', fill = wes_palette("Darjeeling1")[2]) +
  coord_flip()+
  xlab('Region')+
  ylab('Contribution of fisheries industry to regional GDP (%)')+
  scale_y_continuous(expand = c(0, 0), breaks = c(seq(0,40, by = 5)))+
  #geom_label(aes(label = paste(as.character(round(mean_perc_contr,1)))), position = )+
  ggtheme_plot()

PIB_region_plot 

```

Fisheries products are one of the most important exported products of the country. IFOP valued the exportation of fisheries related products to $6,280 million dollars in 2017 (@ifop_boletines_2017). This corresponded to ~ 9 % of the total value of national exportations. The most important exported type of marine products in terms of value are frozen and fresh followed by fish meal and dried algae (@ifop_boletines_2017). 

##### Employment

In terms of employment we estimated a total of ~ 278,000 job positions generated by fisheries and aquaculture. This estimate considers artisanal and industrial fishers, workers in processing facilities and aquaculture centers and indirect jobs generated by the manufacturing sector (See Table X). It does not include indirect jobs generated by the extractive sector nor by the farming in aquaculutre centers. Thus, the number of total people directly and indireclty employed by fisheries and aquaculture in Chile could be higher. 


```{r Employment in fisheries from different sources, echo = FALSE, message=FALSE, warning= FALSE}

Artisanal_extraction <- c('88,968', '2018', 'RPA (SERNAPESCA, 2018)')
Industrial_extraction <- c('3,525', '2018', 'RPI (SERNAPESCA, 2018) and SUBPESCA website (2018)')
Aquaculture <- c('17,631', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Manufacturing <- c('65,451', '2017', 'Maturana et al., 2017 (SERNAPESCA, 2018)')
Indirect <- c('102,758', '2018', 'Own estimate based on IFOP,2015 multipliers' )

table_employment <- rbind(Artisanal_extraction, Industrial_extraction, Aquaculture, Manufacturing, Indirect) %>% 
  data.frame() %>% 
  mutate(Stage_supply_chain = c('Artisanal Extraction', 'Industrial Extraction', 'Aquaculture Centers', 'Manufacturing', 'Indirect employment'), Employees= X1, Year_estimate= X2, Source= X3)
table_employment <- table_employment [4:7]

knitr::kable(table_employment, col.names = c('Stage of the supply chain', 'Employees', 'Year of the estimate', 'Source'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most recent estimates of people employed in different stages of the supply chain of sea products in Chile. Note that the estimate of indirect employment only considers indirect employment generated by the manufacturing stage.') %>% 
kable_styling(latex_options = 'hold_position')


```

The artisanal sector correspond to fishing activities performed by vessels equal or smaller than 18 meters in length and by collectors in the inter-tidal zone (@subpesca_panorama_2018). Based on the current RPA, there are around 89,000 artisanal fishers along with 12,700 vessels in Chile. Figure X, shows the number of artisanal fishers per region and gender. They are concentrated in the X and VIII regions and are mostly men. Based on RPA information, the main activity in the artisanal sector is shore collection followed by fishing and to a lesser extent by diving. 

```{r Artisanal employment, echo= FALSE, fig.cap= 'Number of artisanal fishers in each region by gender. Source: National record of artisanal fishers (SERNAPESCA, 2018)'}

# Importing national records of artisanal fisheres
rpa <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "RPA PESCADORES", skip = 2, n_max = 138294) %>% 
  rename(region = "Regi\u00f3n", year = "A\u00f1o Inscripci\u00f3n")%>% 
  group_by(`RPA Pescador`, region, Genero) %>% 
  summarise(n = n()) %>%
  group_by(region, Genero) %>% 
  summarise(count= n()) %>%
  mutate(Region = ifelse(region ==15, 'XV', ifelse(region == 1, 'I', ifelse(region ==2, 'II', ifelse(region ==3, 'III', ifelse(region ==4, 'IV', ifelse(region == 5, 'V', ifelse(region ==6, 'VI', ifelse(region == 7, 'VII', ifelse(region ==8, 'VIII', ifelse(region == 9, 'IX', ifelse(region == 10, 'X', ifelse(region == 11, 'XI', ifelse(region == 12, 'XII', ifelse(region == 13, 'RM', ifelse(region == 14, 'XIV', 'XV')))))))))))))))) %>% 
  select(gender = Genero, count, Region) 

# Reordering regions
rpa$Region <- factor(rpa$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

rpa <-  ggplot(rpa, aes(x= as.factor(Region), y= count, fill = gender)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Region", x="Number of people", title = 'Artisanal fishers')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  ggtheme_plot()

rpa

```

The industrial sector comprises activities performed by vessels larger than 18 meters. Currently, there are 475 industrial vessels and 164 vessel owners according to the RPI, from which only 25 are individual owners and the rest are firms (@maturana_mujeres_2017). Besides, 3,500 jobs are generated through vessels operations (@subpesca_pesca_2018). Some of the firms that own industrial vessels also own processing facilities whose employees are considered as part of the manufacturing stage. 

A comprehensive study form IFOP estimated a total of ~50,000 job positions in the manufacturing of marine products in processing facilties (@ifop_monitoring_2015). However, a more recent report form SERNAPESCA reports a total of 65,451 jobs (@maturana_mujeres_2017). Figure X displays the geographic distribution of people employed in the manufacturing sector per gender and type of product for 2015. Eployment in the processing facilities is dominated by human consumption (HC) products, whith ~ 40,000 job positions. Products for animal consumption (AC) generated around 5,000 positions while algae products employed less than 3,000 people. Most of the people employed by the industrial sector is located in the southern part of the country and the majority are men. 


```{r Employment in facilities, echo= FALSE, fig.cap= 'Number of employees in the fisheries manufacturing sector in each region by gender in 2015. AC refers to products for animal consumption and HC products for human consumption. Source: Report by IFOP based on facilities survey (IFOP, 2015)}

# Importing IFOP, 2015 estimates of employees per type of facility

manufacturing_employment <- read.csv(here("information/raw_databases/IFOP/employment_per_processing_type_area.csv")) %>% 
  rename(M = ï..man_employee, F = women_employee, Product = product, Zone = region) %>% 
  gather(key = Gender, value = Employees, 1:2)

# Reordering regions
manufacturing_employment$Zone <- factor(manufacturing_employment$Zone , levels = c("South", "Mid South","North" ))

manufacturing_employment <- ggplot(manufacturing_employment, aes(x= as.factor(Zone), y= Employees, fill = Gender)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Employees", x="Zone", title = 'Manufacturing sector')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = wes_palette("Darjeeling1"))+
  facet_wrap(~Product, scales = 'free')+
  ggtheme_plot()
  
manufacturing_employment 

```

There are 17,631 people employed among the 3,683 aquaculture centers currently registered in Chile (@maturana_mujeres_2017). From these, ~4,500 are women. Figure X shows the distribution of aquaculture centers per region and type of farmed species. The main type of product being farmed is fish, mostly salmon, followed by molluscs and algae. The X region hosts more than half of the aquaculture centers in the country. 

```{r Aquaculture centers per region and type of species, echo= FALSE, fig.cap= 'Number of aquaculture centers in each region by type of species in 2017. Source: Report "Subsector Acuicultura 2017" by SERNAPESCA (SERNAPESCA's website, 2017)}

# Importing number of aquac. centers, SERNAPESCA

centers_aquaculture <- read.csv(here("information/raw_databases/SERNAPESCA/employment/centers_per_species_aquaculture.csv")) %>% 
  rename(Region = ï..Region) %>% 
  gather(key = Species, value = 'Centers', 2:7)

# Reordering regions
centers_aquaculture$Region <- factor(centers_aquaculture$Region , levels = c("XII", "XI", "X", "XIV", "IX", "VIII", "VII", "VI", "RM","V", "IV", "III", "II", "I", "XV"))

centers_aquaculture <- ggplot(centers_aquaculture, aes(x= as.factor(Region), y= Centers, fill = Species)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs(y="Centers", x="Regions", title = 'Aquaculture centers')+
  scale_y_continuous(expand = c(0, 0))+
  scale_fill_manual(values = c(wes_palette("Darjeeling1"), wes_palette("Darjeeling2")))+
  #facet_wrap(~Species, scales = 'free', ncol = 2)+
  ggtheme_plot()
  
centers_aquaculture 

```

Comercialization Indirect employment per xx estimated to be. 

** Species-specific estimates of economic contribution **

## Critical Analysis

- Algaes are important in Chile despite not being use for human consumption. They seem to be an important source of employment for artisanal fishers. We do not know much about algaes. 

### Key Messages

### Data uncertainty 

### Fisheries context

### Estimating National Seafood Supply

### Estimating National Economic Participation

- The number of unregistered fishers is low since being registered gives fishers
access to potential benefits like subsidies or territorial users’ rights for fishing. 


## Appendix

```{r Appendix, echo = FALSE, message= FALSE, warning= FALSE}

# Most landed species per region artisanal (for each year between 2013-2017)
knitr::kable(main_species_reg_art, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per region in the artisanal sector') %>% 
kable_styling(latex_options = 'hold_position')

# Most landed species per region industrial (for each year between 2013-2017)
knitr::kable(main_species_reg_ind, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per region in the industrial sector') %>% 
kable_styling(latex_options = 'hold_position')

# Most harvested species per region (for each year between 2012-2017)
knitr::kable(aquaculture_species_per_region, col.names = c('Region', 'Most landed in 2013', 'Most landed in 2014', 'Most landed in 2015', 'Most landed in 2016', 'Most landed in 2017'), format.args = list(decimal.mark = '.', big.mark = ","),"latex", booktabs = T, caption= 'Most landed species per region in the industrial sector') %>% 
kable_styling(latex_options = 'hold_position')

```




