---
title: "Estimating employment per fishery"
author: "Camila Vargas e Ignacia Rivera"
date: "November 19, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(here)
library(magrittr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(wesanderson)

```

## Employment per fishery in the artisanal sector

This is scripts uses the database provided by SERNAPESCA ( requested through the SIAC system) on the landings per vessel and species in the artisanal sector from 2007 to 2017. 

```{r importing raw database}
landings_art_per_vessel <- read_csv(here("information/raw_databases/SERNAPESCA/employment/landings_art_per_vessel.csv"), col_names = FALSE) %>% 
rename(cd_region = X1, nm_folio = X2, nm_tipo_nave = X3, nr_folio_impreso = X4, cd_nave = X5,  nm_nave = X6,  nr_eslora = X7, nr_trg = X8, nr_bodega = X9, nr_matricula = X10, cd_capitania = X11, fc_zarpe = X12,  fc_llegada = X13, nm_provincia = X14, nm_comuna = X15, nm_caleta = X16, nr_resolucion = X17, nm_ano = X18, nr_rpa_buzo = X19, cd_especie = X20, nm_especie = X21, nr_toneladas = X22, cd_arte =  X23 , nm_arte = X24, cd_zona = X25, cd_regimen =  X26, cd_tipo_destino = X27, cd_destino = X28)

```

## Main species landed in the artisanal sector per year 

The main species landed in the artisanal sector between 2010 and 2017 are shown in **Figure XX**. 

```{r main artisanal landings per species per year, echo=FALSE}

## selecting main landings per year, species and gear type for artisanal sector

main_art_species_year <- landings_art_per_vessel %>% 
  select(fc_llegada, nm_especie, nr_toneladas, nm_arte) %>% 
  separate(fc_llegada, c("month", "day", "year"), sep = "/") %>% 
  separate(year, c("year", "rest"), sep = " ") %>% 
  select(- rest) %>% 
  filter(year >2009 & year < 2018) %>% 
  group_by(year, nm_especie) %>% 
  #group_by(year, nm_especie, nm_arte) %>% #to include gear type
  summarise(tons= sum(nr_toneladas)) %>% 
  arrange(year, desc(tons)) %>% 
  group_by(year) %>% 
  top_n(5, tons) %>% 
  data.frame()

## plot 

plot_landings_artisanal <- ggplot(data = main_art_species_year, aes(x = year, y = tons, fill = nm_especie)) + 
    geom_bar(stat= "identity")+
    scale_y_continuous(expand = c(0, 0)) +
    labs(y="Landings (Tons)", x="Year")+
    theme_bw()+
    theme(axis.line = element_line(colour = "black"),
        axis.text=element_text(size=11),
        axis.title=element_text(size=12,face="bold"))+
  scale_fill_manual(name = "",
                    values = c('darkgreen', wes_palette("Darjeeling1", 5, type = "discrete"),"darkorchid4", "blue", "black","khaki4")) 


plot_landings_artisanal


```

## Species dependency pero boat each year

We consider the following information to extrapolate the number of fishers in each type of boat:

 - In Chile, ∼85% of this fleet is composed of small wooden, 7–10 m long, boats operated by a crew of three or four members (Castilla and Fernandez, 1998)

```{r species-dependency index per boat each year, echo=FALSE}

## calculating total tons per boat per year 

total_tons_boat <- landings_art_per_vessel %>% 
  select(cd_nave, nm_especie, nr_toneladas, fc_llegada) %>% 
  separate(fc_llegada, c("month", "day", "year"), sep = "/") %>% 
  separate(year, c("year", "rest"), sep = " ") %>% 
  select(- rest) %>% 
  filter(year >2009 & year < 2018) %>% 
  group_by(year, cd_nave) %>% 
  summarise(total_tons = sum(nr_toneladas)) 

## calculating species dependency index (tons sp / total tons) per boat per year 

sp_dependency_boat <- landings_art_per_vessel %>% 
  select(cd_nave, nm_especie, nr_toneladas, fc_llegada) %>% 
  separate(fc_llegada, c("month", "day", "year"), sep = "/") %>% 
  separate(year, c("year", "rest"), sep = " ") %>% 
  select(- rest) %>% 
  filter(year >2009 & year < 2018) %>% 
  group_by(cd_nave, year, nm_especie)%>%
  summarise(sp_tons = sum(nr_toneladas))%>% 
  left_join(total_tons_boat, by = c('cd_nave', 'year')) %>% 
  mutate(sp_dependency = sp_tons/total_tons) 

## database woth boats features


```
