---
title: 'Scope 2: Fisheries employment and income'
author: "Camila Vargas e Ignacia Rivera"
date: "August 17, 2018"
output: pdf_document
bibliography: Oceana.bib
csl: cell.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(magrittr)
library(tidyr)
library(reshape2)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(readxl)
```


## Top domestic marine, capture fisheries based on employment

### *Artisanal sector*

```{r rpa, echo = FALSE, message= FALSE, warning= FALSE}

## importing Chilean records of artisanal fishers (from SERNAPESCA, 2018)
rpa_fishers <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet= "RPA PESCADORES", skip = 2)

## importing Chilean records of artisanal vessels (from SERNAPESCA, 2018)
rpa_vessels <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet= "RPA EMBARCACIONES", skip = 2)

n_art_fishers <- length(unique(rpa_fishers$`RPA Pescador`))
n_art_vessels <- length(unique(rpa$RPA))
```


In Chile, there are `r n_art_fishers` artisanal fishers along with `r n_art_vessels` artisanal vessels, based SERNAPESCA (2018) records. Unregistered fishers are not common in Chile, hence this estimates are representative of the country's reality (REF.).   

```{r rpa per species}

## Importing rpa per species (SERNAPESCA, 2018)

### gets names of sheets
rpa_list <- list.files(here('information/raw_databases/SERNAPESCA/employment'), pattern= 'rpa_.*.csv', full.names = TRUE)

## Rrad, wrangle and bind all y datasets
rpa_species <- rpa_list %>% 
  map(read_delim,",", escape_double = FALSE, trim_ws = TRUE, col_names = c("species", "region"), skip =1) %>% 
  reduce(rbind)
  
rpa_species_plot <- rpa_all %>% 
  group_by(species, region) %>% 
  summarise(n = n()) %>% 
  arrange(region, desc(n)) %>% 
  group_by(region) %>% 
  top_n(3, n) %>% 
  data.frame() %>% 
   ggplot(., aes(x = region, y = n, fill = species)) + 
    geom_bar(stat= "identity") +
    scale_y_continuous(expand = c(0, 0)) +
    labs(y="Number of permits", x="Region") +
  theme(legend.position = "")+
  #scale_fill_brewer(name= 'BrBG')+
  ggtheme_plot()

rpa_species_plot

```

The main species for artisanal fishers in Chile are XXX if we consider the number of permits issued for different species. **Figure 1** shows the number of permits currently active for the main species in each region based on the 2018 RPA. 

If we use species-specific landings per vessel to evualte species relevance in employment in the artisanal sector (see Apendix for details), the main species are XXXX.

The artisanal sector generates employment beyond the extraction stage. After extraction, most resources are directly selled by the fishers in commercial facilities near their landing points, some are selled to local restaurants *Here we want to include a figure with the value chain of the artisanal catch and hopefully find a way to estimate the contirbution of different group of species to employment in different stages of the value process* 

### Industrial sector

```{r importing Registro Pescadores Industriales, echo=FALSE}

# Importing national records of industrial vessels
rpi <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "RPI", skip = 2, n_max = 472)

# Number of industrial vessels registered by 2018
number_ind_vessels <- length(unique(rpi$`RPI Nave`))

# Number of industrial vessel owners
number_ind_owners <- length(unique(rpi$`Nombre Armador`))
```

For the industrial sector, SERNAPESCA holds records only on the number of boats and its owners. There are currently `r number_ind_vessels` industrial vessels owned by `r number_ind_owners` different companies based on official registers. Estimates from SUBPESCA indicate that the overall operation of industrial vessels provide around 3,500 jobs nationwide (@subpesca_pesca_2018). 

*We have asked for data on species-specific landings or quotas for each industrial vessel. Using this data we will be able to assess how much each vessel has depended on particular species during the past years. We will use the size of the vessel as a proxy for how many people operate in it. Combining this two pieces of information we expect to have an estimate of how relevant are specific fisheries in providing job positions in the extractive industrial sector.*

Most of the catch from the industrial sector ends up going to processing facilities where it is transformed in products for domestic consumption, exportation or as supplies for the aquaculture industry. 

```{r Importing data on jobs in production facilities, echo = FALSE}

# Importing national records of job positions in processing facilities
mano_obra_plantas <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "MANO DE OBRA", skip = 2, n_max = 48140)
colnames(mano_obra_plantas) <- c('year', 'month', 'region', 'port', 'code', 'name', 'mo_perm_m', 'mo_perm_f', 'mo_temp_m', 'mo_temp_f', 'system')


# Average number of permanent and temporal job positions 
number_job_planta <- mano_obra_plantas %>% 
  group_by(year) %>% 
  summarise(mo_perm_m = max(mo_perm_m), mo_perm_f = max(mo_perm_f), mo_temp_m = max(mo_temp_m), mo_temp_f = max(mo_temp_f))


```

SERNAPESCA holds records on the self-reported number of permanent and temporal jobs in each facility. There were 7,751 permanent job positions and another 1,233 temporal positions in processing facilities in 2017. *We can do trends for this and split it by gender, but we first need to double check with SERNAPESCA some of these numbers that are too large to be realistic.* 

*To assess the contirbution of specific fisheries to employment in processing facilities we will use data we already have on (i) the production lines of each facility (e.g. fish meal, canned products, smoked products), (ii) tons of each species use to produce different products (e.g. fish meal, canned products, smoked products) in each region, (iii) the region in which each facility is located, (iv) the number of people working in each facility.*


## Top domestic marine, capture fisheries based on income 

### Artisanal sector

### Industrial sector 

# Cited literature