---
title: 'Scope 2: Fisheries employment and income'
author: "Camila Vargas e Ignacia Rivera"
date: "August 17, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(magrittr)
library(tidyr)
library(reshape2)
library(dplyr)
library(ggplot2)
```

## Introduction

In this section we characterize the top domestic, marine capture fisheries in Chile based on their impact on livelihoods. We use avaialable databases and information from reports to evaluate the impact of fisheries on the employment and income of those harvesting, processing and selling marine resources in Chile.



### Artisanal sector

There are currently around 92,000 artisanal fishers registered in SERNAPESCA along with 12,750 vessels. Artisanal fishers are requiered to register in the official record known as RPA (by its acronym in spanish). The RPA holds information regarding the gender, date of birth, location and type of activities performed by each fisher (e.i. diver, fisher, boat owner, shore gatherer). Although detailed, the RPA does not provide information on the specific cpecies targeted by each fisher making it impossible to determine the contribution of particular species to artisanal fishers livelihoods. To fullfil this information gap we wil.

The figure below shows the number of fishers in each region that perform a particualr activity. 
```{r}

# Importing national records of artisanal fisheres
rpa <- read_excel(here("information/raw_databases/SERNAPESCA/employment/RPA_RPI_mano_obra.xlsx"), sheet = "RPA PESCADORES", skip = 2, n_max = 138294)

# Number of boat owners, divers, fishers and gatherers in each region.

rpa_activity <- rpa %>% 
  select(region= 'Región', gender= Genero, year= 'Año Inscripción', ARMADOR, BUZO, RECOLECTOR, PESCADOR) %>% 
  gather(activity, practice, ARMADOR:PESCADOR) %>% 
  filter(practice == 'X') %>% 
  group_by(region, activity) %>% 
  summarize(count = n())

# Ploting by gender, activity and region 

ggplot(rpa_activity, aes(as.factor(region), count)) +   
  geom_bar(aes(fill = activity), position = "dodge", stat="identity")


```

### Industrial sector

Estimates from SERNAPESCA indicate that the operation of industrial vessels for fishing provide around 3,500 jobs in Chile. The industrial sector also includes processing facilities of marine products, for which 26,502 job positions are estimated by SERNAPESCA.

