---
title: 'Scope 1: Fisheries Context'
author: "Camila Vargas e Ignacia Rivera"
date: "August 16, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
```

## Introduction

The goal of this scope is to combine the data and information gathered from different sources to identify the top domestic fisheries contributing to the landings and national economy of Chile. The following script provides all the methods, information and data that is use to persue this goal. The script is separated in sections that refer to two differen obectives, the first assesing the contribution of different species to volume of landings and the second assesing value of landings. Sections are also based on the sources of data being used. 

## Main sources of information

We will identify the main domestic fisheries regarding volume and value using three different main data sources SERNAPESCA, Sea around us project and FAO estimates. 

## Requested outputs

1) A list of the top 10 domestic marine fisheries in terms of landed volume
2) A list of the top 10 domestic marine fisheries in terms of economic value
3) A clean database use to generate (1)
4) A database use to generate (4)

## Resolution

Where possible, these data should also be broken down by gear type.

## 1) Top domestic marine, capture fisheries based on landed volume

### Using data from SERNAPESCA

```{r importing clean landings SERNAPESCA per species, subsector, month and region}

landings<- read_csv(here("clean_databases/landings_SERNAPESCA.csv")) %>% 
  filter(region != "national") %>% 
  mutate( tons = as.numeric(tons))

```

Ten most important fisheries per region based on landings per region and year from 2014 to 2017

```{r}

##Artisanal

top10_landing_reg_artisanal <- landings %>%
  filter(subsector == "artisanal") %>% 
  arrange(year, region, desc(tons))

##Industrial

top10_landing_reg_industrial <- landings %>%
  filter(subsector == "industrial") %>% 
  arrange(year, region, desc(tons))
  
## frabric_vessels (ADD THEM TO INDUSTRIAL???)

top10_landing_reg_fv <- landings %>%
  filter(subsector == "fabric_vessel") %>% 
  arrange(year, region, desc(tons))

##WHAT TO DO DO WITH INTERNATIONAL WATERS?


```


Adding landings across all reagions to determine the main fisheries nationwise


```{r}



```





## Top ten fisheries per sector en each month based on data from 2014-2017 excluding international waters

### Example with fabric vessels
```{r main fisheries per month per sector}


ggplot() + 
geom_bar(data = filter(mean_landings_SERNAPESCA_month, subsector =='fabric_vessel'), aes(x = month, y= tons, fill = species), stat = 'identity')

```

